### 2014-07-16 bdcloud_paper/latex/paper.tex
Initial version of BDCloud Latex paper

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Domain-Driven Generative Data Model for BigPetStore}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Ronald J. Nowling}
\IEEEauthorblockA{Red Hat,\\
Raleigh, NC \\
Computer Science \& Engineering,\\
University of Notre Dame,\\
Notre Dame, IN 46617\\
Email: rnowling@redhat.com}
\and
\IEEEauthorblockN{Jay Vyas}
\IEEEauthorblockA{Red Hat\\
Raleigh, NC\\
Email: jvyas@redhat.com}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
Big Data applications process large, continuously-changing data sets and with the general goal of information and knowledge extraction.  Novel tools such as Pig, Spark, and Crunch based around Big Data frameworks are being developed every day.  These tools are often developed on desktops but need to be tested on clusters against large input sets.  The lack of frameworks for generating semantically-rich data at arbitrary scale denies developers access to consistent that can be used across a range of environments.

The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout from across the Hadoop and Spark ecosystems. BigPetStore has been incorporated into the Apache BigTop distribution for testing and as a reference application.

In this work, we present a novel domain-drive model for generating arbitrarily large, complex data for testing and benchmarking of software tools in the big data space and incorporate it into the BigPetStore application.   Our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter. Combining the features of the TeraGen (scalability) and MovieLens (semantically rich) input data sets commonly used for Big Data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.  To validate this approach, we apply it as a drop in replacement for the current version of BigPetStore which has already been utilized to test several Hadoop deployments at terabyte scale, and is actively being adopted by a variety of companies in the software industry, and confirm that it generates data with the same desirable properties (global patterns and region specific transactions) without some of the drawbacks (the need for hierarchical imperative coding of attribute selection during data set generation).

To benchmark, one must use inputs which are reliable, but also, which challenge a system in ways which approximate real-world usage.  Over time, more high-quality datasets are becoming publicly available, but these data sets often carry burdens that force developers to make undesirable trade offs for benchmarking. Few data sets are scalable, allowing the user to pick how much data they want.  As a result, users are forced to choose between small data sets which are useful for development and testing on a local machine, but not for production-scale tests or benchmarking, or large data sets which are prohibitively expensive to host, transfer, and store. Furthermore, access to large data sets can be unreliable due to dependence on third party benefactors for hosting and requires significant amounts of bandwidth to download.  Several companies (such as Twitter) offer access to their data but the costs and usage restrictions can be undesirable.

Models for generating synthetic data overcome these obstacles.  The code for generating the data is often small, enabling free hosting, licensing restrictions are often avoided, and the data size can be scaled, enabling seamless usage of the same data set across a range of scenarios.  TeraGen can randomly generate numerical data at any scale, but the data is semantically void and has very little meaning.  Thus, TeraGen cannot be used to test any application other than a simplistic “sort” program.  The Intel Hadoop Benchmark suite include several generative data models such as data for computing the PageRank of web sites for their benchmarks. The models are specific to each benchmark, however, and not useful for general-purpose development and testing.

Given that the driving motivation behind so much of the investment in big data software solutions is “to derive insight from data”, there is thus an obvious, gaping hole in the existing ecosystem: We need a tool that generates meaningful data which contains informational complexity that can be used to derive global patterns in a data set, which can be used to scalably and reliably generate large data sets on the fly - so that we can use it for benchmarking, comparison, and demonstration of the different algorithmic approaches to mining large data sets.

\section{Overview of Model \& Simulations}

\subsection{Stores}

Store records contain a unique identifier and a location in the form of a 5-digit zip code.  The locations of the stores are modeled by a PMF that gives the probability that a store’s location is the given zip code. We’ve written the PMF to give to zip codes in high-population, high-income areas the highest probabilities. The PMF is composed of two individual PMFs. One PMF determines the probability of each zip code as the population of that zip code over the total population:

The second PMF scales the probabilities of the zip codes by their incomes.  The zip code with the highest-income has a probability s-times larger than that of the lowest-income zip code. The values in-between are interpolated using an exponential function:

The combined PMF is given as:

where Z is the normalization factor.  Given the small size (~35,000 zip codes) of the data set, Z can be found directly by iterating over all of the zip codes and summing the scores.

The PMFs are parameterized using external data sets.  The list of zip codes came from a database of zip code latitude and longitudes.  The population and household income data [reference] for the zip codes were taken from from the U.S. Census.

The stores' locations are generated by sampling zip codes from the PMF.  A variety of methods can be used such as Roulette Wheel sampling and Monte Carlo methods.

\subsection{Customers}

Customer records consist of a unique identifier, a name, a location given as a 5-digit zip code, and the number and species of their pets.

The customers’ locations are modeled so that customers are placed within  The customer location is modeled as an exponential distribution over the distance to the nearest store.

[discuss parameters used in later sections, methods for setting number of customers per store]

\subsection{Transactions}

\begin{figure*}[!t]
  \centering
  \includegraphics[width=5.0in]{figures/transaction_modelseps.eps}
  \caption{Flowchart of Transaction Time \& Purchases Simulations}
  \label{fig:fig_sim}
\end{figure*}

\subsubsection{Transaction Times}
Transaction records consist of a customer, a store, a date and time, and a list of items purchased and their prices.

A transaction model is created for each customer and integrated over the simulation's timeframe to generate transactions.  Transactions are triggered when a customer runs out of items such as dog food as well as a random process describing the frequency of "impromptu" shopping trips.

Transaction times are modeled using a Monte Carlo method (Figure 1a).  The usage over time of each category is simulated, from which transaction times are found for each item category. The earliest transaction time is taken as the proposed transaction time.  The probability of the transaction time is calculated using a PDF.  If rejected, new transaction times are proposed.  Otherwise, the purchased items are modeled using a separate process (discussed below). After the items are chosen, the process begins again with the simulation of the item category usages.

In the first stage of the process, the usage of exhaustible items over time is modeled using a simple stochastic differential equation (SDE):

where $a_i$ is the amount of item category $i$ ``stuff'' remaining at time $t$. The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$. From the usage simulations, the exhaustion time for each exhaustible item category is found.

The transaction times are sampled from a collection of Poisson processes, one for each item category as well as one for impromptu shopping trips, modeling the arrival time relative to the exhaustion time.  In other words, the time at which the customer visits the store is found by sampling arrival times from an exponential distribution (parameterized separately for each item category) and subtracting the arrival times from the exhaustion times.  The earliest arrival time is taken as the proposed transaction time.

The acceptance probability of the proposed transaction time is modeled using a PDF that incorporates factors such as store hours.  The PDF could be extended to incorporate additional factors including time-dependent events such as sales, weather, and the amount of money the customer has available at the time.

Until the transaction time is accepted, the process repeatedly samples new transaction times from the Poisson processes. If the transaction time is accepted, a separate model (discussed below) is used to choose the items purchased.  Afterwards, the process starts over with the simulation of the items' usages. The process is repeated until a transaction later than the end time of the simulation is generated.

\subsubsection{Purchased Items}

A customer’s purchases in each transaction are modeled using a Hidden Markov Model (HMM) for the item categories layered on top of separate Markov Models for each item category (Figure 1b).  The current state is the combination of the customer’s current inventory and purchasing preferences. The observables are the item categories and a "stop" state. The observables are assigned probabilities based on which items the user is running out of (based on the usage simulations) and the average number of items purchased per transaction. The item category HMM could be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.

[figure]

After an item category is chosen from the HMM, a Markov model is used to model the customer's buying habits with respect to products in that item category.  Each product in that item category is represented as a state. The current state is that last item purchased in this category, regardless of which transaction the purchase occurred in. The transition probabilities are assigned according to the customer's buying habits.  For example, if a customer tends to buy the same brand of dog food repeatedly, the outgoing edges to all products of the same brand as the product in the current state will be higher than those for other brands' products. If the customer does not care about flavor, the outgoing edges to all products from the current brand will be equal, otherwise the edge weights can be adjusted to prefer a specific combination of brand and flavor.  Likewise, if a customer cares more about flavor or cost, the edge weights will be assigned appropriately.

[example MM]

[equation for determining edge weights]

To model a purchase, the item category Markov model is incremented by one state transition.  After an item is purchased, the customer's inventory is updated and the appropriate usage simulation is re-run.  Afterwards, the process repeats until the stop state is chosen in the HMM.

\section{Implementation}
Prototype using Python

\subsection{Output Format}

In line with the goal of generating data for analytics, the final output represents minimal information.  Although the model and simulation incorporate many internal parameters and state variables, these are ignored in the final output.  The final output consists of a table of transaction items like so:

* Customer Name
* Customer Location
* Store Location
* Transaction Number
* Transaction Date/time
* Item details (string)
* Item cost

[show example output]

\section{Example Analysis of Generated Data}
\subsection{Size, Scale}
[number of stores, number of customers, time window for simulations]

\subsection{Store Locations}
[locations, population, median household income]

\subsection{Customers}
[distances to stores]

\subsection{Transactions}

* stores -- distance, one or multiple stores?
* brand loyalty

\section{Availability}
GitHub, Apache license

\section{Discussion and Conclusion}
implementation
[want a modular framework with clearly-defined extension points, modularity so that users do not need to know about the machinations of the entire model, and so that modules are easy to develop and turn on/off dynamically]

[spark, hadoop ecosystem, parallelization of generating stores, customers, and transactions]

[model extensions]

% no \IEEEPARstart
This demo file is intended to serve as a ``starter file''
for IEEE conference papers produced under \LaTeX\ using
IEEEtran.cls version 1.7 and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
I wish you the best of success.

\hfill mds

\hfill January 11, 2007

\subsection{Subsection Heading Here}
Subsection text here.

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

% \section{Conclusion}
%The conclusion goes here.

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% that's all folks
\end{document}

### 2014-07-16 bdcloud_paper/latex/paper.tex
Updated figures.  Add equations.  Edited model description

@@
 Store records contain a unique identifier and a location in the form of a 5-digit zip code.
-The locations of the stores are modeled by a PMF that gives the probability that a store’s location is the given zip code.
-We’ve written the PMF to give to zip codes in high-population, high-income areas the highest probabilities.
+The locations of the stores are modeled by a PMF that gives the probability that a store's location is the given zip code.
+We've written the PMF to give to zip codes in high-population, high-income areas the highest probabilities.
 The PMF is composed of two individual PMFs.
 One PMF determines the probability of each zip code as the population of that zip code over the total population:
+
+\begin{equation}
+p(\text{location}=z | \text{population}(z)) = \frac{\text{population}(z)}{\sum_{i} \text{population}(i)}
+\end{equation}
 
@@
 
+\begin{equation}
+p(\text{location}=z | \text{income}(z)) = s ^ {\big( \frac{\text{income} - \min_i{\textrm{income}(i)}} {\max_i{\textrm{income}(i)} - \min_i{\textrm{income}(i)}} - 1 \big)}
+\end{equation}
+
 The combined PMF is given as:
 
-where Z is the normalization factor.  Given the small size (~35,000 zip codes) of the data set, Z can be found directly by iterating over all of the zip codes and summing the scores.
+\begin{align}
+p(\text{location}=z) = &Z^-1 p(\text{location}=z | \text{population}(z)) \nonumber \\
+&p(\text{location}=z | \text{income}(z))
+\end{align}
+
+where $Z$ is the normalization factor.  Given the small size ($\approx$30,000 zip codes) of the data set, $Z$ can be found directly by iterating over all of the zip codes and summing the scores.
 
@@
 
-\begin{figure*}[!t]
+\begin{figure}[!t]
   \centering
-  \includegraphics[width=5.0in]{figures/transaction_modelseps.eps}
-  \caption{Flowchart of Transaction Time \& Purchases Simulations}
-  \label{fig:fig_sim}
-\end{figure*}
+  \includegraphics[width=3.75in]{figures/transaction_simulation.eps}
+  \caption{Flowchart of Transaction Monte Carlo Simulation}
+  \label{fig:trans_sim}
+\end{figure}
 
@@
 A transaction model is created for each customer and integrated over the simulation's timeframe to generate transactions.
-Transactions are triggered when a customer runs out of items such as dog food as well as a random process describing the frequency of "impromptu" shopping trips.
+Transactions are triggered when a customer runs out of items such as dog food as well as a random process describing the frequency of ``impromptu'' shopping trips.
 
-Transaction times are modeled using a Monte Carlo method (Figure 1a).
+Transaction times are simulated using a Monte Carlo method (Figure~\ref{fig:trans_sim}).
 The usage over time of each category is simulated, from which transaction times are found for each item category.
@@
 
+\begin{equation}
+\frac{da_i}{dt} = -\mu_i + \sigma_i\, dW_i(t)
+\end{equation}
+
 where $a_i$ is the amount of item category $i$ ``stuff'' remaining at time $t$.
 The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$.
-From the usage simulations, the exhaustion time for each exhaustible item category is found.
+The exhaustion time $ T_{E,i}$ for each item category is found from the usage simulations.
 
-The transaction times are sampled from a collection of Poisson processes, one for each item category as well as one for impromptu shopping trips, modeling the arrival time relative to the exhaustion time.
-In other words, the time at which the customer visits the store is found by sampling arrival times from an exponential distribution (parameterized separately for each item category) and subtracting the arrival times from the exhaustion times.
-The earliest arrival time is taken as the proposed transaction time.
+The proposed transaction time is found from the exhaustion times.
+For each item cateogry, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an Exponential distribution parameterized separately for each customer $c$.
+A proposed transaction time $T_{P, i}$ for each item category is found by subtracting the offset time $T_{O, i}$ from the exhaustion time $T_{E, i}$.
+The earlist proposed transaction time is taken as the overall proposed transaction time $T_T$.
 
-The acceptance probability of the proposed transaction time is modeled using a PDF that incorporates factors such as store hours.
+\begin{align}
+&\beta_c \sim \text{U}(a, b) \\
+&T_{O, i} \sim \text{Exp}(\beta^{-1}_c) \\
+&T_{P, i} = T_{E,i} - T_{P, i} \\
+&T_T = \min_i \, \{  T_{P, i}\}
+\end{align}
+
+The acceptance probability $p(t=T_T)$ of the proposed transaction time $T_T$ is modeled using a PDF that incorporates factors such as store hours.
 The PDF could be extended to incorporate additional factors including time-dependent events such as sales, weather, and the amount of money the customer has available at the time.
 
-Until the transaction time is accepted, the process repeatedly samples new transaction times from the Poisson processes.
-If the transaction time is accepted, a separate model (discussed below) is used to choose the items purchased.
-Afterwards, the process starts over with the simulation of the items' usages.
-The process is repeated until a transaction later than the end time of the simulation is generated.
+\begin{equation}
+p(t=T_T) = p(t=T_T | \text{store hours})
+\end{equation}
 
-\subsubsection{Purchased Items}
+The proposed transaction time is accepted if $p(t=T_T) < r$, where $r \sim \text{U}(0, 1)$.
+Until the proposed transaction time is accepted, a new proposed transaction time is generated by sampling new offset times.
+If the proposed transaction time is accepted, the shopping cart of items is generated through a separate simulation discussed below.
+Transactions are generated until the accepted transaction time is later than the end time given as a simulation parameter.
 
-A customer’s purchases in each transaction are modeled using a Hidden Markov Model (HMM) for the item categories layered on top of separate Markov Models for each item category (Figure 1b).
-The current state is the combination of the customer’s current inventory and purchasing preferences.
-The observables are the item categories and a "stop" state.
-The observables are assigned probabilities based on which items the user is running out of (based on the usage simulations) and the average number of items purchased per transaction.
-The item category HMM could be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.
+\begin{figure*}[!t]
+  \centering
+  \includegraphics[width=6in]{figures/shopping_cart_simulation.eps}
+  \caption{Shopping Cart Hidden Markov Model}
+  \label{fig:shopping_cart_sim}
+\end{figure*}
 
-[figure]
+\subsubsection{Shopping Cart Items}
+
+A customer’s purchases in each transaction are modeled using a layered Hidden Markov Model (HMM) (Figure~\ref{fig:shopping_cart_sim}).
+The HMM has three states: the start state, the purchase states, and the stop state.
+There are an infinite number of purchase states, identified by the exhaustion times of the customer's item categories, the transaction time, and the number of items purchased in the current transaction.
+The observables of the purchase states correspond to the item categories and are Markov models for choosing the item to be purchased from the category.
+The observables are assigned probabilities based on the amounts of time (based on the usage simulations) until the customer uses the remaining amount of each item category.
+
+Markov Models are used to model the customer's purchasing behavior for each item category.
+Each state corresponds to an item in that category.
+The transition probabilities are defined so-as to model the customer's buying preferences.
+
+\begin{equation}
+Pr(X_{n+1}=x|X_n) = \dots
+\end{equation}
 
@@
 
+The item category HMM could be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.
+
+[figure]
+
 [example MM]
@@
 Afterwards, the process repeats until the stop state is chosen in the HMM.
+
+\subsubsection{Other Fields}
 

### 2014-07-16 bdcloud_paper/latex/paper.tex
Add description of numerical integration of SDE

@@
 % http://www.tug.org/applications/pdftex
+
+\usepackage{color}
 
@@
 \author{\IEEEauthorblockN{Ronald J. Nowling}
-\IEEEauthorblockA{Red Hat,\\
-Raleigh, NC \\
+\IEEEauthorblockA{Red Hat, Inc.\\
+Raleigh, NC 27601\\
 Computer Science \& Engineering,\\
@@
 \IEEEauthorblockN{Jay Vyas}
-\IEEEauthorblockA{Red Hat\\
-Raleigh, NC\\
+\IEEEauthorblockA{Red Hat, Inc.\\
+Raleigh, NC 27601\\
 Email: jvyas@redhat.com}}
@@
 
-\subsubsection{Transaction Times}
+\subsubsection{Transaction Times} \label{sec:transaction-times}
 Transaction records consist of a customer, a store, a date and time, and a list of items purchased and their prices.
@@
 
-In the first stage of the process, the usage of exhaustible items over time is modeled using a simple stochastic differential equation (SDE):
+In the first stage of the process, the usage of items over time is modeled using a simple stochastic differential equation (SDE):
 
@@
 The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$.
-The exhaustion time $ T_{E,i}$ for each item category is found from the usage simulations.
+The SDE is integrated using the Euler-Maruyama method \textcolor{red}{reference} with timestep $\Delta t_n$ sampled from a Poisson process:
+
+\begin{align}
+&a_{n+1, i} = a_{n,i} - \mu_{i, c} \, \Delta t_n + \sigma_{i, c} \, \sqrt{\Delta t_n} R_n \\
+&\Delta t_n \sim \text{Exp}(\lambda_i) \nonumber \\
+&R_n \sim \text{N}(0, 1) \nonumber
+\end{align}
+
+where $\lambda_i$ is the item category's average usage rate, $\mu_i$ is the item category's average amount used per use, and $\sigma^2_i$ is the variance of the amount of the item category used per use.
+The parameter $\lambda_i$ is defined for each item cateogry.
+The parameters $\mu_{i, c}$, and $\sigma^2_{i,c}$ are found based on a base rate for each item category multiplied by the number of pets of the appropriate species the customer has.
+The exhaustion time $T_{E,i}$ for each item category is found by simulating the usage until $a_n = 0$.
 
 The proposed transaction time is found from the exhaustion times.
-For each item cateogry, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an Exponential distribution parameterized separately for each customer $c$.
+For each item cateogry, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an exponential distribution parameterized separately for each customer $c$.
 A proposed transaction time $T_{P, i}$ for each item category is found by subtracting the offset time $T_{O, i}$ from the exhaustion time $T_{E, i}$.
@@
 The observables of the purchase states correspond to the item categories and are Markov models for choosing the item to be purchased from the category.
-The observables are assigned probabilities based on the amounts of time (based on the usage simulations) until the customer uses the remaining amount of each item category.
-
-Markov Models are used to model the customer's purchasing behavior for each item category.
-Each state corresponds to an item in that category.
-The transition probabilities are defined so-as to model the customer's buying preferences.
+The observables are \textcolor{red}{assigned probabilities} based on the amounts of time (based on the usage simulations) until the customer uses the remaining amount of each item category.
 
@@
 
-After an item category is chosen from the HMM, a Markov model is used to model the customer's buying habits with respect to products in that item category.
-Each product in that item category is represented as a state.
-The current state is that last item purchased in this category, regardless of which transaction the purchase occurred in.
+Markov Models are used to model the customer's purchasing behavior for each item category.
+Each state corresponds to an item in that category.
 The transition probabilities are assigned according to the customer's buying habits.
@@
 Likewise, if a customer cares more about flavor or cost, the edge weights will be assigned appropriately.
+To simulate an item purchase, the Markov model is transitioned forward by one state.
+The current states of the Markov models are kept between purchases.
 
-The item category HMM could be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.
+\begin{equation}
+Pr(X_{n+1}=x|X_n) = \dots
+\end{equation}
 
-[figure]
+After an item is chosen, the item category's exhaustion time is updated by propogating the item category usage SDE described in Section~\ref{sec:transaction-times}.
+\textcolor{red}{state transition, stop state}
 
-[example MM]
-
-[equation for determining edge weights]
-
-To model a purchase, the item category Markov model is incremented by one state transition.
-After an item is purchased, the customer's inventory is updated and the appropriate usage simulation is re-run.
-Afterwards, the process repeats until the stop state is chosen in the HMM.
+The HMM could be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.
 

### 2014-07-16 bdcloud_paper/latex/paper.tex
Added missing word
Added transaction introduction

@@
 \subsection{Transactions}
+A transaction model is created for each customer and integrated over the simulation's timeframe to generate transactions.
+Transaction records consist of a customer, a store, a date and time, and a list of items purchased and their prices.
+The store is set to the store located closest to the customer.
+The date and time is given as a floating-point number with units of days since the start of the simulation.
+(Timezones are not taken into account.)  Transaction times are generated from a Monte Carlo process based on the customer will use up items such as dog food (Section~\ref{sec:transaction-times}).
+The purchased items are generated by a Hidden Markov Model (HMM) parameterized by the transaction time and amount of time remaining for the items in the customer's inventory (Section~\ref{sec:transaction-purchases}).
 
@@
 \subsubsection{Transaction Times} \label{sec:transaction-times}
-Transaction records consist of a customer, a store, a date and time, and a list of items purchased and their prices.
-
-A transaction model is created for each customer and integrated over the simulation's timeframe to generate transactions.
-Transactions are triggered when a customer runs out of items such as dog food as well as a random process describing the frequency of ``impromptu'' shopping trips.
 
@@
 The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$.
-The SDE is integrated using the Euler-Maruyama method \textcolor{red}{reference} with timestep $\Delta t_n$ sampled from a Poisson process:
+The SDE is numerically integrated using the Euler-Maruyama method \textcolor{red}{reference} with timestep $\Delta t_n$ sampled from a Poisson process:
 
@@
 
-\subsubsection{Shopping Cart Items}
+\subsubsection{Shopping Cart Items} \label{sec:transaction-purchases}
 
@@
 
-The HMM could be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.
-
-\subsubsection{Other Fields}
+The HMM can be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.
 

### 2014-07-16 bdcloud_paper/latex/paper.tex
Add implementation text

@@
 
-\section{Overview of Model \& Simulations}
+\section{Overview of Models \& Simulations}
 
@@
 \section{Implementation}
-Prototype using Python
-
-\subsection{Output Format}
-
-In line with the goal of generating data for analytics, the final output represents minimal information.
-Although the model and simulation incorporate many internal parameters and state variables, these are ignored in the final output.
-The final output consists of a table of transaction items like so:
-
-* Customer Name
-* Customer Location
-* Store Location
-* Transaction Number
-* Transaction Date/time
-* Item details (string)
-* Item cost
-
-[show example output]
+The models and simulations were implemented using Python.
+The source code is available on GitHub \textcolor{red}{reference} at \url{https://github.com/rnowling/bigpetstore-data-generator} under the Apache 2 license.
 
@@
 
-\section{Availability}
-GitHub, Apache license
-
 \section{Discussion and Conclusion}
@@
 % no \IEEEPARstart
-This demo file is intended to serve as a ``starter file''
-for IEEE conference papers produced under \LaTeX\ using
-IEEEtran.cls version 1.7 and later.
+%This demo file is intended to serve as a ``starter file''
+%for IEEE conference papers produced under \LaTeX\ using
+%IEEEtran.cls version 1.7 and later.
 % You must have at least 2 lines in the paragraph with the drop letter
 % (should never be an issue)
-I wish you the best of success.
+%I wish you the best of success.
 
-\hfill mds
+%\hfill mds
 
-\hfill January 11, 2007
-
-\subsection{Subsection Heading Here}
-Subsection text here.
-
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
+%\hfill January 11, 2007
 
@@
 \bibitem{IEEEhowto:kopka}
+%\subsection{Subsection Heading Here}
+%Subsection text here.
+
+%\subsubsection{Subsubsection Heading Here}
+%Subsubsection text here.
 H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus

### 2014-07-16 bdcloud_paper/latex/paper.tex
Add initial benchmark data

@@
 \section{Example Analysis of Generated Data}
-\subsection{Size, Scale}
-[number of stores, number of customers, time window for simulations]
+Transaction data for 10 stores, 10,000 customers, and 4 years of simulated time totaling 100 MB was generated using the Python implementation.
 
@@
 * brand loyalty
+
+\section{Scaling of Data Size and Run Time}
+BigPetStore aims to scale from a local desktop to a large cluster.
+As a result, we benchmarked the generation of data on a laptop with a 2 GHz Intel Core i7 CPU, 8 GB of RAM, and a 256 GB SSD using the Python implementation with a single thread.
+Using the test setup, between 1,500 and 2,000 transactions can be generated per second (Table~\ref{tab:benchmarks}).
+
+\begin{table}[!t]
+%% increase table row spacing, adjust to taste
+\renewcommand{\arraystretch}{1.3}
+% if using array.sty, it might be a good idea to tweak the value of
+% \extrarowheight as needed to properly center the text within the cells
+\caption{Benchmarks of Data Generator}
+\label{tab:benchmarks}
+\centering
+%% Some packages, such as MDW tools, offer better commands for making tables
+%% than the plain LaTeX2e tabular which is used here.
+\begin{tabular}{|p{0.75cm}||p{1.2cm}||p{1cm}||p{1.25cm}||p{1cm}||p{1cm}|}
+\hline
+Stores & Customers & Simulated Time (years) & Transactions & Data Size (MB) & Run Time (min)\\ \hline
+10 & 10,000 & 1 & 279,870 & 18 & 3.5 \\ \hline
+%10 & 100,000 & 1 & 
+100 & 10,000 & 1 & 279,586 & 18 & 3.5 \\ \hline
+% 100 & 100,000 & 1 & 
+10 & 1,000 & 5 & 123,309 & 8 & 1.1 \\ \hline
+100 & 1,000 & 5 & 127,064 & 8 & 1.2 \\ \hline
+10 & 10,000 & 5 & 1,275,542 & 84 & 11.0 \\ \hline
+%10 & 100,000 & 5 & 
+100 & 10,000 & 5 & 1,268,403 & 84 & 10.5 \\ \hline
+\end{tabular}
+\end{table}
+
+The number of transactions (and hence, data size) grows as $O(N_c T)$ where $N_c$ is the number of customers and $T$ is the amount of time to be simulated.
+The amount of data generated can be scaled as large as necessary simply by increasing $N_c$ and $T$.
+Five years of transactions for 100,000 customers will generate approximately 1 GB of data.
+A terabyte of data can be generated by simulating 100 million customers over five years.
 

### 2014-07-16 bdcloud_paper/latex/paper.tex
Edited introduction.  Added placeholders for references

@@
 \section{Introduction}
-Big Data applications process large, continuously-changing data sets and with the general goal of information and knowledge extraction.
-Novel tools such as Pig, Spark, and Crunch based around Big Data frameworks are being developed every day.
-These tools are often developed on desktops but need to be tested on clusters against large input sets.
-The lack of frameworks for generating semantically-rich data at arbitrary scale denies developers access to consistent that can be used across a range of environments.
+Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.
+New tools such as Hive, Pig and Crunch \textcolor{red}{reference} based around big data frameworks such as Hadoop, Storm, and Spark \textcolor{red}{reference} are being released every day.
+Development, testing and benchmarking of big data software, algorithms, and infrastructure pose unique challenges.
+Though software and algorithms will be deployed to large clusters for benchmarking and production usage, engineers develop and test these tools on desktops due to the speed and ease of development and debugging on a local machine and limited access to large clusters due to the high costs of purchasing, provisioning, and maintaining such clusters.
+Software and algorithms are being developed at incredible speed with new releases everyday.
+The high speed of development driven by the need to be the first to market in a competitive field often results in buggy software.
+There is great need for consistent and rigorous testing and benchmarking frameworks to aid in the identification of bugs while maintaining the high-speed of development.
+Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key to increasing access to and development of consistent and rigorour testing and benchmarking frameworks.
 
-The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout from across the Hadoop and Spark ecosystems.
-BigPetStore has been incorporated into the Apache BigTop distribution for testing and as a reference application.
+Over time, more high-quality data sets are becoming publicly available, but these data sets often carry burdens that force developers to make undesirable trade offs.
+Since few data sets are scalable, allowing the user to pick how much data they want, users are forced to choose between small data sets which are useful for development and testing on a local machine, but not for production-scale tests or benchmarking, or large data sets which are prohibitively expensive to host, transfer, and store.
+Furthermore, access to large data sets can be unreliable due to dependence on third party benefactors for hosting and requires significant amounts of bandwidth to download.
+Several companies (such as Twitter \textcolor{red}{reference}) offer access to their data but the costs and usage restrictions can be undesirable.
 
-In this work, we present a novel domain-drive model for generating arbitrarily large, complex data for testing and benchmarking of software tools in the big data space and incorporate it into the BigPetStore application.
-Our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter.
-Combining the features of the TeraGen (scalability) and MovieLens (semantically rich) input data sets commonly used for Big Data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.
-To validate this approach, we apply it as a drop in replacement for the current version of BigPetStore which has already been utilized to test several Hadoop deployments at terabyte scale, and is actively being adopted by a variety of companies in the software industry, and confirm that it generates data with the same desirable properties (global patterns and region specific transactions) without some of the drawbacks (the need for hierarchical imperative coding of attribute selection during data set generation).
+\emph{Ab initio} mathematical models for generating synthetic data overcomes several of these obstacles.
+The code for generating the data is often small, enabling free hosting, licensing restrictions are often avoided, and the data size can be scaled, enabling seamless usage of the same data set across a range of scenarios.
+Common synthetic data set generators such as TeraGen and the Intel Hadoop Benchmark \textcolor{red}{reference} can generate data sets quickly and at any scale, but they are over-specialized for benchmarking, resulting in semantically-void data that lacks usefulness for analytics examples.
+The lack of frameworks for generating semantically-rich data at arbitrary scale denies developers access to consistent data sets that can be used across a range of environments \textcolor{red}{Huppler, Pavlo}.
 
-To benchmark, one must use inputs which are reliable, but also, which challenge a system in ways which approximate real-world usage.
-Over time, more high-quality datasets are becoming publicly available, but these data sets often carry burdens that force developers to make undesirable trade offs for benchmarking.
-Few data sets are scalable, allowing the user to pick how much data they want.
-As a result, users are forced to choose between small data sets which are useful for development and testing on a local machine, but not for production-scale tests or benchmarking, or large data sets which are prohibitively expensive to host, transfer, and store.
-Furthermore, access to large data sets can be unreliable due to dependence on third party benefactors for hosting and requires significant amounts of bandwidth to download.
-Several companies (such as Twitter) offer access to their data but the costs and usage restrictions can be undesirable.
+Attempts have been made to find a middle-ground.
+Myriad and PDGF provide primitives for specifying generative data models \textcolor{red}{reference}.
+Alexandrov, et al.
+proposed Oligos, an approach for generating synthetic data from models trained on reference databases with real data \textcolor{red}{reference}.
+Oligos appears promising but will need to overcome several difficult challenges in automatically characterizing complex reference data to determine relevant constraints and learn accurate models.
+Until such methods have reached maturity, development of a customized domain-driven model is best suited for generating for satisfying the needs of big data software and algorithms development.
 
-Models for generating synthetic data overcome these obstacles.
-The code for generating the data is often small, enabling free hosting, licensing restrictions are often avoided, and the data size can be scaled, enabling seamless usage of the same data set across a range of scenarios.
-TeraGen can randomly generate numerical data at any scale, but the data is semantically void and has very little meaning.
-Thus, TeraGen cannot be used to test any application other than a simplistic “sort” program.
+The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout from across the Hadoop and Spark ecosystems \textcolor{red}{reference}.
+BigPetStore's usefulness has been demonstrated by its incorporation into the Apache BigTop distribution for testing and as a reference application \textcolor{red}{reference}.
+
+In this work, we present a new domain-drive model and simulation for generating arbitrarily large, complex data for testing and benchmarking of software tools in the big data space.
+Compared with BigPetStore's previous data generator, our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter.
+Combining the features of the TeraGen (scalability) \textcolor{red}{reference} and MovieLens (semantically rich) \textcolor{red}{reference} input data sets commonly used for big data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.
+To validate this approach, we apply it as a drop in replacement for the current version of BigPetStore which has already been utilized to test several Hadoop deployments at terabyte scale, and is actively being adopted by a variety of companies in the software industry, and confirm that it generates data with desirable properties (global patterns and region specific transactions).
+
+%Models for synthetic data generation is not the only approach for obtaining data sets. 
+%can randomly generate numerical data at any scale, but the data is semantically void and has very little meaning.
+Thus, TeraGen cannot be used to test any application other than a simplistic sort program.
 The Intel Hadoop Benchmark suite include several generative data models such as data for computing the PageRank of web sites for their benchmarks.
@@
 
-Given that the driving motivation behind so much of the investment in big data software solutions is “to derive insight from data”, there is thus an obvious, gaping hole in the existing ecosystem: We need a tool that generates meaningful data which contains informational complexity that can be used to derive global patterns in a data set, which can be used to scalably and reliably generate large data sets on the fly - so that we can use it for benchmarking, comparison, and demonstration of the different algorithmic approaches to mining large data sets.
+%Given that the driving motivation behind so much of the investment in big data software solutions is ``to derive insight from data,'' there is thus an obvious, gaping hole in the existing ecosystem: We need a tool that generates meaningful data which contains informational complexity that can be used to derive global patterns in a data set, which can be used to scalably and reliably generate large data sets on the fly - so that we can use it for benchmarking, comparison, and demonstration of the different algorithmic approaches to mining large data sets.
 

### 2014-07-16 src/python/item_simulations.py
Fixed integration of SDE

@@ -37,23 +37,19 @@ class ItemCategoryUsageSimulation(object):
         """
         
         # given in days since last usage
-        timestep = random.expovariate(1.0 / self.daily_usage_rate)
+        timestep = random.expovariate(self.daily_usage_rate)
         
-        # Might be more realistic to model usage
-        # in units/time used since items are used
-        # in discrete rather than continuous
-        # quantities
+
+        r = random.normalvariate(0.0, 1.0)
         
         # given in units/day
-        usage_rate = random.normalvariate(self.amount_used_average, self.amount_used_variance)
+        usage_amount = self.amount_used_average * timestep \
+            + np.sqrt(self.amount_used_variance * timestep) * r
         
         # can't use a negative amount :)
-        if usage_rate < 0.0:
-            usage_rate = 0.0
-        
-        # given in units
-        usage_amount = usage_rate * np.sqrt(timestep)
-        
+        if usage_amount < 0.0:
+            usage_amount = 0.0
+                
         self.remaining_amount -= min(usage_amount, self.remaining_amount)
         
         self.time += timestep

### 2014-07-16 src/python/customers.py
Moved transaction and purchase trigger times to customer.  Moved parameters to separate module

@@ -8,10 +8,14 @@ import numpy as np
 
 from samplers import RouletteWheelSampler
 
+import simulation_parameters as sim_param
+
 class Customer(object):
     def __init__(self):
         self.name = None
         self.location = None
+        self.average_transaction_trigger_time = None
+        self.average_purchase_trigger_time = None
         self.pets = {
                         "dog" : 0,
                         "cat" : 0
@@ -23,7 +27,7 @@ class Customer(object):
              self.pets["cat"], self.location)
 
 class LocationSampler(object):
-    def __init__(self, stores=None, zipcode_objs=None, avg_distance=5.0):
+    def __init__(self, stores=None, zipcode_objs=None, avg_distance=None):
         lambd = 1.0 / avg_distance
         self.stores = stores
         self.zipcode_objs = zipcode_objs
@@ -74,7 +78,8 @@ class LocationSampler(object):
 class CustomerGenerator(object):
     def __init__(self, zipcode_objs=None, stores=None):
         self.location_sampler = LocationSampler(stores=stores,
-                                                zipcode_objs=zipcode_objs)
+                                                zipcode_objs=zipcode_objs,
+                                                avg_distance=sim_param.AVERAGE_CUSTOMER_STORE_DISTANCE)
 
 
     def generate(self, n):
@@ -84,9 +89,22 @@ class CustomerGenerator(object):
             customer.name = "Customer_" + str(i)
             customer.location = self.location_sampler.sample()
             
-            num_pets = random.randint(1, 10)
+            num_pets = random.randint(sim_param.MIN_PETS, sim_param.MAX_PETS)
             num_dogs = random.randint(0, num_pets)
             num_cats = num_pets - num_dogs
+
+            # days
+            r = random.normalvariate(sim_param.TRANSACTION_TRIGGER_TIME_AVERAGE,
+                                     sim_param.TRANSACTION_TRIGGER_TIME_VARIANCE)
+            r = max(r, sim_param.TRANSACTION_TRIGGER_TIME_MIN)
+            r = min(r, sim_param.TRANSACTION_TRIGGER_TIME_MAX)
+            customer.average_transaction_trigger_time = r
+
+            r = random.normalvariate(sim_param.PURCHASE_TRIGGER_TIME_AVERAGE,
+                                     sim_param.PURCHASE_TRIGGER_TIME_VARIANCE)
+            r = max(r, sim_param.PURCHASE_TRIGGER_TIME_MIN)
[…]

### 2014-07-16 src/python/item_simulations.py

@@ -95,9 +95,10 @@ class ItemCategorySimulation(object):
         self.daily_usage_rate = item_category.daily_usage_rate
         self.amount_used_average = item_category.base_amount_used_average * num_pets
         self.amount_used_variance = item_category.base_amount_used_variance * num_pets
+
+        self.average_transaction_trigger_time = customer.average_transaction_trigger_time
+        self.average_purchase_trigger_time = customer.average_purchase_trigger_time
         
-        self.transaction_trigger_rate = item_category.transaction_trigger_rate
-        self.transaction_purchase_rate = item_category.transaction_purchase_rate
         self.sim = None
                         
     def record_purchase(self, purchase_time, purchased_amount):
@@ -134,11 +135,12 @@ class ItemCategorySimulation(object):
 
     def purchase_weight(self, time):
         remaining_time = max(self.exhaustion_time() - time, 0.0)
-        lambd = 1.0 / self.transaction_purchase_rate
+        lambd = 1.0 / self.average_purchase_trigger_time
         return lambd * np.exp(-lambd * remaining_time)
         
     def propose_transaction_time(self):
-        time_until_transaction = random.expovariate(1.0 / self.transaction_trigger_rate)
+        lambd = 1.0 / self.average_transaction_trigger_time
+        time_until_transaction = random.expovariate(lambd)
         transaction_time = max(self.exhaustion_time() - time_until_transaction, 0.0)
         return transaction_time
 

### 2014-07-16 src/python/simulation_parameters.py

@@ -1,0 +1,17 @@
+# parameters for simulation
+
+TRANSACTION_TRIGGER_TIME_AVERAGE = 5.0 # days
+TRANSACTION_TRIGGER_TIME_VARIANCE = 2.0
+TRANSACTION_TRIGGER_TIME_MAX = 10.0
+TRANSACTION_TRIGGER_TIME_MIN = 1.0
+
+PURCHASE_TRIGGER_TIME_AVERAGE = 10.0
+PURCHASE_TRIGGER_TIME_VARIANCE = 4.0
+PURCHASE_TRIGGER_TIME_MAX = 20.0
+PURCHASE_TRIGGER_TIME_MIN = 1.0
+
+AVERAGE_CUSTOMER_STORE_DISTANCE = 5.0 # miles
+
+MAX_PETS = 10
+MIN_PETS = 1
+

### 2014-07-16 src/python/simulation.py
Moved store income scaling factor into params module

@@ -1,5 +1,6 @@
 from customers import CustomerGenerator
 from customer_simulation import CustomerState
+import simulation_parameters as sim_param
 from stores import StoreGenerator
 import products
 from transactions import TransactionSimulator
@@ -16,7 +17,7 @@ class Simulator(object):
 
     def generate_stores(self, num=None):
         generator = StoreGenerator(zipcode_objs=self.zipcode_objs,
-                                   income_scaling_factor=100.0)
+                                   income_scaling_factor=sim_param.STORE_INCOME_SCALING_FACTOR)
         self.stores = generator.generate(n=num)
 
     def generate_customers(self, num=None):
@@ -109,13 +110,13 @@ if __name__ == "__main__":
         store_writer.append(store)
 
     print "Generating customers..."
-    sim.generate_customers(num=1000)
+    sim.generate_customers(num=100)
 
     for customer in sim.customers:
         customer_writer.append(customer)
 
     print "Generating transactions..."
-    for trans in sim.generate_transactions(end_time=365.0*4.0):
+    for trans in sim.generate_transactions(end_time=365.0*1.0):
         trans_writer.append(trans)
 
     print

### 2014-07-16 src/python/simulation_parameters.py

@@ -15,3 +15,4 @@ AVERAGE_CUSTOMER_STORE_DISTANCE = 5.0 # miles
 MAX_PETS = 10
 MIN_PETS = 1
 
+STORE_INCOME_SCALING_FACTOR = 100.0

### 2014-07-16 src/python/stores.py

@@ -3,7 +3,7 @@ from samplers import RouletteWheelSampler
 import numpy as np
 
 class ZipcodeSampler(object):
-    def __init__(self, zipcode_objs, income_scaling_factor=100.0):
+    def __init__(self, zipcode_objs, income_scaling_factor=None):
 
         pop_probs = dict()
         income_probs = dict()
@@ -57,7 +57,7 @@ class Store(object):
 
 
 class StoreGenerator(object):
-    def __init__(self, zipcode_objs=None, income_scaling_factor=100.0):
+    def __init__(self, zipcode_objs=None, income_scaling_factor=None):
         self.zipcode_objs = zipcode_objs
         self.zipcode_sampler = ZipcodeSampler(zipcode_objs=zipcode_objs,
                                                 income_scaling_factor=income_scaling_factor)

### 2014-07-16 bdcloud_paper/latex/paper.tex
Added description of HMM observable probabilities

@@
 
-\section{Overview of Models \& Simulations}
+\section{Overview of Model \& Simulation Procedures}
+
+\textcolor{red}{need an intro paragrpah}
 
@@
   \centering
-  \includegraphics[width=3.75in]{figures/transaction_simulation.eps}
+  \includegraphics[width=3.5in]{figures/transaction_simulation.eps}
   \caption{Flowchart of Transaction Monte Carlo Simulation}
@@
 &T_{O, i} \sim \text{Exp}(\beta^{-1}_c) \\
-&T_{P, i} = T_{E,i} - T_{P, i} \\
+&T_{P, i} = T_{E,i} - T_{O, i} \\
 &T_T = \min_i \, \{  T_{P, i}\}
@@
   \includegraphics[width=6in]{figures/shopping_cart_simulation.eps}
-  \caption{Shopping Cart Hidden Markov Model}
+  \caption{Example Shopping Cart Hidden Markov Model}
   \label{fig:shopping_cart_sim}
@@
 
-A customer’s purchases in each transaction are modeled using a layered Hidden Markov Model (HMM) (Figure~\ref{fig:shopping_cart_sim}).
+A customer's purchases in each transaction are modeled using a layered Hidden Markov Model (HMM) (Figure~\ref{fig:shopping_cart_sim}).
 The HMM has three states: the start state, the purchase states, and the stop state.
 There are an infinite number of purchase states, identified by the exhaustion times of the customer's item categories, the transaction time, and the number of items purchased in the current transaction.
-The observables of the purchase states correspond to the item categories and are Markov models for choosing the item to be purchased from the category.
-The observables are \textcolor{red}{assigned probabilities} based on the amounts of time (based on the usage simulations) until the customer uses the remaining amount of each item category.
+The observables of the states correspond to the item categories.
+Each item category $i$ is assigned a weight $w_i$ according to the amount of time between the transaction time $T_t$ and $T_{E,i}$ when each item category will be exhausted (based on the usage simulations):
 
 \begin{equation}
-Pr(X_{n+1}=x|X_n) = \dots
+w_i = - \lambda_i \exp(-\lambda_i (T_{E, i}- T_{T}))
 \end{equation}
+
+The probability $p(X=i)$ of choosing item category $i$ is given by the normalizing the weight $w_i$ by the sum of all of the weights $W$:
+
+\begin{align}
+&p(X=i) = W^{-1} w_i \\
+&W = \sum_i w_i \nonumber
+\end{align}
+
+The item category for the purchase is chosen using a Roulette Wheel sampler.
 

### 2014-07-16 bdcloud_paper/latex/paper.tex
starting adding references

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Domain-Driven Generative Data Model for BigPetStore}
+\title{Domain-Driven, Generative Data Model for BigPetStore}
 
@@
 The lack of frameworks for generating semantically-rich data at arbitrary scale denies developers access to consistent data sets that can be used across a range of environments \textcolor{red}{Huppler, Pavlo}.
+\textcolor{red}{standard benchmarks}
 
@@
 The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout from across the Hadoop and Spark ecosystems \textcolor{red}{reference}.
-BigPetStore's usefulness has been demonstrated by its incorporation into the Apache BigTop distribution for testing and as a reference application \textcolor{red}{reference}.
+BigPetStore's usefulness has been demonstrated by its incorporation into the Apache BigTop distribution \cite{BigTop} for testing and as a reference application.
 
@@
 The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$.
-The SDE is numerically integrated using the Euler-Maruyama method \textcolor{red}{reference} with timestep $\Delta t_n$ sampled from a Poisson process:
+The SDE is numerically integrated using the Euler-Maruyama method\cite{Klouden13} with timestep $\Delta t_n$ sampled from a Poisson process:
 
@@
 Until the proposed transaction time is accepted, a new proposed transaction time is generated by sampling new offset times.
-If the proposed transaction time is accepted, the shopping cart of items is generated through a separate simulation discussed below.
+If the proposed transaction time is accepted, the shopping cart of items is generated through a separate simulation discussed in Section~\ref{sec:transaction-purchases} below.
 Transactions are generated until the accepted transaction time is later than the end time given as a simulation parameter.
@@
 The models and simulations were implemented using Python.
-The source code is available on GitHub \textcolor{red}{reference} at \url{https://github.com/rnowling/bigpetstore-data-generator} under the Apache 2 license.
+The source code is available on GitHub \cite{GitHub} at \url{https://github.com/rnowling/bigpetstore-data-generator} under the Apache 2 license.
 
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
+\bibliographystyle{plain}
 % argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
+%\bibliography{IEEEabrv,paper}
+\bibliography{paper}
 %
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
+%\begin{thebibliography}{1}
 
-\bibitem{IEEEhowto:kopka}
+%\bibitem{IEEEhowto:kopka}
 %\subsection{Subsection Heading Here}
@@
 %Subsubsection text here.
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
+%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
 
-\end{thebibliography}
+%\end{thebibliography}
 

### 2014-07-16 bdcloud_paper/latex/paper.tex
small edits.  add references

@@
 Store records contain a unique identifier and a location in the form of a 5-digit zip code.
-The locations of the stores are modeled by a PMF that gives the probability that a store's location is the given zip code.
-We've written the PMF to give to zip codes in high-population, high-income areas the highest probabilities.
-The PMF is composed of two individual PMFs.
+The locations of the stores are modeled by a PDF that gives the probability that a store's location is the given zip code.
+We designed the PDF to give to zip codes in high-population, high-income areas the highest probabilities.
+The PDF is composed of two individual PMFs.
 One PMF determines the probability of each zip code as the population of that zip code over the total population:
@@
 
-The second PMF scales the probabilities of the zip codes by their incomes.
-The zip code with the highest-income has a probability s-times larger than that of the lowest-income zip code.
+The second PDF scales the probabilities of the zip codes by their incomes.
+The zip code with the highest-income has a probability $s$-times larger than that of the lowest-income zip code.
 The values in-between are interpolated using an exponential function:
@@
 
-The combined PMF is given as:
+The combined PDF is given as:
 
@@
 
-The PMFs are parameterized using external data sets.
-The list of zip codes came from a database of zip code latitude and longitudes.
-The population and household income data [reference] for the zip codes were taken from from the U.S.
-Census.
+The PDFs are parameterized using external data sets.
+The zip codes' latitudes and longitudes came from the Zip Code Database Project \cite{Zips}.
+The population and household income data for the zip codes were taken from from the U.S.
+Census American Community Survey \cite{ACS}.
 
-The stores' locations are generated by sampling zip codes from the PMF.  A variety of methods can be used such as Roulette Wheel sampling and Monte Carlo methods.
+The stores' locations are generated by sampling zip codes from the PDF.  A variety of methods can be used such as Roulette Wheel sampling and Monte Carlo methods.
 

### 2014-07-17 src/python/transactions.py
Choose MSM weights differently to create separation of customer preferences

@@ -21,9 +21,12 @@ class ItemCategoryMarkovModelBuilder(object):
         self.field_weights = dict()
         self.field_similarity_weights = dict()
         for field in self.item_category.fields:
-            self.field_weights[field] = random.choice([0.1, 0.25, 0.5, 0.75, 0.9])
-            self.field_similarity_weights[field] = random.choice([0.1, 0.25, 0.5, 0.76, 0.9])
-        self.loopback_weight = random.choice([0.25, 0.5, 0.9])
+            avg = random.choice([0.15, 0.85])
+            self.field_weights[field] = min(0.95, max(0.05, random.normalvariate(avg, 0.1)))
+            avg = random.choice([0.15, 0.85])
+            self.field_similarity_weights[field] = min(0.95, max(0.05, random.normalvariate(avg, 0.1)))
+        avg = random.choice([0.25, 0.75])
+        self.loopback_weight = min(0.95, max(0.05, random.normalvariate(avg, 0.1)))
 
     def similarity_weight(self, rec1, rec2):
         weight = 0.0

### 2014-07-18 bdcloud_paper/latex/paper.tex
Updated paper with clustering figure
Added text for example analyses

@@
 %\boldmath
-The abstract goes here.
+Abstract
 \end{abstract}
@@
 \section{Example Analysis of Generated Data}
-Transaction data for 10 stores, 10,000 customers, and 4 years of simulated time totaling 100 MB was generated using the Python implementation.
+Using data for 10 stores, 10,000 customers, and five years of transactions generated from the model, we performed two example analyses in support of a fictional advertising campaign.
+The analyses were designed to be realistic and driven by real-world business concerns.
+Due to limited advertising budgets, businesses need to decide who to target and what sort of advertisements are likely to be effective for each type of customer.
+We analyzed data generated from the model to determine where the advertising campaigns should be targeted geographically and identify customer purchasing profiles.
 
-\subsection{Store Locations}
-[locations, population, median household income]
+First, we wanted to identify how close most customers live to a store since advertising to people who live or work too far away from a store is unlikely to be effective and will increase our costs.
+We analyzed the distribution of distances between customers' homes and their nearest stores.
+We found that most customers tend to live within 10 miles of their closest store, suggesting that we should limit our advertising to a 10-mile radius around each our store.
 
-\subsection{Customers}
-[distances to stores]
+\textcolor{red}{figure}
 
-\subsection{Transactions}
+Secondly, we profiled our customers' purchasing habits to optimize the advertising campaign's effectivness by customizing the advertized products for each customer.
+For each customer, we generated a feature vector by computing the frequency with which they would purchase the same brand or flavor from one transaction to the next.
+We clustered the customers' feature vectors using the KMean algorithms with a range of cluster counts.
+The clustering identified 16 centers, determined by the convergence of the error.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=3.5in]{figures/cluster_analysis.eps}
+  \caption{Clustering of Brand and Flavor Purchasing Preferences}
+  \label{fig:cluster_analysis}
+\end{figure}
+
+Analysis of the feature vectors and clusters revealed four predominant purchasing profiles (Figure~\ref{fig:cluster_analysis}).
+High frequencies of purchasing the same flavors repeatedly were tightly-correlated with high frequencies of purchasing the same brands -- these customers were likely to be happy with a particular item and kept purchasing the same item repeatedly.
+For our advertising campaign, it would be unlikely to get these customers to purchase different items, so we should create incentives to purchase larger quantites, especially when inventory levels are high and needed to be depleted.
+Other customers had a tendency to purchase either the same flavor or brand repeatedly, but varied in their choice of the other.
+For customers who prefer a particular brand, we should target our advertising campaign to suggest other flavors sold by that brand.
+Likewise, for customers who prefer a particular flavor, we should target our advertising campaign to suggest other brands with that flavor.
+Lastly, some customers had very low frequencies of purchasing neither the same brand nr flavor.
+These customers may be driven by other factors, such as cost, not represented in this analysis and will need further study.
+
+\subsection{Customer Locations}
+
+\subsection{Customer Purchasing Profiles}
+
+%\begin{figure*}[!t]
+%  \centerline{\subfloat[Case I}{\includegraphics[width=2.5in]{figures/cluster_error.eps}
+%    \label{fig:cluster_error}}
+%  \hfil
+%  \subfloat[Case II]{\includegraphics[width=2.5in]{figures/cluster_analysis.eps}
+%    \label{fig:cluster_analysis}}}
+%  \caption{Clustering of Brand and Flavor Preferences}
+%  \label{fig:clustering}
+%\end{figure*}
 

### 2014-07-18 bdcloud_paper/latex/paper.tex
Small edits of analysis

@@
 For each customer, we generated a feature vector by computing the frequency with which they would purchase the same brand or flavor from one transaction to the next.
-We clustered the customers' feature vectors using the KMean algorithms with a range of cluster counts.
+We clustered the customers' feature vectors using the KMean algorithm with a range of cluster counts.
 The clustering identified 16 centers, determined by the convergence of the error.
@@
 Lastly, some customers had very low frequencies of purchasing neither the same brand nr flavor.
-These customers may be driven by other factors, such as cost, not represented in this analysis and will need further study.
+Other factors, such as cost, not represented in this analysis may be driving these customers' purchasing habits and will need further study.
 

### 2014-07-18 bdcloud_paper/latex/paper.tex
Updated cluster analysis for 10,000 customers

@@
 We clustered the customers' feature vectors using the KMean algorithm with a range of cluster counts.
-The clustering identified 16 centers, determined by the convergence of the error.
+Twenty clusters converged the error.
 
@@
 
-\subsection{Customer Locations}
-
-\subsection{Customer Purchasing Profiles}
-
-%\begin{figure*}[!t]
-%  \centerline{\subfloat[Case I}{\includegraphics[width=2.5in]{figures/cluster_error.eps}
-%    \label{fig:cluster_error}}
-%  \hfil
-%  \subfloat[Case II]{\includegraphics[width=2.5in]{figures/cluster_analysis.eps}
-%    \label{fig:cluster_analysis}}}
-%  \caption{Clustering of Brand and Flavor Preferences}
-%  \label{fig:clustering}
-%\end{figure*}
-
-* stores -- distance, one or multiple stores?
-* brand loyalty
-
 \section{Scaling of Data Size and Run Time}
@@
 Using the test setup, between 1,500 and 2,000 transactions can be generated per second (Table~\ref{tab:benchmarks}).
+As the customers' transaction simulations are independent of one another, the transaction generation can easily be parallelized so that 1,500-2,000 transactions can be generated per thread per second.
 

### 2014-07-18 bdcloud_paper/latex/paper.tex
Add customer-store distance data

@@
 We analyzed the distribution of distances between customers' homes and their nearest stores.
-We found that most customers tend to live within 10 miles of their closest store, suggesting that we should limit our advertising to a 10-mile radius around each our store.
+We found that most customers tend to live within 15 miles of their closest store, suggesting that we should limit our advertising to a 15-mile radius around each our store.
 
-\textcolor{red}{figure}
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=3.5in]{figures/customer_store_distances.eps}
+  \caption{Distributions of Distances of Customers to Closest Stores}
+  \label{fig:cluster_analysis}
+\end{figure}
 

### 2014-07-18 bdcloud_paper/latex/paper.tex
Added description of customre locations and pets

@@
 Store records contain a unique identifier and a location in the form of a 5-digit zip code.
-The locations of the stores are modeled by a PDF that gives the probability that a store's location is the given zip code.
+The locations of the stores are modeled by a probability density function (PDF) that gives the probability that a store's location is the given zip code.
 We designed the PDF to give to zip codes in high-population, high-income areas the highest probabilities.
-The PDF is composed of two individual PMFs.
-One PMF determines the probability of each zip code as the population of that zip code over the total population:
+The PDF is composed of two individual PDFs.
+One PDF determines the probability of each zip code as the population of that zip code over the total population:
 
@@
 
-The stores' locations are generated by sampling zip codes from the PDF.  A variety of methods can be used such as Roulette Wheel sampling and Monte Carlo methods.
+The stores' locations are generated by sampling zip codes from the PDF using a Roulette Wheel sampler.
 
@@
 
-The customers’ locations are modeled so that customers are placed within  The customer location is modeled as an exponential distribution over the distance to the nearest store.
+\textcolor{red}{names}
 
-[discuss parameters used in later sections, methods for setting number of customers per store]
+For each proposed customer location zipcode $z$, we compute the distance between $z$ and every store's zipcode to find the closest store.
+We compute the distance $d$ between $z$ and the zipcode $s_z$ of the closest store.
+The weight $w_z$ for each zipcode $z$ is modeled using an exponential distribution with the average distance $\beta$.
+
+\begin{align}
+&w_z = \beta^{-1} \exp(-\beta^{-1} \, d_m(z)) \\
+&d_m(z) = \min_{s_z} \, d(z, s_z) \nonumber
+\end{align}
+
+To find the probabilities, the weights are normalized:
+
+\begin{align}
+&p(\text{location} = \text{z}) = W^{-1} w_z \\
+&W = \sum_{z} w(z) \nonumber
+\end{align}
+
+The customer locations are sampled from the PDF using a Roulette Wheel sampler.
+
+We determine the number of pets $N_p$ each customer has by sampling from a uniform distribution of integers.
+We then sample the number of dogs $N_d$ as a uniform distribution between 0 and $N_p$.
+The remaining number of pets are assigned to be cats.
+
+\begin{align}
+N_p \sim U(1, b) \\
+N_d \sim U(0, N_p) \nonumber \\
+N_c = N_p - N_d \nonumber
+\end{align}
 

### 2014-07-18 bdcloud_paper/latex/paper.tex
Created evaluation section

@@
 
-We determine the number of pets $N_p$ each customer has by sampling from a uniform distribution of integers.
-We then sample the number of dogs $N_d$ as a uniform distribution between 0 and $N_p$.
+We determine the number of pets $N_p$ each customer has by sampling from a discrete uniform distribution of integers.
+We then sample the number of dogs $N_d$ as a discrete uniform distribution of integers between 0 and $N_p$.
 The remaining number of pets are assigned to be cats.
@@
 
-\section{Example Analysis of Generated Data}
+\section{Evaluation of the Model}
+
+\textcolor{red}{intro paragraph}
+
+\subsection{Example Analysis of Generated Data}
 Using data for 10 stores, 10,000 customers, and five years of transactions generated from the model, we performed two example analyses in support of a fictional advertising campaign.
@@
 
-\section{Scaling of Data Size and Run Time}
+\subsection{Scaling of Data Size and Run Time}
 BigPetStore aims to scale from a local desktop to a large cluster.

### 2014-07-19 bdcloud_paper/latex/paper.tex
Added references to introduction

@@
 
-\emph{Ab initio} mathematical models for generating synthetic data overcomes several of these obstacles.
-The code for generating the data is often small, enabling free hosting, licensing restrictions are often avoided, and the data size can be scaled, enabling seamless usage of the same data set across a range of scenarios.
-Common synthetic data set generators such as TeraGen and the Intel Hadoop Benchmark \textcolor{red}{reference} can generate data sets quickly and at any scale, but they are over-specialized for benchmarking, resulting in semantically-void data that lacks usefulness for analytics examples.
+Methods for generating synthetic data offer multiple advantages over pre-generated data sets.
+The code for generating the data is often small, making it easy and cheap to host, transfer, and store.
+Licensing restrictions and data privacy issues are avoided \textcolor{red}{cite}.
+And, the size of the data sets can be scaled arbitrarily, enabling seamless usage of the same data set across a range of situations.
+
+Multiple synethic data set generators exist.
+TeraGen \textcolor{red}{cite} and the Intel Hadoop Benchmark Suite \cite{Huang2010} can generate data sets quickly and at any scale.
+Targetted towards specific benchmarks such as Map-Reduce jobs and Hive queries, the resulting data is semantically-void and not useful for other tasks.
+The PGDF (Parallel Data Generation Framework) can quickly generate data and queries that structurally satifies a user-defined relational database schema \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011}.
+The Myriad Toolkit improves on PGDF through a more descriptive model specification language \cite{Alexandrov2012}.
+Myriad allows the user to specify the distributions for sampling values, simple arithmetic equations for setting the value of fields from other fields, and restrictions on fields domains.
+Due to the inability to specify underlying dynamic models with latent variables, the data generated by PGDF and Myriad lacks informational complexity and genuine semantics, making it inapprioriate for analytics.
+
+More advanced approaches such as Oligas \cite{Alexandrov2013} and SMOTE \cite{Chawla2002} attempt to generate or extend real data sets with synthetic data sets that satisfy characteristics learned from a real reference data sets.
+\textcolor{red}{privacy, accuracy without details of underlying process, SMOTE - feature space} Oligos appears promising but will need to overcome several difficult challenges in automatically characterizing complex reference data to determine relevant constraints and learn accurate models.
+
+Until such methods have reached maturity, the development of customized, domain-driven models is best suited for generating for satisfying the needs of big data software and algorithms development.
+For example, the Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006} was developed for the Department of Homeland Security to develop synthetic data for effectiveness of IDASs designed to identify potential terrorism threat scenarios.
+
+.
 The lack of frameworks for generating semantically-rich data at arbitrary scale denies developers access to consistent data sets that can be used across a range of environments \textcolor{red}{Huppler, Pavlo}.
@@
 
-Attempts have been made to find a middle-ground.
-Myriad and PDGF provide primitives for specifying generative data models \textcolor{red}{reference}.
-Alexandrov, et al.
-proposed Oligos, an approach for generating synthetic data from models trained on reference databases with real data \textcolor{red}{reference}.
-Oligos appears promising but will need to overcome several difficult challenges in automatically characterizing complex reference data to determine relevant constraints and learn accurate models.
-Until such methods have reached maturity, development of a customized domain-driven model is best suited for generating for satisfying the needs of big data software and algorithms development.
+The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout from across the Hadoop and Spark ecosystems \textcolor{red}{reference}.
+BigPetStore's usefulness has been demonstrated by its incorporation into the open-source Apache BigTop distribution \cite{BigTop} for testing and as a reference application.
 
-The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout from across the Hadoop and Spark ecosystems \textcolor{red}{reference}.
-BigPetStore's usefulness has been demonstrated by its incorporation into the Apache BigTop distribution \cite{BigTop} for testing and as a reference application.
-
-In this work, we present a new domain-drive model and simulation for generating arbitrarily large, complex data for testing and benchmarking of software tools in the big data space.
+In this work, we present a new domain-driven model and simulation for generating arbitrarily large, complex data for testing and benchmarking of algorithms, software, and infrastructure in the big data space.
 Compared with BigPetStore's previous data generator, our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter.

### 2014-07-19 bdcloud_paper/latex/paper.tex
Small edits to intro

@@
 There is great need for consistent and rigorous testing and benchmarking frameworks to aid in the identification of bugs while maintaining the high-speed of development.
-Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key to increasing access to and development of consistent and rigorour testing and benchmarking frameworks.
+Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key to increasing access to and development of consistent and rigorous testing and benchmarking frameworks.
 
@@
 More advanced approaches such as Oligas \cite{Alexandrov2013} and SMOTE \cite{Chawla2002} attempt to generate or extend real data sets with synthetic data sets that satisfy characteristics learned from a real reference data sets.
-\textcolor{red}{privacy, accuracy without details of underlying process, SMOTE - feature space} Oligos appears promising but will need to overcome several difficult challenges in automatically characterizing complex reference data to determine relevant constraints and learn accurate models.
+\textcolor{red}{privacy, accuracy without details of underlying process, SMOTE - feature space} Oligos appears promising but will need to overcome several difficult challenges in automatically characterizing complex reference data to determine relevant constraints and learn accurate models that capture the effect of latent variables.
 
-Until such methods have reached maturity, the development of customized, domain-driven models is best suited for generating for satisfying the needs of big data software and algorithms development.
+Until such methods have reached maturity, the development of customized, domain-driven models is best suited for satisfying the needs of big data software and algorithms development.
 For example, the Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006} was developed for the Department of Homeland Security to develop synthetic data for effectiveness of IDASs designed to identify potential terrorism threat scenarios.
 
-.
-The lack of frameworks for generating semantically-rich data at arbitrary scale denies developers access to consistent data sets that can be used across a range of environments \textcolor{red}{Huppler, Pavlo}.
-\textcolor{red}{standard benchmarks}
+\textcolor{red}{Huppler, Pavlo}.  \textcolor{red}{standard benchmarks}
 
@@
 Combining the features of the TeraGen (scalability) \textcolor{red}{reference} and MovieLens (semantically rich) \textcolor{red}{reference} input data sets commonly used for big data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.
-To validate this approach, we apply it as a drop in replacement for the current version of BigPetStore which has already been utilized to test several Hadoop deployments at terabyte scale, and is actively being adopted by a variety of companies in the software industry, and confirm that it generates data with desirable properties (global patterns and region specific transactions).
+To validate this approach, we perform an example analysis designed to guide a fictional advertising campaign (a typical business use case) to demonstrate that the model generates data with desirable properties (global patterns and region specific transactions).
+We have made the model's implementation available as open source software.
 

### 2014-07-19 bdcloud_paper/latex/paper.tex
Added more references

@@
 
-Multiple synethic data set generators exist.
+Several synethic data set generators exist.
 TeraGen \textcolor{red}{cite} and the Intel Hadoop Benchmark Suite \cite{Huang2010} can generate data sets quickly and at any scale.
 Targetted towards specific benchmarks such as Map-Reduce jobs and Hive queries, the resulting data is semantically-void and not useful for other tasks.
-The PGDF (Parallel Data Generation Framework) can quickly generate data and queries that structurally satifies a user-defined relational database schema \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011}.
-The Myriad Toolkit improves on PGDF through a more descriptive model specification language \cite{Alexandrov2012}.
-Myriad allows the user to specify the distributions for sampling values, simple arithmetic equations for setting the value of fields from other fields, and restrictions on fields domains.
-Due to the inability to specify underlying dynamic models with latent variables, the data generated by PGDF and Myriad lacks informational complexity and genuine semantics, making it inapprioriate for analytics.
+Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005}.
+The more advanced approaches of these allow users to specify relationships between field values through simple arithmetic equations for setting the value of fields from other fields \cite{Alexandrov2012} or use constraint-solving techniques \cite{Arasu2011}.
+However, due to the inability to specify underlying dynamic models with latent variables, the generated data lacks informational complexity and genuine semantics, making it inapprioriate for analytics.
 
-More advanced approaches such as Oligas \cite{Alexandrov2013} and SMOTE \cite{Chawla2002} attempt to generate or extend real data sets with synthetic data sets that satisfy characteristics learned from a real reference data sets.
-\textcolor{red}{privacy, accuracy without details of underlying process, SMOTE - feature space} Oligos appears promising but will need to overcome several difficult challenges in automatically characterizing complex reference data to determine relevant constraints and learn accurate models that capture the effect of latent variables.
+Recent work is focused on generating synthetic data sets that satisfy characteristics learned from a real reference data sets \cite{Alexandrov2013}.
+Such approaches appear promising but will need to overcome several difficult challenges in automatically characterizing complex reference data to determine relevant constraints and learn accurate models that capture the effect of latent variables.
 
 Until such methods have reached maturity, the development of customized, domain-driven models is best suited for satisfying the needs of big data software and algorithms development.
-For example, the Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006} was developed for the Department of Homeland Security to develop synthetic data for effectiveness of IDASs designed to identify potential terrorism threat scenarios.
+Several customized, domain-driven models have seen usage in real-world situations.
+Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006} was developed for the Department of Homeland Security to develop synthetic data for effectiveness of IDASs designed to identify potential terrorism threat scenarios.
 

### 2014-07-19 bdcloud_paper/latex/paper.tex
Added discussion and conclusion

@@
 \section{Discussion and Conclusion}
-implementation
-[want a modular framework with clearly-defined extension points, modularity so that users do not need to know about the machinations of the entire model, and so that modules are easy to develop and turn on/off dynamically]
+We have described a domain-driven mathematical model and accompanying simulation for generating semantically-rich data.
+We validated the method by analyzing generated data to inform decisions in a fictional advertising campaign.
+Scalability testing combined with analysis the implementation suggests that the data generation method can scale from small data sets appropriate for desktop development to large data sets for testing and benchmarking of clusters.
 
-[spark, hadoop ecosystem, parallelization of generating stores, customers, and transactions]
+We see many opportunities for building on the work presented.
+We intend to expand the model to incorporate additional factors.
+Deeper integration of existing fields such as location and purchasing profiles (e.g., to model regional purchasing preferences) will increase the variety of the semantic information encoded in the generated data.
+The incorporation of weather and climate data can be used to influence when customers shop and the types of products they buy.
+For example, customers would be less likely to shop during snow storms and more likely to buy winter apparel for their pets.
+Modeling of time-dependent events such as sales, evolution of customer purchasing profiles over time, and ``life events,'' such as the birth or passing of pets, will enable more interesting time-series analysis of the data.
 
-[model extensions]
+The current implementation was prototyped in Python.
+We are working to implement the model using in Hadoop and Spark.
+We will commit the resulting implementations to the Apache BigTop project to enable ease of access and immediate benefit to current users.
 

### 2014-07-19 bdcloud_paper/latex/paper.tex
Added to discussion and conclusion
Added acknowledgements

@@
 Modeling of time-dependent events such as sales, evolution of customer purchasing profiles over time, and ``life events,'' such as the birth or passing of pets, will enable more interesting time-series analysis of the data.
+We would also like to expand the scope of the model to incorporate business processes (such as inventory management, customer complaints, employees, etc.), thus enabling queries about how internal business process affect customer behavior.
 
@@
 \section*{Acknowledgment}
-
-The authors would like to thank...
+The authors would like to thank Brian P.
+Clare and Casey Robinson for useful and interesting discussion.
+JV would like to thank the Apache BigTop community for their interest in and support of BigPetStore.
+We would also like to thank Matt Fenwick, Nigel Savage, and Bhashit Parikh for their contributions to BigPetStore.
+The authors would like to thank Red Hat, Inc.
+and the University of Notre Dame for their support of this work.
 

### 2014-07-19 bdcloud_paper/latex/paper.tex
Small edits

@@
 We validated the method by analyzing generated data to inform decisions in a fictional advertising campaign.
-Scalability testing combined with analysis the implementation suggests that the data generation method can scale from small data sets appropriate for desktop development to large data sets for testing and benchmarking of clusters.
+Scalability testing combined with analysis of the implementation suggests that the data generation method can scale from small data sets appropriate for desktop development to large data sets for testing and benchmarking of clusters.
+We have released the implementation source code under the Apache 2 open-source license.
 

### 2014-07-20 bdcloud_paper/latex/paper.tex
Highlighted abstract / keywords
Added keywords
edits to intro

@@
 %\boldmath
-Abstract
+\textcolor{red}{Abstract}
+
 \end{abstract}
@@
 % no keywords
+
+\begin{IEEEkeywords}
+big data, synthetic, data set, data generation, benchmarking, testing, probabilistic models
+\end{IEEEkeywords}
 
@@
 Targetted towards specific benchmarks such as Map-Reduce jobs and Hive queries, the resulting data is semantically-void and not useful for other tasks.
-Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005}.
-The more advanced approaches of these allow users to specify relationships between field values through simple arithmetic equations for setting the value of fields from other fields \cite{Alexandrov2012} or use constraint-solving techniques \cite{Arasu2011}.
-However, due to the inability to specify underlying dynamic models with latent variables, the generated data lacks informational complexity and genuine semantics, making it inapprioriate for analytics.
+Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}.
+Several approaches use domain-specific langugages \cite{Bruno2005,Hoag2007} to allow users to specify additional constraints such as which distributions to sample from or allow modeling basic relationships between fields such as simple arithmetic equations \cite{Alexandrov2012}.
+Arasu, et al.
+\cite{Arasu2011} demonstrated how constraint-solving techniques could be used as an alternative to procedural approaches.
+These frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for analytics.
 
-Recent work is focused on generating synthetic data sets that satisfy characteristics learned from a real reference data sets \cite{Alexandrov2013}.
-Such approaches appear promising but will need to overcome several difficult challenges in automatically characterizing complex reference data to determine relevant constraints and learn accurate models that capture the effect of latent variables.
+Realizing the difficulty in creating a generic framework capable of modeling the semantics of deep domain knowledge, recent work\cite{Alexandrov2013} has focused on generating synthetic data sets that satisfy characteristics learned from a real reference data sets.
+Such approaches appear promising but will need to overcome the difficulties of accurately training models, especially on raw data instead of features.
 

### 2014-07-20 bdcloud_paper/latex/paper.tex
Added references

@@
 Several synethic data set generators exist.
-TeraGen \textcolor{red}{cite} and the Intel Hadoop Benchmark Suite \cite{Huang2010} can generate data sets quickly and at any scale.
+TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} can generate data sets quickly and at any scale.
 Targetted towards specific benchmarks such as Map-Reduce jobs and Hive queries, the resulting data is semantically-void and not useful for other tasks.
@@
 Compared with BigPetStore's previous data generator, our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter.
-Combining the features of the TeraGen (scalability) \textcolor{red}{reference} and MovieLens (semantically rich) \textcolor{red}{reference} input data sets commonly used for big data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.
+Combining the features of the TeraGen (scalability)  and MovieLens \cite{MovieLens} (semantically rich) input data sets commonly used for big data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.
 To validate this approach, we perform an example analysis designed to guide a fictional advertising campaign (a typical business use case) to demonstrate that the model generates data with desirable properties (global patterns and region specific transactions).

### 2014-07-20 bdcloud_paper/latex/paper.tex
Added references

@@
 Furthermore, access to large data sets can be unreliable due to dependence on third party benefactors for hosting and requires significant amounts of bandwidth to download.
-Several companies (such as Twitter \textcolor{red}{reference}) offer access to their data but the costs and usage restrictions can be undesirable.
+Several companies (such as Twitter\cite{Twitter}) offer access to their data but the costs and usage restrictions can be undesirable.
 
@@
 The code for generating the data is often small, making it easy and cheap to host, transfer, and store.
-Licensing restrictions and data privacy issues are avoided \textcolor{red}{cite}.
+Licensing restrictions and data privacy issues are avoided with synthetic data sets.
 And, the size of the data sets can be scaled arbitrarily, enabling seamless usage of the same data set across a range of situations.

### 2014-07-20 bdcloud_paper/latex/paper.tex
Added references

@@
 Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.
-New tools such as Hive, Pig and Crunch \textcolor{red}{reference} based around big data frameworks such as Hadoop, Storm, and Spark \textcolor{red}{reference} are being released every day.
+New tools such as Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Shark \cite{Engle2012} based around big data frameworks such as Hadoop \cite{Hadoop}, Storm \cite{Storm}, and Spark \cite{Zaharia2010,Zaharia2012} are being released every day.
 Development, testing and benchmarking of big data software, algorithms, and infrastructure pose unique challenges.

### 2014-07-20 bdcloud_paper/latex/paper.tex
Added references

@@
 
-The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout from across the Hadoop and Spark ecosystems \textcolor{red}{reference}.
+The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout \cite{Mahout} from across the Hadoop and Spark ecosystems.
 BigPetStore's usefulness has been demonstrated by its incorporation into the open-source Apache BigTop distribution \cite{BigTop} for testing and as a reference application.

### 2014-07-21 bdcloud_paper/latex/paper.tex
Added equations for MSM
Added intro paragraph to evaluation

@@
 
+\textcolor{red}{description. min, max}
+
+\begin{align}
+Pr(X_{n+1}&=x|X_n=y) = \\
+& \left\{ 
+  \begin{array}{l l}
+   (1.0 - w_l) W^{-1} \sum_f w_f w_{f,s}(x, y)  & \quad \text{if $x \neq y$}\\
+   w_l & \quad \text{if $x = y$}
+  \end{array} \right.
+\end{align}
+
 \begin{equation}
-Pr(X_{n+1}=x|X_n) = \dots
+w_{f,s}(x,y) = \left\{ 
+  \begin{array}{l l}
+   w_{f,s}  & \quad \text{if $x$.$f$ = $y$.$f$}\\
+   1 - w_{f, s} & \quad \text{if $x$.$f$ $\neq$ $y$.$f$}
+  \end{array} \right.
+\end{equation}
+
+\begin{align}
+&w_l \sim N(\mu, \sigma^2) \\
+&w_f \sim N(\mu, \sigma^2) \\
+&w_{f, s} \sim N(\mu, \sigma^2)
+\end{align}
+
+\begin{equation}
+W = \sum_i \sum_{j \neq i} \sum_f w_f w_{f,s}(i, j)
 \end{equation}
@@
 \section{Evaluation of the Model}
-
-\textcolor{red}{intro paragraph}
+In accordance with our original design goals, we evaluated the model and implementation in terms of their abilities to generate semantically-interesting data and scale from a single local machine to a cluster.
 
 \subsection{Example Analysis of Generated Data}
-Using data for 10 stores, 10,000 customers, and five years of transactions generated from the model, we performed two example analyses in support of a fictional advertising campaign.
+To evaluate the semantic content of the generated data, we performed two example analyses in support of a fictional advertising campaign using data for 10 stores, 10,000 customers, and five years of transactions generated from the model.
 The analyses were designed to be realistic and driven by real-world business concerns.

### 2014-07-21 bdcloud_paper/latex/paper.tex
Edited intro to scaling

@@
 BigPetStore aims to scale from a local desktop to a large cluster.
-As a result, we benchmarked the generation of data on a laptop with a 2 GHz Intel Core i7 CPU, 8 GB of RAM, and a 256 GB SSD using the Python implementation with a single thread.
+To evaluate the scaling of the model and implementation, we benchmarked the data generator on a laptop with a 2 GHz Intel Core i7 CPU, 8 GB of RAM, and a 256 GB SSD using the Python implementation with a single thread.
 Using the test setup, between 1,500 and 2,000 transactions can be generated per second (Table~\ref{tab:benchmarks}).

### 2014-07-21 bdcloud_paper/latex/paper.tex
Added intro paragraph to model

@@
 \section{Overview of Model \& Simulation Procedures}
+Our generatorive model combines various well-known mathematical modeling techniques to simulate the factors affecting customers' purchasing habits.
+Each part of the model is derived from \emph{ab initio} assumptions.
+In several cases, real data is used to parameterize parts of the model.
 
-\textcolor{red}{need an intro paragrpah}
+\textcolor{red}{relational data model diagram?}
 

### 2014-07-21 bdcloud_paper/latex/paper.tex
Use IEEE citation format
updated keywords
Added abstract

@@
 %
-%\usepackage{cite}
+\usepackage{cite}
 % cite.sty was written by Donald Arseneau
@@
 \begin{abstract}
-%\boldmath
-\textcolor{red}{Abstract}
-
+Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key for consistent development, testing, and benchmarking of big data software, algorithms, and infrastructure.
+BigPetStore is a big data application blueprint, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics, built around a fictional chain of pet store which has been used for testing and benchmarking in the Apache BigTop distribution.
+We present a new generative data model for BigPetStore and demonstrate the model's ability to generate semantically-rich data that can be used for complex analytics workflows at variable scale ranging from a single machine to a large cluster.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-big data, synthetic, data set, data generation, benchmarking, testing, probabilistic models
+big data, synthetic data sets, data generation, benchmarking, testing, probabilistic models
 \end{IEEEkeywords}
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliographystyle{plain}
+\bibliographystyle{IEEEtran}
 % argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,paper}
-\bibliography{paper}
+\bibliography{IEEEabrv,paper}
+%\bibliography{paper}
 %

### 2014-07-21 bdcloud_paper/latex/paper.tex
Small edits

@@
 Over time, more high-quality data sets are becoming publicly available, but these data sets often carry burdens that force developers to make undesirable trade offs.
-Since few data sets are scalable, allowing the user to pick how much data they want, users are forced to choose between small data sets which are useful for development and testing on a local machine, but not for production-scale tests or benchmarking, or large data sets which are prohibitively expensive to host, transfer, and store.
+Since few data sets are scalable (allowing the user to pick how much data they want), users are forced to choose between small data sets which are useful for development and testing on a local machine, but not for production-scale tests or benchmarking, or large data sets which are prohibitively expensive to host, transfer, and store.
 Furthermore, access to large data sets can be unreliable due to dependence on third party benefactors for hosting and requires significant amounts of bandwidth to download.
@@
 
-Store records contain a unique identifier and a location in the form of a 5-digit zip code.
+Store records contain both a unique identifier and a location in the form of a 5-digit zip code.
 The locations of the stores are modeled by a probability density function (PDF) that gives the probability that a store's location is the given zip code.

### 2014-07-21 bdcloud_paper/latex/paper.tex
Spell checking:

@@
 
-Several synethic data set generators exist.
+Several synthetic data set generators exist.
 TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} can generate data sets quickly and at any scale.
-Targetted towards specific benchmarks such as Map-Reduce jobs and Hive queries, the resulting data is semantically-void and not useful for other tasks.
+Targeted towards specific benchmarks such as Map-Reduce jobs and Hive queries, the resulting data is semantically-void and not useful for other tasks.
 Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}.
-Several approaches use domain-specific langugages \cite{Bruno2005,Hoag2007} to allow users to specify additional constraints such as which distributions to sample from or allow modeling basic relationships between fields such as simple arithmetic equations \cite{Alexandrov2012}.
+Several approaches use domain-specific languages \cite{Bruno2005,Hoag2007} to allow users to specify additional constraints such as which distributions to sample from or allow modeling basic relationships between fields such as simple arithmetic equations \cite{Alexandrov2012}.
 Arasu, et al.
-\cite{Arasu2011} demonstrated how constraint-solving techniques could be used as an alternative to procedural approaches.
+\cite{Arasu2011} demonstrated that constraint-solving techniques could be used as an alternative to procedural approaches.
 These frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for analytics.
@@
 \section{Overview of Model \& Simulation Procedures}
-Our generatorive model combines various well-known mathematical modeling techniques to simulate the factors affecting customers' purchasing habits.
+Our generative model combines various well-known mathematical modeling techniques to simulate the factors affecting customers' purchasing habits.
 Each part of the model is derived from \emph{ab initio} assumptions.
@@
 
-For each proposed customer location zipcode $z$, we compute the distance between $z$ and every store's zipcode to find the closest store.
-We compute the distance $d$ between $z$ and the zipcode $s_z$ of the closest store.
-The weight $w_z$ for each zipcode $z$ is modeled using an exponential distribution with the average distance $\beta$.
+For each proposed customer location zip code $z$, we compute the distance between $z$ and every store's zip code to find the closest store.
+We compute the distance $d$ between $z$ and the zip code $s_z$ of the closest store.
+The weight $w_z$ for each zip code $z$ is modeled using an exponential distribution with the average distance $\beta$.
 
@@
 \subsection{Transactions}
-A transaction model is created for each customer and integrated over the simulation's timeframe to generate transactions.
+A transaction model is created for each customer and integrated over the simulation's time frame to generate transactions.
 Transaction records consist of a customer, a store, a date and time, and a list of items purchased and their prices.
@@
 The date and time is given as a floating-point number with units of days since the start of the simulation.
-(Timezones are not taken into account.)  Transaction times are generated from a Monte Carlo process based on the customer will use up items such as dog food (Section~\ref{sec:transaction-times}).
+(Time zones are not taken into account.)  Transaction times are generated from a Monte Carlo process based on the customer will use up items such as dog food (Section~\ref{sec:transaction-times}).
 The purchased items are generated by a Hidden Markov Model (HMM) parameterized by the transaction time and amount of time remaining for the items in the customer's inventory (Section~\ref{sec:transaction-purchases}).
@@
 The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$.
-The SDE is numerically integrated using the Euler-Maruyama method\cite{Klouden13} with timestep $\Delta t_n$ sampled from a Poisson process:
+The SDE is numerically integrated using the Euler-Maruyama method\cite{Klouden13} with time step $\Delta t_n$ sampled from a Poisson process:
 
@@
 where $\lambda_i$ is the item category's average usage rate, $\mu_i$ is the item category's average amount used per use, and $\sigma^2_i$ is the variance of the amount of the item category used per use.
-The parameter $\lambda_i$ is defined for each item cateogry.
+The parameter $\lambda_i$ is defined for each item category.
 The parameters $\mu_{i, c}$, and $\sigma^2_{i,c}$ are found based on a base rate for each item category multiplied by the number of pets of the appropriate species the customer has.
@@
 The proposed transaction time is found from the exhaustion times.
-For each item cateogry, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an exponential distribution parameterized separately for each customer $c$.
+For each item category, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an exponential distribution parameterized separately for each customer $c$.
 A proposed transaction time $T_{P, i}$ for each item category is found by subtracting the offset time $T_{O, i}$ from the exhaustion time $T_{E, i}$.
-The earlist proposed transaction time is taken as the overall proposed transaction time $T_T$.
+The earliest proposed transaction time is taken as the overall proposed transaction time $T_T$.
 
@@
 
-After an item is chosen, the item category's exhaustion time is updated by propogating the item category usage SDE described in Section~\ref{sec:transaction-times}.
+After an item is chosen, the item category's exhaustion time is updated by propagating the item category usage SDE described in Section~\ref{sec:transaction-times}.
 \textcolor{red}{state transition, stop state}
@@
 
-Secondly, we profiled our customers' purchasing habits to optimize the advertising campaign's effectivness by customizing the advertized products for each customer.
+Secondly, we profiled our customers' purchasing habits to optimize the advertising campaign's effectiveness by customizing the advertised products for each customer.
 For each customer, we generated a feature vector by computing the frequency with which they would purchase the same brand or flavor from one transaction to the next.
-We clustered the customers' feature vectors using the KMean algorithm with a range of cluster counts.
+We clustered the customers' feature vectors using the KMeans algorithm with a range of cluster counts.
 Twenty clusters converged the error.
@@
 High frequencies of purchasing the same flavors repeatedly were tightly-correlated with high frequencies of purchasing the same brands -- these customers were likely to be happy with a particular item and kept purchasing the same item repeatedly.
-For our advertising campaign, it would be unlikely to get these customers to purchase different items, so we should create incentives to purchase larger quantites, especially when inventory levels are high and needed to be depleted.
+For our advertising campaign, it would be unlikely to get these customers to purchase different items, so we should create incentives to purchase larger quantities, especially when inventory levels are high and needed to be depleted.
 Other customers had a tendency to purchase either the same flavor or brand repeatedly, but varied in their choice of the other.
@@
 Likewise, for customers who prefer a particular flavor, we should target our advertising campaign to suggest other brands with that flavor.
-Lastly, some customers had very low frequencies of purchasing neither the same brand nr flavor.
+Lastly, some customers had very low frequencies of purchasing neither the same brand nor flavor.
 Other factors, such as cost, not represented in this analysis may be driving these customers' purchasing habits and will need further study.

### 2014-07-22 bdcloud_paper/latex/paper.tex
Added relational data model

@@
 
-\textcolor{red}{relational data model diagram?}
+\subsection{Relational Data Model}
+The model generates data for four types of records: stores, customers, transactions, and transaction items (Figure~\ref{fig:relational-data-model}).
+Store records contain both a unique identifier and a location in the form of a 5-digit zip code.
+Customer records consist of a unique identifier, a name, and a location given as a 5-digit zip code.
+Transaction records consist of a customer identifier, a store identifier, and a time.
+Transaction item records contain a transaction identifier and an item description.
+The item description is a list of key-value pairs stored as a string.
+Key-value pairs are used since item characteristics differ depending on the type of item.
+For example, dog food has a brand, a flavor, a size (in lbs), and a per-unit cost, while poop bags have a brand, a color, a count, and a per-unit cost.
 
-\subsection{Stores}
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=3.5in]{figures/transactions_data_model.eps}
+  \caption{Relational Data Model for Generated Data}
+  \label{fig:relational-data-model}
+\end{figure}
+
+\subsection{Generation of Stores}
 
@@
 
-\subsection{Customers}
-
+\subsection{Generation of Customers}
 Customer records consist of a unique identifier, a name, a location given as a 5-digit zip code, and the number and species of their pets.
@@
 
-\subsection{Transactions}
+\subsection{Simulation of Transactions}
 A transaction model is created for each customer and integrated over the simulation's time frame to generate transactions.
@@
 
-\subsubsection{Transaction Times} \label{sec:transaction-times}
+\subsubsection{Simulation of Transaction Times} \label{sec:transaction-times}
 

### 2014-07-22 bdcloud_paper/latex/paper.tex
Edits of simulation text

@@
 Customer records consist of a unique identifier, a name, and a location given as a 5-digit zip code.
-Transaction records consist of a customer identifier, a store identifier, and a time.
+Transaction records consist of a customer identifier, a store identifier, and a time given in days since the beginning of the simulation.
 Transaction item records contain a transaction identifier and an item description.
@@
 \subsection{Generation of Stores}
-
-Store records contain both a unique identifier and a location in the form of a 5-digit zip code.
 The locations of the stores are modeled by a probability density function (PDF) that gives the probability that a store's location is the given zip code.
@@
 
-\begin{equation}
+\begin{equation*}
 p(\text{location}=z | \text{population}(z)) = \frac{\text{population}(z)}{\sum_{i} \text{population}(i)}
-\end{equation}
+\end{equation*}
 
@@
 
-\begin{equation}
+\begin{equation*}
 p(\text{location}=z | \text{income}(z)) = s ^ {\big( \frac{\text{income} - \min_i{\textrm{income}(i)}} {\max_i{\textrm{income}(i)} - \min_i{\textrm{income}(i)}} - 1 \big)}
-\end{equation}
+\end{equation*}
 
@@
 
-\begin{align}
-p(\text{location}=z) = &Z^-1 p(\text{location}=z | \text{population}(z)) \nonumber \\
+\begin{align*}
+p(\text{location}=z) = &Z^-1 p(\text{location}=z | \text{population}(z)) \\
 &p(\text{location}=z | \text{income}(z))
-\end{align}
+\end{align*}
 
-where $Z$ is the normalization factor.  Given the small size ($\approx$30,000 zip codes) of the data set, $Z$ can be found directly by iterating over all of the zip codes and summing the scores.
-
-The PDFs are parameterized using external data sets.
-The zip codes' latitudes and longitudes came from the Zip Code Database Project \cite{Zips}.
+where $Z$ is the normalization factor.
+Given the small size ($\approx$30,000 zip codes) of the data set, $Z$ can be found directly by iterating over all of the zip codes and summing the scores.
 The population and household income data for the zip codes were taken from from the U.S.
 Census American Community Survey \cite{ACS}.
-
 The stores' locations are generated by sampling zip codes from the PDF using a Roulette Wheel sampler.
@@
 \subsection{Generation of Customers}
-Customer records consist of a unique identifier, a name, a location given as a 5-digit zip code, and the number and species of their pets.
-
-\textcolor{red}{names}
-
 For each proposed customer location zip code $z$, we compute the distance between $z$ and every store's zip code to find the closest store.
 We compute the distance $d$ between $z$ and the zip code $s_z$ of the closest store.
+The zip codes' latitudes and longitudes came from the Zip Code Database Project \cite{Zips}.
 The weight $w_z$ for each zip code $z$ is modeled using an exponential distribution with the average distance $\beta$.
 
-\begin{align}
+\begin{align*}
 &w_z = \beta^{-1} \exp(-\beta^{-1} \, d_m(z)) \\
-&d_m(z) = \min_{s_z} \, d(z, s_z) \nonumber
-\end{align}
+&d_m(z) = \min_{s_z} \, d(z, s_z)
+\end{align*}
 
@@
 
-\begin{align}
+\begin{align*}
 &p(\text{location} = \text{z}) = W^{-1} w_z \\
 &W = \sum_{z} w(z) \nonumber
-\end{align}
+\end{align*}
 
 The customer locations are sampled from the PDF using a Roulette Wheel sampler.
+
+\textcolor{red}{names}
 
@@
 
-\begin{align}
+\begin{align*}
 N_p \sim U(1, b) \\
-N_d \sim U(0, N_p) \nonumber \\
-N_c = N_p - N_d \nonumber
-\end{align}
+N_d \sim U(0, N_p) \\
+N_c = N_p - N_d
+\end{align*}
 
 \subsection{Simulation of Transactions}
-A transaction model is created for each customer and integrated over the simulation's time frame to generate transactions.
-Transaction records consist of a customer, a store, a date and time, and a list of items purchased and their prices.
-The store is set to the store located closest to the customer.
-The date and time is given as a floating-point number with units of days since the start of the simulation.
-(Time zones are not taken into account.)  Transaction times are generated from a Monte Carlo process based on the customer will use up items such as dog food (Section~\ref{sec:transaction-times}).
+A transaction simulation is run for each customer.
+The transaction's store is set to the store located closest to the customer.
+Transaction times are generated from a Monte Carlo process that proposes transaction times based on the usage of the customer's items and Poisson processes modeling the amount of time between the customer visiting the store and running out of the items (Section~\ref{sec:transaction-times}).
 The purchased items are generated by a Hidden Markov Model (HMM) parameterized by the transaction time and amount of time remaining for the items in the customer's inventory (Section~\ref{sec:transaction-purchases}).
@@
 \subsubsection{Simulation of Transaction Times} \label{sec:transaction-times}
-
 Transaction times are simulated using a Monte Carlo method (Figure~\ref{fig:trans_sim}).
-The usage over time of each category is simulated, from which transaction times are found for each item category.
-The earliest transaction time is taken as the proposed transaction time.
+The usage over time of each item category is simulated.
+The time between the customer's visit to the store to buy more items in each category and the exhaustion time of that item category is modeled using a Poisson process.
+Proposed transaction times for each item category are computed based on the exhaustion time and time sampled from the Poisson process.
+The earliest transaction time is taken as the overall proposed transaction time.
 The probability of the transaction time is calculated using a PDF.
@@
 
-\begin{equation}
+\begin{equation*}
 \frac{da_i}{dt} = -\mu_i + \sigma_i\, dW_i(t)
-\end{equation}
+\end{equation*}
 
@@
 The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$.
-The SDE is numerically integrated using the Euler-Maruyama method\cite{Klouden13} with time step $\Delta t_n$ sampled from a Poisson process:
+The SDE is numerically integrated using the Euler-Maruyama method\cite{Klouden13} with time step $\Delta t_n$ sampled from an exponential distribution:
 
-\begin{align}
+\begin{align*}
 &a_{n+1, i} = a_{n,i} - \mu_{i, c} \, \Delta t_n + \sigma_{i, c} \, \sqrt{\Delta t_n} R_n \\
-&\Delta t_n \sim \text{Exp}(\lambda_i) \nonumber \\
-&R_n \sim \text{N}(0, 1) \nonumber
-\end{align}
+&\Delta t_n \sim \text{Exp}(\lambda_i) \\
+&R_n \sim \text{N}(0, 1)
+\end{align*}
 
-where $\lambda_i$ is the item category's average usage rate, $\mu_i$ is the item category's average amount used per use, and $\sigma^2_i$ is the variance of the amount of the item category used per use.
+where $\lambda_i$ is the item category's average usage rate, $\mu_i$ is the item category's average amount used per time used, and $\sigma^2_i$ is the variance of the amount of the item category used per time used.
 The parameter $\lambda_i$ is defined for each item category.
@@
 
-The proposed transaction time is found from the exhaustion times.
+The proposed transaction time is found from the exhaustion times (Eq.~\ref{eq:transaction-times}).
 For each item category, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an exponential distribution parameterized separately for each customer $c$.
@@
 
-\begin{align}
-&\beta_c \sim \text{U}(a, b) \\
-&T_{O, i} \sim \text{Exp}(\beta^{-1}_c) \\
-&T_{P, i} = T_{E,i} - T_{O, i} \\
+\begin{align} \label{eq:transaction-times}
+&\beta_c \sim \text{U}(a, b) \nonumber \\
+&T_{O, i} \sim \text{Exp}(\beta^{-1}_c) \nonumber \\
+&T_{P, i} = T_{E,i} - T_{O, i} \nonumber\\
 &T_T = \min_i \, \{  T_{P, i}\}
@@
 
-The acceptance probability $p(t=T_T)$ of the proposed transaction time $T_T$ is modeled using a PDF that incorporates factors such as store hours.
-The PDF could be extended to incorporate additional factors including time-dependent events such as sales, weather, and the amount of money the customer has available at the time.
+The acceptance probability $p(t_{n+1}=T_T|t_n)$ of the proposed transaction time $T_T$ is modeled using a PDF (Eq.~\ref{eq:transaction_time_pdf}).
+For now, the PDF only ensures that the proposed transaction time is more recent than the last transaction time.
 
-\begin{equation}
-p(t=T_T) = p(t=T_T | \text{store hours})
+%The PDF could be extended to incorporate additional factors including time-dependent events such as sales, weather, and the amount of money the customer has available at the time.
+
+\begin{equation} \label{eq:transaction_time_pdf}
+p(t_{n+1}=T_T|t_n) = \left\{ 
+  \begin{array}{l l}
+   1 & \quad t_{n+1} \geq t_n\\
+   0 & \quad t_{n+1} < t_n
+  \end{array} \right.
 \end{equation}
 
-The proposed transaction time is accepted if $p(t=T_T) < r$, where $r \sim \text{U}(0, 1)$.
+The proposed transaction time is accepted if $p(t_{n+1}=T_T) < r$, where $r \sim \text{U}(0, 1)$.
 Until the proposed transaction time is accepted, a new proposed transaction time is generated by sampling new offset times.
@@
 
-\subsubsection{Shopping Cart Items} \label{sec:transaction-purchases}
+\subsubsection{Simulation of Transaction Items} \label{sec:transaction-purchases}
 
 A customer's purchases in each transaction are modeled using a layered Hidden Markov Model (HMM) (Figure~\ref{fig:shopping_cart_sim}).
-The HMM has three states: the start state, the purchase states, and the stop state.
+The HMM has three types of states: the start state, the purchase states, and the stop state.
 There are an infinite number of purchase states, identified by the exhaustion times of the customer's item categories, the transaction time, and the number of items purchased in the current transaction.
@@
 
-\begin{equation}
+\begin{equation*}
 w_i = - \lambda_i \exp(-\lambda_i (T_{E, i}- T_{T}))
-\end{equation}
+\end{equation*}
 
@@
 
-\begin{align}
+\begin{align*}
 &p(X=i) = W^{-1} w_i \\
 &W = \sum_i w_i \nonumber
-\end{align}
+\end{align*}
 
@@
 
-\begin{align}
+\begin{align*}
 Pr(X_{n+1}&=x|X_n=y) = \\
@@
   \end{array} \right.
-\end{align}
+\end{align*}
 
-\begin{equation}
+\begin{equation*}
 w_{f,s}(x,y) = \left\{ 
@@
   \end{array} \right.
-\end{equation}
+\end{equation*}
 
-\begin{align}
+\begin{align*}
 &w_l \sim N(\mu, \sigma^2) \\
@@
 &w_{f, s} \sim N(\mu, \sigma^2)
-\end{align}
+\end{align*}
 
-\begin{equation}
+\begin{equation*}
 W = \sum_i \sum_{j \neq i} \sum_f w_f w_{f,s}(i, j)
-\end{equation}
+\end{equation*}
 

### 2014-07-22 bdcloud_paper/latex/paper.tex
Edits to MSM transition probabilities

@@
 Markov Models are used to model the customer's purchasing behavior for each item category.
-Each state corresponds to an item in that category.
-The transition probabilities are assigned according to the customer's buying habits.
-For example, if a customer tends to buy the same brand of dog food repeatedly, the outgoing edges to all products of the same brand as the product in the current state will be higher than those for other brands' products.
-If the customer does not care about flavor, the outgoing edges to all products from the current brand will be equal, otherwise the edge weights can be adjusted to prefer a specific combination of brand and flavor.
-Likewise, if a customer cares more about flavor or cost, the edge weights will be assigned appropriately.
-To simulate an item purchase, the Markov model is transitioned forward by one state.
-The current states of the Markov models are kept between purchases.
+Each state $X_n$ corresponds to an item in that category.
+The customer's buying habits are determined by the transition probabilities (Eq.~\ref{eq:transition_probabilities}).
+The transition probabilities depend on three types of parameters: loopback weight $w_l$ indicating the probability of purchasing the same item again, weight $w_f$ of a particular field $f$, and the weight $w_{f,s}$ when two items $x$ and $y$ have the same value for the field $f$.
 
-\textcolor{red}{description. min, max}
-
-\begin{align*}
-Pr(X_{n+1}&=x|X_n=y) = \\
+\begin{align} \label{eq:transition_probabilities}
+Pr(X_{n+1}&=x|X_n) = \nonumber \\
 & \left\{ 
   \begin{array}{l l}
-   (1.0 - w_l) W^{-1} \sum_f w_f w_{f,s}(x, y)  & \quad \text{if $x \neq y$}\\
-   w_l & \quad \text{if $x = y$}
+   (1.0 - w_l) W^{-1} \sum_f w_f w_{f,s}(x, y)  & \quad \text{if $x \neq X_n$}\\
+   w_l & \quad \text{if $x = X_n$}
   \end{array} \right.
-\end{align*}
+\end{align}
+
+\textcolor{red}{todo}
 
@@
 
-\begin{align*}
-&w_l \sim N(\mu, \sigma^2) \\
-&w_f \sim N(\mu, \sigma^2) \\
-&w_{f, s} \sim N(\mu, \sigma^2)
-\end{align*}
+\textcolor{red}{todo}
 
@@
 \end{equation*}
+
+The weights $w_l$, $w_f$, and $w_{f, s}$ are chosen randomly for each customer.
+
+\textcolor{red}{description. min, max}
+
+\begin{align} \label{eq:item-weights}
+&w_l \sim N(\mu, \sigma^2) \nonumber \\
+&w_f \sim N(\mu, \sigma^2) \nonumber \\
+&w_{f, s} \sim N(\mu, \sigma^2)
+\end{align}
+
+\textcolor{red}{reorder so that probilities describe habits}
+
+For example, if a customer tends to buy the same brand of dog food repeatedly, the outgoing edges to all products of the same brand as the product in the current state will be higher than those for other brands' products.
+If the customer does not care about flavor, the outgoing edges to all products from the current brand will be equal, otherwise the edge weights can be adjusted to prefer a specific combination of brand and flavor.
+Likewise, if a customer cares more about flavor or cost, the edge weights will be assigned appropriately.
+
+To simulate an item purchase, the Markov model is transitioned forward by one state.  The current states of the Markov models are kept between purchases.
 
@@
 The current implementation was prototyped in Python.
-We are working to implement the model using in Hadoop and Spark.
+We are working to implement the model using Hadoop and Spark.
 We will commit the resulting implementations to the Apache BigTop project to enable ease of access and immediate benefit to current users.

### 2014-07-22 src/python/customers.py
Updated output to match diagram

@@ -12,6 +12,7 @@ import simulation_parameters as sim_param
 
 class Customer(object):
     def __init__(self):
+        self.id = None
         self.name = None
         self.location = None
         self.average_transaction_trigger_time = None
@@ -22,8 +23,8 @@ class Customer(object):
                     }
 
     def __repr__(self):
-        return "(%s, %s dogs, %s cats, %s)" % \
-            (self.name, self.pets["dog"], 
+        return "(%s, %s, %s dogs, %s cats, %s)" % \
+            (self.id, self.name, self.pets["dog"], 
              self.pets["cat"], self.location)
 
 class LocationSampler(object):
@@ -80,12 +81,15 @@ class CustomerGenerator(object):
         self.location_sampler = LocationSampler(stores=stores,
                                                 zipcode_objs=zipcode_objs,
                                                 avg_distance=sim_param.AVERAGE_CUSTOMER_STORE_DISTANCE)
+        self.current_id = 0
 
 
     def generate(self, n):
         customers = list()
         for i in xrange(n):
             customer = Customer()
+            customer.id = self.current_id
+            self.current_id += 1
             customer.name = "Customer_" + str(i)
             customer.location = self.location_sampler.sample()
             

### 2014-07-22 src/python/simulation.py

@@ -29,40 +29,13 @@ class Simulator(object):
         for customer in self.customers:
             state = CustomerState(item_categories=self.item_categories,
                     customer=customer)
-            trans_sim = TransactionSimulator(customer_state=state,
+            trans_sim = TransactionSimulator(stores=self.stores,
+                                             customer_state=state,
                                              item_categories=self.item_categories)
             for trans in trans_sim.simulate(end_time):
                 yield trans
 
-
-
-class CustomerWriter(object):
-    def __init__(self, filename=None):
-        self.fl = open(filename, "w")
-
-    def append(self, customer):
-        string = "%s,%s,%s,%s\n" % (customer.name, 
-                                    customer.location,
-                                    customer.pets["dog"],
-                                    customer.pets["cat"])
-        self.fl.write(string)
-
-    def close(self):
-        self.fl.close()
-
-class StoreWriter(object):
-    def __init__(self, filename=None):
-        self.fl = open(filename, "w")
-
-    def append(self, store):
-        string = "%s,%s\n" % (store.name, store.zipcode)
-        self.fl.write(string)
-
-    def close(self):
-        self.fl.close()
-
-
-class TransactionWriter(object):
+class TransactionItemWriter(object):
     def __init__(self, filename=None):
         self.fl = open(filename, "w")
 
@@ -80,13 +53,49 @@ class TransactionWriter(object):
             else:
                 item_str = "%s:%s:%s" % \
                     (item["category"], item["brand"], item["size"])
-            
 
+            self.fl.write("%s,%s\n" % (trans.transaction_id(),
+                                  item_str))
+    def close(self):
+        self.fl.close()
+    
+
+class CustomerWriter(object):
+    def __init__(self, filename=None):
+        self.fl = open(filename, "w")
+
+    def append(self, customer):
+        string = "%s,%s,%s\n" % (customer.id,
+                                 customer.name, 
+                                 customer.location)
+                     
+        self.fl.write(string)
+
+    def close(self):
+        self.fl.close()
+
+class StoreWriter(object):
+    def __init__(self, filename=None):
+        self.fl = open(filename, "w")
+
+    def append(self, store):
+        string = "%s,%s\n" % (store.id, store.zipcode)
+        self.fl.write(string)
+
+    def close(self):
+        self.fl.close()
+
+
+class TransactionWriter(object):
+    def __init__(self, filename=None):
[…]

### 2014-07-22 src/python/stores.py

@@ -48,6 +48,7 @@ class ZipcodeSampler(object):
 
 class Store(object):
     def __init__(self):
+        self.id = None
         self.name = None
         self.zipcode = None
         self.coords = None
@@ -61,11 +62,14 @@ class StoreGenerator(object):
         self.zipcode_objs = zipcode_objs
         self.zipcode_sampler = ZipcodeSampler(zipcode_objs=zipcode_objs,
                                                 income_scaling_factor=income_scaling_factor)
+        self.current_id = 0
         
     def generate(self, n):
         stores = list()
         for i in xrange(n):
             store = Store()
+            store.id = self.current_id
+            self.current_id += 1
             store.name = "Store_" + str(i)
             store.zipcode = self.zipcode_sampler.sample()
             store.coords = self.zipcode_objs[store.zipcode].coords

### 2014-07-22 src/python/transactions.py

@@ -3,6 +3,8 @@ from collections import defaultdict
 import math
 import random
 
+import hashlib
+
 from markovmodel import MarkovModelBuilder
 from samplers import RouletteWheelSampler
 
@@ -62,17 +64,22 @@ class ItemCategoryMarkovModelBuilder(object):
 
 
 class Transaction(object):
-    def __init__(self, customer=None, trans_time=None, purchased_items=None, inventory_before=None, inventory_after=None):
+    def __init__(self, customer=None, trans_time=None, purchased_items=None, store=None,
+                 trans_count=None):
+        self.store = store
         self.customer = customer
         self.trans_time = trans_time
         self.purchased_items = purchased_items
-        self.inventory_before = inventory_before
-        self.inventory_after = inventory_after
+        self.trans_count = trans_count
+
+    def transaction_id(self):
+        return hashlib.md5(repr(self)).hexdigest()
 
     def __repr__(self):
-        return "(%s, Time: %s, Purchased: %s\n\tInventory Before: %s\n\tInventory After: %s)" % (self.customer,
-            self.trans_time, len(self.purchased_items), self.inventory_before,
-                self.inventory_after)
+        return "(%s, %s, %s, %s)" % (self.store.id,
+                                     self.customer.id,
+                                     self.trans_time,
+                                     self.trans_count)
 
 
 class TransactionPurchasesSimulator(object):
@@ -162,10 +169,12 @@ class TransactionTimeSampler(object):
 
 
 class TransactionSimulator(object):
-    def __init__(self, customer_state=None, item_categories=None):
+    def __init__(self, stores=None, customer_state=None, item_categories=None):
+        self.stores = stores
         self.customer_state = customer_state
         self.trans_time_sampler = TransactionTimeSampler(customer_state=customer_state)
         self.purchase_sim = TransactionPurchasesSimulator(customer_state=self.customer_state, item_categories=item_categories) 
+        self.trans_count = 0
     
     def simulate(self, end_time):
         last_trans_time = 0.0
@@ -175,14 +184,13 @@ class TransactionSimulator(object):
             if trans_time > end_time:
                 break
             
-            remaining_before = self.customer_state.get_inventory_amounts(trans_time)
             purchased_items = self.purchase_sim.simulate(trans_time=trans_time)
-            remaining_after = self.customer_state.get_inventory_amounts(trans_time)
             trans = Transaction(customer=self.customer_state.customer,
-                        purchased_items=purchased_items,
[…]

### 2014-07-22 bdcloud_paper/latex/paper.tex
Edits to model description

@@
 \subsection{Generation of Customers}
-For each proposed customer location zip code $z$, we compute the distance between $z$ and every store's zip code to find the closest store.
-We compute the distance $d$ between $z$ and the zip code $s_z$ of the closest store.
-The zip codes' latitudes and longitudes came from the Zip Code Database Project \cite{Zips}.
-The weight $w_z$ for each zip code $z$ is modeled using an exponential distribution with the average distance $\beta$.
+For each proposed customer location zip code $z$, we compute the distance $d_m(z)$ between $z$ and each store's zip code $s_z$ to find the closest store.
+The zip codes' latitudes and longitudes (taken from the Zip Code Database Project \cite{Zips}) are used to compute the distances.
+Each zip code $z$ is assigned a weight $w_z$ according to its distance $d_m(z)$ to the nearest store using an exponential distribution with the average distance $\beta$:
 
@@
 
-To find the probabilities, the weights are normalized:
+The probability of choosing zip code $z$ is given by
 
-\begin{align*}
-&p(\text{location} = \text{z}) = W^{-1} w_z \\
-&W = \sum_{z} w(z) \nonumber
-\end{align*}
+\begin{equation*}
+p(\text{location} = \text{z}) = \frac{w_z}{\sum_{z} w(z)}
+\end{equation*}
+
+\textcolor{red}{names}
 
 The customer locations are sampled from the PDF using a Roulette Wheel sampler.
-
-\textcolor{red}{names}
 
@@
 The usage over time of each item category is simulated.
-The time between the customer's visit to the store to buy more items in each category and the exhaustion time of that item category is modeled using a Poisson process.
+Items categories are groups of items which are interchangeable such as ``dry dog food,'' ``dry cat food,'' and ``kitty litter.'' The time between the customer's visit to the store to buy more items in each category and the exhaustion time of that item category is modeled using a Poisson process.
 Proposed transaction times for each item category are computed based on the exhaustion time and time sampled from the Poisson process.
@@
 
-In the first stage of the process, the usage of items over time is modeled using a simple stochastic differential equation (SDE):
+In the first stage of the process, the usage of items over time is modeled using a stochastic differential equation (SDE):
 
@@
 
-The probability $p(X=i)$ of choosing item category $i$ is given by the normalizing the weight $w_i$ by the sum of all of the weights $W$:
+The probability $p(X=i)$ of choosing item category $i$ is given by
 
-\begin{align*}
-&p(X=i) = W^{-1} w_i \\
-&W = \sum_i w_i \nonumber
-\end{align*}
+\begin{equation*}
+p(X=i) = \frac{w_i}{\sum_i w_i}.
+\end{equation*}
 
@@
   \includegraphics[width=3.5in]{figures/cluster_analysis.eps}
-  \caption{Clustering of Brand and Flavor Purchasing Preferences}
+  \caption{Clustering of Brand and Flavor Purchasing Preferences. Grey plus signs (+) represent customer data points, and cyan circles represent cluster centers.}
   \label{fig:cluster_analysis}

### 2014-07-23 src/python/customers.py
Added generation of names

@@ -27,6 +27,35 @@ class Customer(object):
             (self.id, self.name, self.pets["dog"], 
              self.pets["cat"], self.location)
 
+class NameSampler(object):
+    def __init__(self, first_names, last_names):
+        normalized_first_names = self.normalize(first_names)
+        normalized_last_names = self.normalize(last_names)
+
+        self.first_name_sampler = RouletteWheelSampler(normalized_first_names)
+        self.last_name_sampler = RouletteWheelSampler(normalized_last_names)
+
+    def normalize(self, names):
+        normalized_names = []
+
+        weight_sum = 0.0
+        for name, weight in names:
+            weight_sum += weight
+
+        for name, weight in names:
+            normalized_names.append((name, weight / weight_sum))
+
+        return normalized_names
+
+    def sample(self):
+        names = []
+        names.append(self.first_name_sampler.sample())
+        names.append(self.last_name_sampler.sample())
+
+        return " ".join(names)
+        
+
+        
 class LocationSampler(object):
     def __init__(self, stores=None, zipcode_objs=None, avg_distance=None):
         lambd = 1.0 / avg_distance
@@ -77,10 +106,12 @@ class LocationSampler(object):
         return self.sampler.sample()
 
 class CustomerGenerator(object):
-    def __init__(self, zipcode_objs=None, stores=None):
+    def __init__(self, zipcode_objs=None, stores=None, first_names=None,
+                 last_names=None):
         self.location_sampler = LocationSampler(stores=stores,
                                                 zipcode_objs=zipcode_objs,
                                                 avg_distance=sim_param.AVERAGE_CUSTOMER_STORE_DISTANCE)
+        self.name_sampler = NameSampler(first_names, last_names)
         self.current_id = 0
 
 
@@ -90,7 +121,7 @@ class CustomerGenerator(object):
             customer = Customer()
             customer.id = self.current_id
             self.current_id += 1
-            customer.name = "Customer_" + str(i)
+            customer.name = self.name_sampler.sample()
             customer.location = self.location_sampler.sample()
             
             num_pets = random.randint(sim_param.MIN_PETS, sim_param.MAX_PETS)
@@ -114,6 +145,21 @@ class CustomerGenerator(object):
             customer.pets["cat"] = num_cats
             customers.append(customer)
         return customers
+
+def load_names():
+    name_fl = open("../../resources/namedb/data/data.dat")
+    first_names = []
+    last_names = []
[…]

### 2014-07-23 src/python/simulation.py

@@ -1,4 +1,5 @@
 from customers import CustomerGenerator
+from customers import load_names
 from customer_simulation import CustomerState
 import simulation_parameters as sim_param
 from stores import StoreGenerator
@@ -14,6 +15,7 @@ class Simulator(object):
     def load_data(self):
         self.item_categories = products.load_products_json()
         self.zipcode_objs = load_zipcode_data()
+        self.first_names, self.last_names = load_names()
 
     def generate_stores(self, num=None):
         generator = StoreGenerator(zipcode_objs=self.zipcode_objs,
@@ -22,7 +24,9 @@ class Simulator(object):
 
     def generate_customers(self, num=None):
         generator = CustomerGenerator(zipcode_objs=self.zipcode_objs,
-                                      stores=self.stores)
+                                      stores=self.stores,
+                                      first_names=self.first_names,
+                                      last_names=self.last_names)
         self.customers = generator.generate(num)
 
     def generate_transactions(self, end_time=None):

### 2014-07-23 src/python/simulation_parameters.py

@@ -16,3 +16,4 @@ MAX_PETS = 10
 MIN_PETS = 1
 
 STORE_INCOME_SCALING_FACTOR = 100.0
+

### 2014-07-23 bdcloud_paper/latex/paper.tex
Added namedb description

@@
 Census American Community Survey \cite{ACS}.
-The stores' locations are generated by sampling zip codes from the PDF using a Roulette Wheel sampler.
+The stores' locations are generated by sampling zip codes from the PDF.
 
@@
 
-The probability of choosing zip code $z$ is given by
+The customer's zip code is chosen by sampling from the zip codes with the probabiltiy of each zip code $z$ proportionate to its weight $w_z$.
 
-\begin{equation*}
-p(\text{location} = \text{z}) = \frac{w_z}{\sum_{z} w(z)}
-\end{equation*}
-
-\textcolor{red}{names}
-
-The customer locations are sampled from the PDF using a Roulette Wheel sampler.
+Names are generated using data from the Name Database\cite{NameDB}.
+Each record in the database gives a name, a weight, and flag indicatings if the name can be used as a first name, a last name, or both.
+PDFs generated for the first and last names using the weights.
+The customer's name is generated by sampling a first name and a last name from each PDF respectively.
 
@@
 
-The probability $p(X=i)$ of choosing item category $i$ is given by
-
-\begin{equation*}
-p(X=i) = \frac{w_i}{\sum_i w_i}.
-\end{equation*}
-
-The item category for the purchase is chosen using a Roulette Wheel sampler.
+The item category for each purchase is chosen by sampling from the item categories with the probability of each item category $i$ proportionate to its weight $w_i$.
 

### 2014-07-23 bdcloud_paper/latex/paper.tex
Updated description of SDE

@@
 \begin{align*}
-N_p \sim U(1, b) \\
+N_c = N_p - N_d \\
 N_d \sim U(0, N_p) \\
-N_c = N_p - N_d
+N_p \sim U(1, b)
 \end{align*}
@@
 \begin{equation*}
-\frac{da_i}{dt} = -\mu_i + \sigma_i\, dW_i(t)
+\frac{da_i}{dt} = \min \{-\mu_i + \sigma_i\, dW_i(t), 0\}
 \end{equation*}
@@
 where $a_i$ is the amount of item category $i$ ``stuff'' remaining at time $t$.
-The variable $u_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$.
+The variable $\mu_i$ gives the average usage rate and $\sigma^2_i$ gives the variance of the usage rate for item category $i$.
 The SDE is numerically integrated using the Euler-Maruyama method\cite{Klouden13} with time step $\Delta t_n$ sampled from an exponential distribution:
@@
 \begin{align*}
-&a_{n+1, i} = a_{n,i} - \mu_{i, c} \, \Delta t_n + \sigma_{i, c} \, \sqrt{\Delta t_n} R_n \\
-&\Delta t_n \sim \text{Exp}(\lambda_i) \\
+&a_{n+1, i} = a_{n,i} - \min \{\mu_{i, c} \, \Delta t_n + \sigma_{i, c} \, \sqrt{\Delta t_n} R_n, 0.0\} \\
+&t_{n+1} = t_n + \Delta t_n \\
+&\Delta t_n \sim \text{Exp}(\beta^{-1}_{i,c}) \\
 &R_n \sim \text{N}(0, 1)
@@
 
-where $\lambda_i$ is the item category's average usage rate, $\mu_i$ is the item category's average amount used per time used, and $\sigma^2_i$ is the variance of the amount of the item category used per time used.
-The parameter $\lambda_i$ is defined for each item category.
-The parameters $\mu_{i, c}$, and $\sigma^2_{i,c}$ are found based on a base rate for each item category multiplied by the number of pets of the appropriate species the customer has.
-The exhaustion time $T_{E,i}$ for each item category is found by simulating the usage until $a_n = 0$.
+where $a_{n, i}$ is the amount remaining of item category $i$ at time step $t_n$, $\mu_{i,c}$ is the average amount used per time used, $\sigma^2_{i,c}$ is the variance of the amount used per time used, and $\beta_{i,c}$ is the item category's average amount of time between uses for customer $c$.
+The parameters $\mu_{i, c}$, and $\sigma^2_{i,c}$ are computed from a base rate for each item category multiplied by the number of pets of the appropriate species the customer $c$ has.
+The exhaustion time $T_{E,i}$ for each item category is found by simulating the usage until $a_{n,i} \leq 0$.
 
@@
 \begin{align} \label{eq:transaction-times}
-&\beta_c \sim \text{U}(a, b) \nonumber \\
+&T_T = \min_i \, \{  T_{P, i}\} \\
+&T_{P, i} = T_{E,i} - T_{O, i} \nonumber\\
 &T_{O, i} \sim \text{Exp}(\beta^{-1}_c) \nonumber \\
-&T_{P, i} = T_{E,i} - T_{O, i} \nonumber\\
-&T_T = \min_i \, \{  T_{P, i}\}
+&\beta_c \sim \text{U}(a, b) \nonumber
 \end{align}
@@
 Until the proposed transaction time is accepted, a new proposed transaction time is generated by sampling new offset times.
-If the proposed transaction time is accepted, the shopping cart of items is generated through a separate simulation discussed in Section~\ref{sec:transaction-purchases} below.
+If the proposed transaction time is accepted, the purchased items are chosen through a separate simulation discussed in Section~\ref{sec:transaction-purchases} below.
 Transactions are generated until the accepted transaction time is later than the end time given as a simulation parameter.

### 2014-07-23 src/python/products.py
Fixed bug where daily_usage_rate for item categories were inverted.

@@ -17,7 +17,7 @@ class ItemCategory(object):
         self.items = items
         self.species = species
         self.trigger_transaction = trigger_transaction
-        self.daily_usage_rate = daily_usage_rate
+        self.daily_usage_rate = 1.0 / daily_usage_rate
         self.base_amount_used_average = base_amount_used_average
         self.base_amount_used_variance = base_amount_used_average
         self.transaction_trigger_rate = transaction_trigger_rate

### 2014-07-23 src/python/products.py
revert last change

@@ -17,7 +17,7 @@ class ItemCategory(object):
         self.items = items
         self.species = species
         self.trigger_transaction = trigger_transaction
-        self.daily_usage_rate = 1.0 / daily_usage_rate
+        self.daily_usage_rate = daily_usage_rate
         self.base_amount_used_average = base_amount_used_average
         self.base_amount_used_variance = base_amount_used_average
         self.transaction_trigger_rate = transaction_trigger_rate

### 2014-07-23 bdcloud_paper/latex/paper.tex
rewrote intro and abstract

@@
 \begin{abstract}
-Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key for consistent development, testing, and benchmarking of big data software, algorithms, and infrastructure.
+Both developers and users of big data systems benefit from realistic example applications.
+For developers, such as applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.
+
+We present a new domain-driven, generative data model for BigPetStore, a big data application blueprint for the Hadoop ecosystem and included in the Apache BigTop distribution.
+We describe the model and demonstrate the its ability to generate semantically-rich data at variable scale ranging from a single machine to a large cluster.
+We validate the model by using the generated data to answer questions about customer locations and purchasing habits for a fictional targeted advertising campaign, a common business use case.
+
+%Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key for consistent development, testing, and benchmarking of big data software, algorithms, and infrastructure.
 BigPetStore is a big data application blueprint, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics, built around a fictional chain of pet store which has been used for testing and benchmarking in the Apache BigTop distribution.
@@
 Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.
-New tools such as Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Shark \cite{Engle2012} based around big data frameworks such as Hadoop \cite{Hadoop}, Storm \cite{Storm}, and Spark \cite{Zaharia2010,Zaharia2012} are being released every day.
+With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from realistic example applications.
+For developers, such as applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.
+
+BigPetStore is a big data application blueprint built around processing transaction data for a fictional chain of pet stores.
+BigPetStore mainly targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}.
+BigPetStore has been incorporated with the open-source Apache BigTop distribution \cite{BigTop}, where it is used for testing and as a reference application.
+One of BigPetStore's key features is the ability to scale from a single machine to a large cluster, making it easy to develop projects on a local machine and transfer the application to a large cluster for production testing.
+
+To achieve the project's goals of providing high-quality, realistic examples, BigPetStore requires semantically-rich, complex data.
+At its core, BigPetStore relies on a generative data model for producing synthetic transaction data.
+Despite the growing number of real data sets now publically available, synthetic data generators have a number of advantages for BigPetStore over real data sets.
+The synthetic data generator can be packaged with BigPetStore, avoiding the cost, time, and infrastructure needed to host, transfer, and store large data sets.
+Synthetic data generators are scalable, allowing the user to choose how much data to generate -- a requirement for supporting BigPetStore's goal of running on both single machines and large clusters.
+Licensing and privacy issues are avoided with the use of synthetic data sets.
+And lastly, synthetic data generators are tunable by the user, allowing the user to generate data with specific criteria amenable to testing.
+For example, a user may generate transactions from a small number of purchasable items so that clustering results can be easily visualized.
+
+% New tools such as Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Shark \cite{Engle2012} based around big data frameworks such as Hadoop \cite{Hadoop}, Storm \cite{Storm}, and Spark \cite{Zaharia2010,Zaharia2012} are being released every day.
 Development, testing and benchmarking of big data software, algorithms, and infrastructure pose unique challenges.
@@
 
-Over time, more high-quality data sets are becoming publicly available, but these data sets often carry burdens that force developers to make undesirable trade offs.
+%Over time, more high-quality data sets are becoming publicly available, but these data sets often carry burdens that force developers to make undesirable trade offs.
 Since few data sets are scalable (allowing the user to pick how much data they want), users are forced to choose between small data sets which are useful for development and testing on a local machine, but not for production-scale tests or benchmarking, or large data sets which are prohibitively expensive to host, transfer, and store.
@@
 
-Methods for generating synthetic data offer multiple advantages over pre-generated data sets.
+%Methods for generating synthetic data offer multiple advantages over pre-generated data sets.
 The code for generating the data is often small, making it easy and cheap to host, transfer, and store.
@@
 
-Several synthetic data set generators exist.
-TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} can generate data sets quickly and at any scale.
-Targeted towards specific benchmarks such as Map-Reduce jobs and Hive queries, the resulting data is semantically-void and not useful for other tasks.
+A variety of approaches for generating synthetic data set exist.
+TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} are popular tools that can generate data sets quickly and at any scale, but the resulting data is semantically-void and not useful for much more than simple benchmarks.
 Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}.
-Several approaches use domain-specific languages \cite{Bruno2005,Hoag2007} to allow users to specify additional constraints such as which distributions to sample from or allow modeling basic relationships between fields such as simple arithmetic equations \cite{Alexandrov2012}.
+Several approaches even provide domain-specific languages \cite{Bruno2005,Hoag2007} for specifying additional constraints such as which distributions to sample from and allow for modeling basic relationships between fields using simple arithmetic equations \cite{Alexandrov2012}.
 Arasu, et al.
 \cite{Arasu2011} demonstrated that constraint-solving techniques could be used as an alternative to procedural approaches.
-These frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for analytics.
+These frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for BigPetStore's analytics examples.
 
-Realizing the difficulty in creating a generic framework capable of modeling the semantics of deep domain knowledge, recent work\cite{Alexandrov2013} has focused on generating synthetic data sets that satisfy characteristics learned from a real reference data sets.
+Realizing the difficulty in creating a generic framework capable of modeling the semantics of real data, recent work\cite{Alexandrov2013} has focused on generating synthetic data sets that satisfy characteristics learned from real reference data.
 Such approaches appear promising but will need to overcome the difficulties of accurately training models, especially on raw data instead of features.
 
-Until such methods have reached maturity, the development of customized, domain-driven models is best suited for satisfying the needs of big data software and algorithms development.
-Several customized, domain-driven models have seen usage in real-world situations.
-Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006} was developed for the Department of Homeland Security to develop synthetic data for effectiveness of IDASs designed to identify potential terrorism threat scenarios.
+Until such methods have reached maturity, BigPetStore's needs are best met by a customized, domain-driven model.
+Another example of such a model is The Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006}.
+IDSG was developed to to test the effectiveness of IDASs in identifying potential terrorism threat scenarios using synthetic data.
+Like BigPetStore, the developers of IDSG needed to avoid licensing restrictions and privacy issues associated with real data.
 
-\textcolor{red}{Huppler, Pavlo}.  \textcolor{red}{standard benchmarks}
+%\textcolor{red}{Huppler, Pavlo}.  \textcolor{red}{standard benchmarks}
 
-The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout \cite{Mahout} from across the Hadoop and Spark ecosystems.
+%The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout \cite{Mahout} from across the Hadoop and Spark ecosystems.
 BigPetStore's usefulness has been demonstrated by its incorporation into the open-source Apache BigTop distribution \cite{BigTop} for testing and as a reference application.
 
-In this work, we present a new domain-driven model and simulation for generating arbitrarily large, complex data for testing and benchmarking of algorithms, software, and infrastructure in the big data space.
+BigPetStore's current model has been used successfully to generate terabytes of synthetic data and is used regularly for testing in Apache BigTop, showcasing the value of the approach.
+The current model is limited, however, in its ability to generate data that is semantically rich, limiting progress on BigPetStore's goals of providing realistic analytics examples.
+
+In this work, we present a new domain-driven model and simulation method for BigPetStore.
 Compared with BigPetStore's previous data generator, our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter.
 Combining the features of the TeraGen (scalability)  and MovieLens \cite{MovieLens} (semantically rich) input data sets commonly used for big data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.
-To validate this approach, we perform an example analysis designed to guide a fictional advertising campaign (a typical business use case) to demonstrate that the model generates data with desirable properties (global patterns and region specific transactions).
-We have made the model's implementation available as open source software.
+To demonstrate that the model generates data with desirable properties, we performed an example analysis designed to guide a fictional advertising campaign (a typical business use case).
+The model's implementation available as open source software.
 
@@
 The earliest proposed transaction time is taken as the overall proposed transaction time $T_T$.
+\textcolor{red}{$\beta_c$}
 
@@
 The item category for each purchase is chosen by sampling from the item categories with the probability of each item category $i$ proportionate to its weight $w_i$.
+\textcolor{red}{$\lambda_i$ - maybe these should be per customer / item category instead of just per item category}
 

### 2014-07-23 bdcloud_paper/latex/paper.tex
small edits

@@
 Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.
-With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from realistic example applications.
+With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from the availability of realistic example applications.
 For developers, such as applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.
@@
 BigPetStore mainly targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}.
-BigPetStore has been incorporated with the open-source Apache BigTop distribution \cite{BigTop}, where it is used for testing and as a reference application.
+BigPetStore has been incorporated into the open-source Apache BigTop distribution \cite{BigTop}, where it is used for testing and as a reference application.
 One of BigPetStore's key features is the ability to scale from a single machine to a large cluster, making it easy to develop projects on a local machine and transfer the application to a large cluster for production testing.
@@
 Licensing and privacy issues are avoided with the use of synthetic data sets.
-And lastly, synthetic data generators are tunable by the user, allowing the user to generate data with specific criteria amenable to testing.
+And lastly, the generated data can be customized by the user, allowing the user to generate data with specific criteria amenable to testing.
 For example, a user may generate transactions from a small number of purchasable items so that clustering results can be easily visualized.
@@
 
-\subsection{Relational Data Model}
+\subsection{Description of Generated Data}
 The model generates data for four types of records: stores, customers, transactions, and transaction items (Figure~\ref{fig:relational-data-model}).
@@
 
-The customer's zip code is chosen by sampling from the zip codes with the probabiltiy of each zip code $z$ proportionate to its weight $w_z$.
+The customer's zip code is chosen by sampling from the zip codes with the probabiltiy of choosing each zip code $z$ proportional to its weight $w_z$.
 
@@
 
-The item category for each purchase is chosen by sampling from the item categories with the probability of each item category $i$ proportionate to its weight $w_i$.
+The item category for each purchase is chosen by sampling from the item categories with the probability of choosing each item category $i$ proportional to its weight $w_i$.
 \textcolor{red}{$\lambda_i$ - maybe these should be per customer / item category instead of just per item category}

### 2014-07-23 bdcloud_paper/latex/paper.tex
Edited abstract, intro

@@
 Both developers and users of big data systems benefit from realistic example applications.
-For developers, such as applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.
+For developers, such applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.
 
-We present a new domain-driven, generative data model for BigPetStore, a big data application blueprint for the Hadoop ecosystem and included in the Apache BigTop distribution.
-We describe the model and demonstrate the its ability to generate semantically-rich data at variable scale ranging from a single machine to a large cluster.
+We present a new domain-driven, generative data model for BigPetStore, a big data application blueprint for the Hadoop ecosystem included in the Apache BigTop distribution.
+We describe the model and demonstrate its ability to generate semantically-rich data at variable scale ranging from a single machine to a large cluster.
 We validate the model by using the generated data to answer questions about customer locations and purchasing habits for a fictional targeted advertising campaign, a common business use case.
@@
 With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from the availability of realistic example applications.
-For developers, such as applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.
+For developers, such applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.
 
 BigPetStore is a big data application blueprint built around processing transaction data for a fictional chain of pet stores.
-BigPetStore mainly targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}.
+BigPetStore mainly targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating, and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}.
 BigPetStore has been incorporated into the open-source Apache BigTop distribution \cite{BigTop}, where it is used for testing and as a reference application.
@@
 Compared with BigPetStore's previous data generator, our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter.
-Combining the features of the TeraGen (scalability)  and MovieLens \cite{MovieLens} (semantically rich) input data sets commonly used for big data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.
+Combining the features of the TeraGen (scalability)  and MovieLens \cite{MovieLens} (semantically rich) input data sets commonly used for big data benchmarking, we present a synthetic data set generator which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.
 To demonstrate that the model generates data with desirable properties, we performed an example analysis designed to guide a fictional advertising campaign (a typical business use case).
@@
 \section{Implementation}
-The models and simulations were implemented using Python.
-The source code is available on GitHub \cite{GitHub} at \url{https://github.com/rnowling/bigpetstore-data-generator} under the Apache 2 license.
+The models and simulations were implemented using Python. The source code is available at \url{https://github.com/rnowling/bigpetstore-data-generator} under the Apache 2 license.
 

### 2014-07-23 bdcloud_paper/latex/paper.tex
Edited MSM equations

@@
 Each state $X_n$ corresponds to an item in that category.
-The customer's buying habits are determined by the transition probabilities (Eq.~\ref{eq:transition_probabilities}).
-The transition probabilities depend on three types of parameters: loopback weight $w_l$ indicating the probability of purchasing the same item again, weight $w_f$ of a particular field $f$, and the weight $w_{f,s}$ when two items $x$ and $y$ have the same value for the field $f$.
+The customer's buying habits are determined by the transition probabilities:
 
-\begin{align} \label{eq:transition_probabilities}
-Pr(X_{n+1}&=x|X_n) = \nonumber \\
+\begin{align}
+Pr(X_{n+1}&=x|X_n=y) = \nonumber \\
 & \left\{ 
   \begin{array}{l l}
-   (1.0 - w_l) W^{-1} \sum_f w_f w_{f,s}(x, y)  & \quad \text{if $x \neq X_n$}\\
-   w_l & \quad \text{if $x = X_n$}
+   (1.0 - p_l) \, Pr(x,y)  & \quad \text{if $x \neq y$}\\
+   p_l & \quad \text{if $x = y$}
   \end{array} \right.
@@
 
-\textcolor{red}{todo}
+where $w_l$ is the loopback probability, giving the probability of choosing the same item $y$ again.
+The function $Pr(x,y)$ gives the probability of choosing the item $x$ given that the previous item was $y$:
+
+\begin{equation*}
+Pr(x,y) = \frac{\sum_f w_f w_{f,s}(x, y)}{\sum_{j \neq x} \sum_f w_f w_{f,s}(x, j)}
+\end{equation*}
+
+where $w_f$ is the weight of a particular field $f$.   The pair of items $x$ and $y$ are weighted  according to the equality of the values for each field:
 
@@
 
-\textcolor{red}{todo}
+The weights $w_l$, $w_f$, and $w_{f, s}$ are chosen randomly for each customer:
 
-\begin{equation*}
-W = \sum_i \sum_{j \neq i} \sum_f w_f w_{f,s}(i, j)
-\end{equation*}
+\begin{align*}
+&w_l \sim N(\mu, \sigma^2) \\
+&w_f \sim N(\mu, \sigma^2) \\
+&w_{f, s} \sim N(\mu, \sigma^2)
+\end{align*}
 
-The weights $w_l$, $w_f$, and $w_{f, s}$ are chosen randomly for each customer.
-
-\textcolor{red}{description. min, max}
-
-\begin{align} \label{eq:item-weights}
-&w_l \sim N(\mu, \sigma^2) \nonumber \\
-&w_f \sim N(\mu, \sigma^2) \nonumber \\
-&w_{f, s} \sim N(\mu, \sigma^2)
-\end{align}
+If the weights are greater than 1, they are rounded down to 1.  Likese, if the weights are less than 0, they are rounded up to 0.
 

### 2014-07-23 bdcloud_paper/latex/paper.tex
Edited HMM equations

@@
 The proposed transaction time is found from the exhaustion times (Eq.~\ref{eq:transaction-times}).
-For each item category, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an exponential distribution parameterized separately for each customer $c$.
+For each item category, the offset time $T_{O, i}$ between when a customer would visit the store and the exhaustion time is sampled from an exponential distribution.
+The distribution is parameterized by $\beta_c$, the average number of days before triggering a transaction.
+The parameter $\beta_c$ is set separately for each customer $c$ by sampling from a uniform distribution.
 A proposed transaction time $T_{P, i}$ for each item category is found by subtracting the offset time $T_{O, i}$ from the exhaustion time $T_{E, i}$.
 The earliest proposed transaction time is taken as the overall proposed transaction time $T_T$.
-\textcolor{red}{$\beta_c$}
 
@@
 
-\begin{equation*}
-w_i = - \lambda_i \exp(-\lambda_i (T_{E, i}- T_{T}))
-\end{equation*}
+\begin{align*}
+&w_i = - \beta^{-1}_c \exp(-\beta^{-1}_c (T_{E, i}- T_T)) \\
+&\beta^{-1}_c \sim U(a,b)
+\end{align*}
 
-The item category for each purchase is chosen by sampling from the item categories with the probability of choosing each item category $i$ proportional to its weight $w_i$.
-\textcolor{red}{$\lambda_i$ - maybe these should be per customer / item category instead of just per item category}
+where $\beta_c$ the average number of days between when a customer purchases an item and the exhaustion time.
+The value $\beta_c$ is set separately for each customer by sampling from a uniform distribution.
+The value $w_i$ is used to model the propensity for a customer to purchase an item they will run out at time $T_{E, i}$ in the future when at the store at time $T_T$.
+
+The item category is chosen by sampling from the item categories with the probability of choosing each item category $i$ proportional to its weight $w_i$.
 
@@
 
-If the weights are greater than 1, they are rounded down to 1.  Likese, if the weights are less than 0, they are rounded up to 0.
+If the weights are greater than 1, they are rounded down to 1.  Likewise, if the weights are less than 0, they are rounded up to 0.
 

### 2014-07-23 bdcloud_paper/latex/paper.tex
Updated HMM equations

@@
 The HMM has three types of states: the start state, the purchase states, and the stop state.
-There are an infinite number of purchase states, identified by the exhaustion times of the customer's item categories, the transaction time, and the number of items purchased in the current transaction.
+There are an infinite number of purchase states, parameterized by the exhaustion times of the customer's item categories and the transaction time.
+
 The observables of the states correspond to the item categories.
@@
 
-\begin{align*}
+\begin{align} \label{eq:category-weights}
 &w_i = - \beta^{-1}_c \exp(-\beta^{-1}_c (T_{E, i}- T_T)) \\
-&\beta^{-1}_c \sim U(a,b)
-\end{align*}
+&\beta^{-1}_c \sim U(a,b) \nonumber
+\end{align}
 
@@
 
-\textcolor{red}{reorder so that probilities describe habits}
+%\textcolor{red}{reorder so that probilities describe habits}
 
-For example, if a customer tends to buy the same brand of dog food repeatedly, the outgoing edges to all products of the same brand as the product in the current state will be higher than those for other brands' products.
+%For example, if a customer tends to buy the same brand of dog food repeatedly, the outgoing edges to all products of the same brand as the product in the current state will be higher than those for other brands' products.
 If the customer does not care about flavor, the outgoing edges to all products from the current brand will be equal, otherwise the edge weights can be adjusted to prefer a specific combination of brand and flavor.
@@
 
-To simulate an item purchase, the Markov model is transitioned forward by one state.  The current states of the Markov models are kept between purchases.
+To simulate an item purchase, the chosen Markov model is transitioned forward by one state.
+The current states of the Markov models are kept between purchases.
+After an item is chosen, the item category's exhaustion time is updated by propagating the item category usage SDE described in Section~\ref{sec:transaction-times}, resulting in an implicit creation of the next possible purchase state in the HMM.
 
-After an item is chosen, the item category's exhaustion time is updated by propagating the item category usage SDE described in Section~\ref{sec:transaction-times}.
-\textcolor{red}{state transition, stop state}
+After each purchase, the HMM's current state is transitioned to either the new purchase state or the stop state.  The probability of choosing the next state is
 
-The HMM can be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.
+\begin{align*}
+&Pr(X_{n+1}=\text{stop}|X_n) = p(\text{stop})\\
+&Pr(X_{n+1}=\text{purchase}|X_n) = 1 - p(\text{stop})
+\end{align*}
+
+where $p(\text{stop})$ is the probability of choosing the stop state and is given by
+
+\begin{equation*}
+ p(\text{stop}) = \frac{w_{\text{stop}}}{w_{\text{stop}} + \sum_i w_i}
+\end{equation*}
+
+where $w_{\text{stop}}$ is the weight of the stop state and is a simulation parameter and the weight $w_i$ of each item category $i$ was given earlier in Eq.~\ref{eq:category-weights}.
+
+%The HMM can be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.
 

### 2014-07-23 bdcloud_paper/latex/paper.tex
Updated HMM conclusion
edits to discussion & conclusion

@@
 
+If the stop state is chosen, the transaction is finished and the next transaction is started according to the procedure presented in Section~\ref{sec:transaction-times}.
+
 %The HMM can be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.
@@
 \section{Implementation}
-The models and simulations were implemented using Python. The source code is available at \url{https://github.com/rnowling/bigpetstore-data-generator} under the Apache 2 license.
+The models and simulations were implemented using Python. The source code is available at \url{https://github.com/rnowling/bigpetstore-data-generator} under the Apache Public License v2.
 
@@
 Scalability testing combined with analysis of the implementation suggests that the data generation method can scale from small data sets appropriate for desktop development to large data sets for testing and benchmarking of clusters.
-We have released the implementation source code under the Apache 2 open-source license.
+We have released the implementation source code under the open-source Apache Public License v2.
 
@@
 Deeper integration of existing fields such as location and purchasing profiles (e.g., to model regional purchasing preferences) will increase the variety of the semantic information encoded in the generated data.
-The incorporation of weather and climate data can be used to influence when customers shop and the types of products they buy.
-For example, customers would be less likely to shop during snow storms and more likely to buy winter apparel for their pets.
+The incorporation of weather and climate data can be used to influence when customers shop, the types of products they buy, and the amount of products purchased.
+For example, customers would be less likely to shop during snow storms, more likely to buy items such as winter apparel for their pets, and more likely to purchase bulk quantities to reduce the number of transactions.
 Modeling of time-dependent events such as sales, evolution of customer purchasing profiles over time, and ``life events,'' such as the birth or passing of pets, will enable more interesting time-series analysis of the data.

### 2014-07-23 bdcloud_paper/latex/paper.tex
edits to discussion & conclusion

@@
 Modeling of time-dependent events such as sales, evolution of customer purchasing profiles over time, and ``life events,'' such as the birth or passing of pets, will enable more interesting time-series analysis of the data.
-We would also like to expand the scope of the model to incorporate business processes (such as inventory management, customer complaints, employees, etc.), thus enabling queries about how internal business process affect customer behavior.
+We would also like to expand the scope of the model to incorporate business processes (such as inventory management, customer complaints, employees, etc.), thus enabling queries about the relationship between internal business process and customer behavior.
 

### 2014-07-23 bdcloud_paper/latex/paper.tex
edits

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Domain-Driven, Generative Data Model for BigPetStore}
+\title{A Domain-Driven, Generative Data Model for BigPetStore}
 
@@
 \begin{abstract}
-Both developers and users of big data systems benefit from realistic example applications.
+Generating large amounts of semantically-rich data for testing big data workflows is paramount for scalable performance benchmarking and quality assurance in modern machine-learning and analytics workloads.
+The most obvious use case for such a generative algorithm is in conjunction with a big data application blueprint, which can be used by developers (to test their emerging big data solutions) as well as end users (as a starting point for validating infrastructure installations, building novel applications, and learning analytics methods).
+
+%Both developers and users of big data systems benefit from realistic example applications.
 For developers, such applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.
@@
 
-%Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key for consistent development, testing, and benchmarking of big data software, algorithms, and infrastructure.
-BigPetStore is a big data application blueprint, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics, built around a fictional chain of pet store which has been used for testing and benchmarking in the Apache BigTop distribution.
-We present a new generative data model for BigPetStore and demonstrate the model's ability to generate semantically-rich data that can be used for complex analytics workflows at variable scale ranging from a single machine to a large cluster.
 \end{abstract}
-% IEEEtran.cls defaults to using nonbold math in the Abstract.
-% This preserves the distinction between vectors and scalars. However,
-% if the conference you are submitting to favors bold math in the abstract,
-% then you can use LaTeX's standard command \boldmath at the very start
-% of the abstract to achieve this. Many IEEE journals/conferences frown on
-% math in the abstract anyway.
-
-% no keywords
 
@@
 \section{Introduction}
-Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.
+Big data applications process large, dynamics, multidimensional data sets with the general goal of information and knowledge extraction.
 With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from the availability of realistic example applications.
@@
 
-BigPetStore is a big data application blueprint built around processing transaction data for a fictional chain of pet stores.
-BigPetStore mainly targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating, and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}.
+BigPetStore is one such big data application blueprint built around processing transaction data for a fictional chain of pet stores.
+BigPetStore targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating, and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}.
 BigPetStore has been incorporated into the open-source Apache BigTop distribution \cite{BigTop}, where it is used for testing and as a reference application.
@@
 
-A variety of approaches for generating synthetic data set exist.
+A variety of approaches for generating synthetic data sets exist.
 TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} are popular tools that can generate data sets quickly and at any scale, but the resulting data is semantically-void and not useful for much more than simple benchmarks.
-Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}.
-Several approaches even provide domain-specific languages \cite{Bruno2005,Hoag2007} for specifying additional constraints such as which distributions to sample from and allow for modeling basic relationships between fields using simple arithmetic equations \cite{Alexandrov2012}.
-Arasu, et al.
+Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}, and several approaches even provide domain-specific languages \cite{Bruno2005,Hoag2007} for specifying additional constraints (such as which distributions to sample from and allow for modeling basic relationships between fields using simple arithmetic equations \cite{Alexandrov2012}).
+
+Recent work has addressed some aspects of the need for more dynamic data set generation: Arasu, et al.
 \cite{Arasu2011} demonstrated that constraint-solving techniques could be used as an alternative to procedural approaches.
-These frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for BigPetStore's analytics examples.
+Such frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for BigPetStore's analytics examples.
 
@@
 
-Until such methods have reached maturity, BigPetStore's needs are best met by a customized, domain-driven model.
+Until such purely generic methods reach maturity, BigPetStore's needs are best met by a customized, domain-driven model.
 Another example of such a model is The Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006}.
@@
 To demonstrate that the model generates data with desirable properties, we performed an example analysis designed to guide a fictional advertising campaign (a typical business use case).
-The model's implementation available as open source software.
+The model's implementation was made available as open source software.
 
@@
 Customer records consist of a unique identifier, a name, and a location given as a 5-digit zip code.
-Transaction records consist of a customer identifier, a store identifier, and a time given in days since the beginning of the simulation.
+Transaction records consist of a transaction identifier, a customer identifier, a store identifier, and a time given in days since the beginning of the simulation.
 Transaction item records contain a transaction identifier and an item description.
@@
 \begin{align*}
-p(\text{location}=z) = &Z^-1 p(\text{location}=z | \text{population}(z)) \\
+p(\text{location}=z) = &Z p(\text{location}=z | \text{population}(z)) \\
 &p(\text{location}=z | \text{income}(z))
@@
 We are working to implement the model using Hadoop and Spark.
-We will commit the resulting implementations to the Apache BigTop project to enable ease of access and immediate benefit to current users.
+We will commit the resulting implementations to BigPetStore hosted in the Apache BigTop distribution to enable ease of access and immediate benefit to current users.
 

### 2014-07-23 bdcloud_paper/latex/paper.tex
added to acknowledgements

@@
 Clare and Casey Robinson for useful and interesting discussion.
+Trevor M.
+Cickovski, Scott A.
+Hale, Hank Jakiela, Casey Robinson, and Douglas Thain provided invaluable feedback on the manuscript.
 JV would like to thank the Apache BigTop community for their interest in and support of BigPetStore.
-We would also like to thank Matt Fenwick, Nigel Savage, and Bhashit Parikh for their contributions to BigPetStore.
+The authors would also like to thank Matt Fenwick, Nigel Savage, and Bhashit Parikh for their contributions to BigPetStore.
 The authors would like to thank Red Hat, Inc.

### 2014-10-12 bdcloud_paper/latex/paper.tex
delete deade code
Fix typos found by reviewers

@@
 
-%Both developers and users of big data systems benefit from realistic example applications.
-For developers, such applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.
-
 We present a new domain-driven, generative data model for BigPetStore, a big data application blueprint for the Hadoop ecosystem included in the Apache BigTop distribution.
@@
 \section{Introduction}
-Big data applications process large, dynamics, multidimensional data sets with the general goal of information and knowledge extraction.
+Big data applications process large, dynamic, multidimensional data sets with the general goal of information and knowledge extraction.
 With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from the availability of realistic example applications.
@@
 Another example of such a model is The Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006}.
-IDSG was developed to to test the effectiveness of IDASs in identifying potential terrorism threat scenarios using synthetic data.
+IDSG was developed to test the effectiveness of IDASs in identifying potential terrorism threat scenarios using synthetic data.
 Like BigPetStore, the developers of IDSG needed to avoid licensing restrictions and privacy issues associated with real data.

### 2014-10-12 bdcloud_paper/latex/paper.tex
Added discussion of Java implementation.

@@
 The current implementation was prototyped in Python.
-We are working to implement the model using Hadoop and Spark.
-We will commit the resulting implementations to BigPetStore hosted in the Apache BigTop distribution to enable ease of access and immediate benefit to current users.
+We are finishing a Java implementation that can be used from the command-line, Hadoop, and Spark, enabling massive parallelization and improved scaling.
+As part of the effort to re-write the framework in Java, we are abstracting classes for statistical modeling and simulation for reuse in developing data generators for other domains.
+We will commit the resulting implementation to BigPetStore hosted in the Apache BigTop distribution to enable ease of access and immediate benefit to current users.
 

### 2014-10-12 bdcloud_paper/latex/paper.tex
Add people to acknowledgements.
Add log synth reference

@@
 TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} are popular tools that can generate data sets quickly and at any scale, but the resulting data is semantically-void and not useful for much more than simple benchmarks.
-Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}, and several approaches even provide domain-specific languages \cite{Bruno2005,Hoag2007} for specifying additional constraints (such as which distributions to sample from and allow for modeling basic relationships between fields using simple arithmetic equations \cite{Alexandrov2012}).
+Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}, and several approaches even provide domain-specific languages \cite{LogSynth,Bruno2005,Hoag2007} for specifying additional constraints (such as which distributions to sample from and allow for modeling basic relationships between fields using simple arithmetic equations \cite{Alexandrov2012}).
 
@@
 Clare and Casey Robinson for useful and interesting discussion.
-Trevor M.
-Cickovski, Scott A.
+Will Benton, Trevor M.
+Cickovski, Erik Erlandson, Scott A.
 Hale, Hank Jakiela, Casey Robinson, and Douglas Thain provided invaluable feedback on the manuscript.

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/purchasingprofile/PurchasingProfileSamplerBuilder.java
Move things into proper packages

@@ -6,6 +6,7 @@ import java.util.Map;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
+import com.github.rnowling.bps.datagenerator.samplers.purchasingprofile.ProductCategoryMarkovModelSampler;
 import com.github.rnowling.bps.datagenerator.samplers.purchasingprofile.PurchasingProfileSampler;
 import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/datamodels/simulation/PurchasingProfileBuilder.java

@@ -1,9 +1,7 @@
-package com.github.rnowling.bps.datagenerator.builders.purchasingprofile;
+package com.github.rnowling.bps.datagenerator.datamodels.simulation;
 
 import java.util.Map;
 
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
 import com.google.common.collect.Maps;
 

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/samplers/purchasingprofile/ProductCategoryMarkovModelSampler.java

@@ -1,4 +1,4 @@
-package com.github.rnowling.bps.datagenerator.builders.purchasingprofile;
+package com.github.rnowling.bps.datagenerator.samplers.purchasingprofile;
 
 import java.util.Map;
 

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/samplers/purchasingprofile/PurchasingProfileSampler.java

@@ -2,10 +2,10 @@ package com.github.rnowling.bps.datagenerator.samplers.purchasingprofile;
 
 import java.util.Map;
 
-import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
 import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
 

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/builders/transaction/TestPurchasingProcessesBuilder.java

@@ -8,10 +8,10 @@ import java.util.Map;
 import org.junit.Test;
 
 import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProcessesBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.samplers.transaction.PurchasingProcesses;
 import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/datamodels/simulation/TestPurchasingProfileBuilder.java

@@ -1,4 +1,4 @@
-package com.github.rnowling.bps.datagenerator.builders.purchasingprofile;
+package com.github.rnowling.bps.datagenerator.datamodels.simulation;
 
 import static org.junit.Assert.assertNotNull;
 import static org.junit.Assert.assertThat;
@@ -9,9 +9,9 @@ import java.util.Map;
 
 import org.junit.Test;
 
-import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModelBuilder;
 import com.google.common.collect.Lists;

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/samplers/purchasingprofile/TestProductCategoryMarkovModelSampler.java

@@ -1,4 +1,4 @@
-package com.github.rnowling.bps.datagenerator.builders.purchasingprofile;
+package com.github.rnowling.bps.datagenerator.samplers.purchasingprofile;
 
 import static org.junit.Assert.assertNotNull;
 import static org.junit.Assert.assertTrue;
@@ -8,10 +8,10 @@ import java.util.List;
 import org.junit.Test;
 
 import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.ProductCategoryMarkovModelSampler;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datareaders.ProductsReader;
+import com.github.rnowling.bps.datagenerator.samplers.purchasingprofile.ProductCategoryMarkovModelSampler;
 import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
 

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/samplers/transaction/TestTransactionPurchasesHiddenMarkovModel.java

@@ -9,7 +9,6 @@ import java.util.Map;
 import org.junit.Test;
 
 import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerTransactionParametersSamplerBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
@@ -20,6 +19,7 @@ import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Purchase;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProcessesBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datareaders.ProductCategoryBuilder;
 import com.github.rnowling.bps.datagenerator.samplers.transaction.PurchasingProcesses;
 import com.github.rnowling.bps.datagenerator.samplers.transaction.TransactionPurchasesHiddenMarkovModel;

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/transaction/CustomerInventoryBuilder.java
Re-organize packages

@@ -1,10 +1,13 @@
-package com.github.rnowling.bps.datagenerator.datamodels.simulation;
+package com.github.rnowling.bps.datagenerator.builders.transaction;
 
 import java.util.Collection;
 import java.util.List;
 import java.util.Map;
 
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerInventory;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.ProductCategoryInventory;
 import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
 import com.google.common.collect.Lists;
 import com.google.common.collect.Maps;

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/transaction/CustomerTransactionParametersBuilder.java

@@ -1,6 +1,7 @@
-package com.github.rnowling.bps.datagenerator.datamodels.simulation;
+package com.github.rnowling.bps.datagenerator.builders.transaction;
 
 import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
 import com.google.common.collect.HashMultiset;
 import com.google.common.collect.Multiset;
 

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/transaction/PurchasingProcessesBuilder.java

@@ -1,7 +1,9 @@
-package com.github.rnowling.bps.datagenerator.datamodels.simulation;
+package com.github.rnowling.bps.datagenerator.builders.transaction;
 
 import java.util.Map;
 
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
 import com.github.rnowling.bps.datagenerator.samplers.transaction.PurchasingProcesses;
 import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/transaction/PurchasingProfileBuilder.java

@@ -1,7 +1,9 @@
-package com.github.rnowling.bps.datagenerator.datamodels.simulation;
+package com.github.rnowling.bps.datagenerator.builders.transaction;
 
 import java.util.Map;
 
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
 import com.google.common.collect.Maps;
 

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/transaction/TransactionPurchasesSamplerBuilder.java

@@ -4,10 +4,8 @@ import java.util.Collection;
 
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerInventory;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerInventoryBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Purchase;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProcessesBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
 import com.github.rnowling.bps.datagenerator.samplers.transaction.PurchasingProcesses;
 import com.github.rnowling.bps.datagenerator.samplers.transaction.TransactionPurchasesHiddenMarkovModel;

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/samplers/purchasingprofile/PurchasingProfileSampler.java

@@ -2,10 +2,10 @@ package com.github.rnowling.bps.datagenerator.samplers.purchasingprofile;
 
 import java.util.Map;
 
+import com.github.rnowling.bps.datagenerator.builders.transaction.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
 import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
 

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/samplers/transaction/CustomerTransactionParametersSampler.java

@@ -1,8 +1,8 @@
 package com.github.rnowling.bps.datagenerator.samplers.transaction;
 
+import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerTransactionParametersBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParametersBuilder;
 import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
 
 public class CustomerTransactionParametersSampler implements Sampler<CustomerTransactionParameters>

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/builders/transaction/TestCustomerInventoryBuilder.java

@@ -12,7 +12,6 @@ import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerTransa
 import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerInventory;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerInventoryBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datareaders.ProductCategoryBuilder;

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/builders/transaction/TestCustomerTransactionParametersBuilder.java

@@ -1,12 +1,12 @@
-package com.github.rnowling.bps.datagenerator.datamodels.simulation;
+package com.github.rnowling.bps.datagenerator.builders.transaction;
 
 import static org.junit.Assert.assertTrue;
 
 import org.junit.Test;
 
+import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerTransactionParametersBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParametersBuilder;
 
 public class TestCustomerTransactionParametersBuilder
 {

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/builders/transaction/TestPurchasingProcessesBuilder.java

@@ -9,9 +9,7 @@ import org.junit.Test;
 
 import com.github.rnowling.bps.datagenerator.Constants;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProcessesBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.samplers.transaction.PurchasingProcesses;
 import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/builders/transaction/TestPurchasingProfileBuilder.java

@@ -1,4 +1,4 @@
-package com.github.rnowling.bps.datagenerator.datamodels.simulation;
+package com.github.rnowling.bps.datagenerator.builders.transaction;
 
 import static org.junit.Assert.assertNotNull;
 import static org.junit.Assert.assertThat;
@@ -9,9 +9,9 @@ import java.util.Map;
 
 import org.junit.Test;
 
+import com.github.rnowling.bps.datagenerator.builders.transaction.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModelBuilder;
 import com.google.common.collect.Lists;

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/samplers/transaction/TestTransactionPurchasesHiddenMarkovModel.java

@@ -9,17 +9,17 @@ import java.util.Map;
 import org.junit.Test;
 
 import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerInventoryBuilder;
 import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerTransactionParametersSamplerBuilder;
+import com.github.rnowling.bps.datagenerator.builders.transaction.PurchasingProcessesBuilder;
+import com.github.rnowling.bps.datagenerator.builders.transaction.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerInventory;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerInventoryBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Purchase;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProcessesBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datareaders.ProductCategoryBuilder;
 import com.github.rnowling.bps.datagenerator.samplers.transaction.PurchasingProcesses;
 import com.github.rnowling.bps.datagenerator.samplers.transaction.TransactionPurchasesHiddenMarkovModel;

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/purchasingprofile/PurchasingProfileBuilder.java
Re-organize packages

@@ -1,4 +1,4 @@
-package com.github.rnowling.bps.datagenerator.builders.transaction;
+package com.github.rnowling.bps.datagenerator.builders.purchasingprofile;
 
 import java.util.Map;
 

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/samplers/purchasingprofile/PurchasingProfileSampler.java

@@ -2,7 +2,7 @@ package com.github.rnowling.bps.datagenerator.samplers.purchasingprofile;
 
 import java.util.Map;
 
-import com.github.rnowling.bps.datagenerator.builders.transaction.PurchasingProfileBuilder;
+import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/builders/purchasingprofile/TestPurchasingProfileBuilder.java

@@ -1,4 +1,4 @@
-package com.github.rnowling.bps.datagenerator.builders.transaction;
+package com.github.rnowling.bps.datagenerator.builders.purchasingprofile;
 
 import static org.junit.Assert.assertNotNull;
 import static org.junit.Assert.assertThat;
@@ -9,7 +9,7 @@ import java.util.Map;
 
 import org.junit.Test;
 
-import com.github.rnowling.bps.datagenerator.builders.transaction.PurchasingProfileBuilder;
+import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/builders/transaction/TestPurchasingProcessesBuilder.java

@@ -8,6 +8,7 @@ import java.util.Map;
 import org.junit.Test;
 
 import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
 import com.github.rnowling.bps.datagenerator.samplers.transaction.PurchasingProcesses;

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/samplers/transaction/TestTransactionPurchasesHiddenMarkovModel.java

@@ -9,10 +9,10 @@ import java.util.Map;
 import org.junit.Test;
 
 import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerInventoryBuilder;
 import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerTransactionParametersSamplerBuilder;
 import com.github.rnowling.bps.datagenerator.builders.transaction.PurchasingProcessesBuilder;
-import com.github.rnowling.bps.datagenerator.builders.transaction.PurchasingProfileBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerInventory;

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/purchasingprofile/PurchasingProfileSamplerBuilder.java
Abstract ProductCategoryMarkovModelSampler a bit

@@ -3,6 +3,7 @@ package com.github.rnowling.bps.datagenerator.builders.purchasingprofile;
 import java.util.Collection;
 import java.util.Map;
 
+import com.github.rnowling.bps.datagenerator.Constants;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
@@ -10,6 +11,7 @@ import com.github.rnowling.bps.datagenerator.samplers.purchasingprofile.ProductC
 import com.github.rnowling.bps.datagenerator.samplers.purchasingprofile.PurchasingProfileSampler;
 import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
+import com.github.rnowling.bps.datagenerator.statistics.samplers.BoundedMultiModalGaussianSampler;
 import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
 import com.google.common.collect.ImmutableList;
 import com.google.common.collect.Maps;
@@ -18,11 +20,29 @@ public class PurchasingProfileSamplerBuilder
 {
 	final ImmutableList<ProductCategory> productCategories;
 	final SeedFactory seedFactory;
+	final Sampler<Double> fieldWeightSampler;
+	final Sampler<Double> fieldSimilarityWeightSampler;
+	final Sampler<Double> loopbackWeightSampler;
 	
 	public PurchasingProfileSamplerBuilder(Collection<ProductCategory> productCategories, SeedFactory seedFactory)
 	{
 		this.productCategories = ImmutableList.copyOf(productCategories);
 		this.seedFactory = seedFactory;
+		
+		this.fieldWeightSampler = new BoundedMultiModalGaussianSampler(Constants.PRODUCT_MSM_FIELD_WEIGHT_GAUSSIANS, 
+				Constants.PRODUCT_MSM_FIELD_WEIGHT_LOWERBOUND, 
+				Constants.PRODUCT_MSM_FIELD_WEIGHT_UPPERBOUND,
+				seedFactory);
+	
[…]

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/samplers/purchasingprofile/ProductCategoryMarkovModelSampler.java

@@ -2,45 +2,33 @@ package com.github.rnowling.bps.datagenerator.samplers.purchasingprofile;
 
 import java.util.Map;
 
-import com.github.rnowling.bps.datagenerator.Constants;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModelBuilder;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.BoundedMultiModalGaussianSampler;
 import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
 import com.google.common.collect.Maps;
 
 public class ProductCategoryMarkovModelSampler implements Sampler<MarkovModel<Product>>
 {
-	ProductCategory productCategory;
-	Sampler<Double> fieldWeightSampler;
-	Sampler<Double> fieldSimilarityWeightSampler;
-	Sampler<Double> loopbackWeightSampler;
+	final ProductCategory productCategory;
+	final Sampler<Double> fieldWeightSampler;
+	final Sampler<Double> fieldSimilarityWeightSampler;
+	final Sampler<Double> loopbackWeightSampler;
 	
 	Map<String, Double> fieldWeights;
 	Map<String, Double> fieldSimilarityWeights;
 	double loopbackWeight;
 	
-	public ProductCategoryMarkovModelSampler(ProductCategory productCategory, SeedFactory seedFactory)
+	public ProductCategoryMarkovModelSampler(ProductCategory productCategory, 
+			Sampler<Double> fieldWeightSampler, Sampler<Double> fieldSimilarityWeightSampler,
+			Sampler<Double> loopbackWeightSampler)
 	{
 		this.productCategory = productCategory;
 		
-		this.fieldWeightSampler = new BoundedMultiModalGaussianSampler(Constants.PRODUCT_MSM_FIELD_WEIGHT_GAUSSIANS, 
[…]

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/statistics/samplers/UniformSampler.java

@@ -1,0 +1,20 @@
+package com.github.rnowling.bps.datagenerator.statistics.samplers;
+
+import java.util.Random;
+
+import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
+
+public class UniformSampler implements Sampler<Double>
+{
+	final Random rng;
+	
+	public UniformSampler(SeedFactory seedFactory)
+	{
+		rng = new Random(seedFactory.getNextSeed());
+	}
+	
+	public Double sample()
+	{
+		return rng.nextDouble();
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/samplers/purchasingprofile/TestProductCategoryMarkovModelSampler.java

@@ -11,9 +11,9 @@ import com.github.rnowling.bps.datagenerator.Constants;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
 import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
 import com.github.rnowling.bps.datagenerator.datareaders.ProductsReader;
-import com.github.rnowling.bps.datagenerator.samplers.purchasingprofile.ProductCategoryMarkovModelSampler;
 import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
 import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
+import com.github.rnowling.bps.datagenerator.statistics.samplers.UniformSampler;
 
 public class TestProductCategoryMarkovModelSampler
 {
@@ -29,7 +29,9 @@ public class TestProductCategoryMarkovModelSampler
 		
 		ProductCategory productCategory = productCategories.get(0);
 		
-		ProductCategoryMarkovModelSampler generator = new ProductCategoryMarkovModelSampler(productCategory, seedFactory);
+		ProductCategoryMarkovModelSampler generator = new ProductCategoryMarkovModelSampler(productCategory, 
+				new UniformSampler(seedFactory), new UniformSampler(seedFactory), new UniformSampler(seedFactory)
+				);
 		
 		MarkovModel<Product> model = generator.sample();
 		

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/CustomerGenerator.java
Create high-level generator classes

@@ -1,0 +1,26 @@
+package com.github.rnowling.bps.datagenerator;
+
+import java.util.List;
+
+import com.github.rnowling.bps.datagenerator.builders.customer.CustomerSamplerBuilder;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.InputData;
+import com.github.rnowling.bps.datagenerator.datamodels.outputs.Customer;
+import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
+import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
+import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
+
+public class CustomerGenerator
+{
+	final Sampler<Customer> sampler;
+	
+	public CustomerGenerator(InputData inputData, List<Store> stores, SeedFactory seedFactory)
+	{
+		CustomerSamplerBuilder builder = new CustomerSamplerBuilder(stores, inputData, seedFactory);
+		sampler = builder.build();
+	}
+	
+	public Customer generate() throws Exception
+	{
+		return sampler.sample();
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/PurchasingProfileGenerator.java

@@ -1,0 +1,25 @@
+package com.github.rnowling.bps.datagenerator;
+
+import java.util.Collection;
+
+import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.PurchasingProfileSamplerBuilder;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
+import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
+import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
+
+public class PurchasingProfileGenerator
+{
+	final Sampler<PurchasingProfile> sampler;
+	
+	public PurchasingProfileGenerator(Collection<ProductCategory> productCategories, SeedFactory seedFactory) throws Exception
+	{
+		PurchasingProfileSamplerBuilder builder = new PurchasingProfileSamplerBuilder(productCategories, seedFactory);
+		sampler = builder.build();
+	}
+	
+	public PurchasingProfile generate() throws Exception
+	{
+		return sampler.sample();
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/StoreGenerator.java

@@ -1,0 +1,23 @@
+package com.github.rnowling.bps.datagenerator;
+
+import com.github.rnowling.bps.datagenerator.builders.store.StoreSamplerBuilder;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.InputData;
+import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
+import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
+import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
+
+public class StoreGenerator
+{
+	final Sampler<Store> sampler;
+	
+	public StoreGenerator(InputData inputData, SeedFactory seedFactory)
+	{	
+		StoreSamplerBuilder builder = new StoreSamplerBuilder(inputData.getZipcodeTable(), seedFactory);
+		sampler = builder.build();
+	}
+	
+	public Store generate() throws Exception
+	{
+		return sampler.sample();
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/TransactionGenerator.java

@@ -1,0 +1,30 @@
+package com.github.rnowling.bps.datagenerator;
+
+import java.util.Collection;
+import java.util.List;
+
+import com.github.rnowling.bps.datagenerator.builders.transaction.TransactionSamplerBuilder;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
+import com.github.rnowling.bps.datagenerator.datamodels.outputs.Customer;
+import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
+import com.github.rnowling.bps.datagenerator.datamodels.outputs.Transaction;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
+import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
+import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
+
+public class TransactionGenerator
+{
+	Sampler<Transaction> sampler;
+	
+	public TransactionGenerator(Customer customer, PurchasingProfile profile, List<Store> stores,
+			Collection<ProductCategory> productCategories, SeedFactory seedFactory) throws Exception
+	{
+		sampler = new TransactionSamplerBuilder(stores, productCategories,
+				customer, profile, seedFactory).build();
+	}
+	
+	public Transaction generate() throws Exception
+	{
+		return sampler.sample();
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/cli/Simulation.java

@@ -6,10 +6,10 @@ import java.util.List;
 import java.util.Vector;
 
 import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.builders.customer.CustomerSamplerBuilder;
-import com.github.rnowling.bps.datagenerator.builders.purchasingprofile.PurchasingProfileSamplerBuilder;
-import com.github.rnowling.bps.datagenerator.builders.store.StoreSamplerBuilder;
-import com.github.rnowling.bps.datagenerator.builders.transaction.TransactionSamplerBuilder;
+import com.github.rnowling.bps.datagenerator.CustomerGenerator;
+import com.github.rnowling.bps.datagenerator.PurchasingProfileGenerator;
+import com.github.rnowling.bps.datagenerator.StoreGenerator;
+import com.github.rnowling.bps.datagenerator.TransactionGenerator;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.InputData;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.Names;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
@@ -22,7 +22,6 @@ import com.github.rnowling.bps.datagenerator.datareaders.NameReader;
 import com.github.rnowling.bps.datagenerator.datareaders.ProductsReader;
 import com.github.rnowling.bps.datagenerator.datareaders.ZipcodeReader;
 import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
 import com.google.common.collect.Lists;
 
 public class Simulation
@@ -36,6 +35,7 @@ public class Simulation
 	
 	List<Store> stores;
 	List<Customer> customers;
+	List<PurchasingProfile> purchasingProfiles;
 	List<Transaction> transactions;
 	
 	public Simulation(int nStores, int nCustomers, double simulationTime, long seed)
@@ -69,16 +69,15 @@ public class Simulation
 		inputData = new InputData(zipcodeTable, names);
 	}
 	
-	private void generateStores(int nStores) throws Exception
+	private void generateStores() throws Exception
 	{
 		System.out.println("Generating stores");
-		StoreSamplerBuilder builder = new StoreSamplerBuilder(inputData.getZipcodeTable(), this.seedFactory);
-		Sampler<Store> storeSampler = builder.build();
[…]

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/Generator.java
Remove unneeded interface

@@ -1,6 +1,0 @@
-package com.github.rnowling.bps.datagenerator.builders;
-
-public interface Generator<T>
-{
-	public T generate() throws Exception;
-}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/CustomerGenerator.java
Re-organize packages

@@ -2,12 +2,12 @@ package com.github.rnowling.bps.datagenerator;
 
 import java.util.List;
 
-import com.github.rnowling.bps.datagenerator.builders.customer.CustomerSamplerBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.InputData;
 import com.github.rnowling.bps.datagenerator.datamodels.outputs.Customer;
 import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.github.rnowling.bps.datagenerator.generators.customer.CustomerSamplerBuilder;
 
 public class CustomerGenerator
 {

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/StoreGenerator.java

@@ -1,10 +1,10 @@
 package com.github.rnowling.bps.datagenerator;
 
-import com.github.rnowling.bps.datagenerator.builders.store.StoreSamplerBuilder;
 import com.github.rnowling.bps.datagenerator.datamodels.inputs.InputData;
 import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.github.rnowling.bps.datagenerator.generators.store.StoreSamplerBuilder;
 
 public class StoreGenerator
 {

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/customer/CustomerSamplerBuilder.java

@@ -1,46 +1,0 @@
-package com.github.rnowling.bps.datagenerator.builders.customer;
-
-import java.util.List;
-
-import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.InputData;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ZipcodeRecord;
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Customer;
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
-import com.github.rnowling.bps.datagenerator.pdfs.customer.CustomerLocationPDF;
-import com.github.rnowling.bps.datagenerator.samplers.customer.CustomerSampler;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.pdfs.ProbabilityDensityFunction;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.RouletteWheelSampler;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.SequenceSampler;
-
-public class CustomerSamplerBuilder
-{
-	private final List<Store> stores;
-	private final InputData inputData;
-	private final SeedFactory seedFactory;
-	
-	public CustomerSamplerBuilder(List<Store> stores, InputData inputData, SeedFactory seedFactory)
-	{
-		this.stores = stores;
-		this.seedFactory = seedFactory;
-		this.inputData = inputData;
-		
-	}
-	
-	public Sampler<Customer> build()
-	{
-		ProbabilityDensityFunction<ZipcodeRecord> locationPDF = new CustomerLocationPDF(inputData.getZipcodeTable(), stores,
-				Constants.AVERAGE_CUSTOMER_STORE_DISTANCE);
-		
-		Sampler<Integer> idSampler = new SequenceSampler();
-		Sampler<String> firstNameSampler = RouletteWheelSampler.create(inputData.getNames().getFirstNames(), seedFactory);
-		Sampler<String> lastNameSampler = RouletteWheelSampler.create(inputData.getNames().getLastNames(), seedFactory);
-		Sampler<ZipcodeRecord> locationSampler = RouletteWheelSampler.create(inputData.getZipcodeTable(), 
-				locationPDF, seedFactory);
-		
[…]

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/store/StoreSamplerBuilder.java

@@ -1,45 +1,0 @@
-package com.github.rnowling.bps.datagenerator.builders.store;
-
-import java.util.List;
-
-import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ZipcodeRecord;
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
-import com.github.rnowling.bps.datagenerator.pdfs.store.StoreLocationIncomePDF;
-import com.github.rnowling.bps.datagenerator.pdfs.store.StoreLocationPopulationPDF;
-import com.github.rnowling.bps.datagenerator.samplers.store.StoreSampler;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.pdfs.JointPDF;
-import com.github.rnowling.bps.datagenerator.statistics.pdfs.ProbabilityDensityFunction;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.RouletteWheelSampler;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.SequenceSampler;
-
-public class StoreSamplerBuilder
-{
-	private final List<ZipcodeRecord> zipcodeTable;
-	private final SeedFactory seedFactory;
-	
-	public StoreSamplerBuilder(List<ZipcodeRecord> zipcodeTable, SeedFactory seedFactory)
-	{
-		this.zipcodeTable = zipcodeTable;
-		this.seedFactory = seedFactory;
-	}
-	
-	public Sampler<Store> build()
-	{
-		Sampler<Integer> idSampler = new SequenceSampler();
-		
-		ProbabilityDensityFunction<ZipcodeRecord> locationPopulationPDF = 
-				new StoreLocationPopulationPDF(zipcodeTable);
-		ProbabilityDensityFunction<ZipcodeRecord> locationIncomePDF = 
-				new StoreLocationIncomePDF(zipcodeTable, Constants.INCOME_SCALING_FACTOR);
-		ProbabilityDensityFunction<ZipcodeRecord> locationJointPDF = 
-				new JointPDF<ZipcodeRecord>(zipcodeTable, locationPopulationPDF, locationIncomePDF);
-		
[…]

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/transaction/CustomerTransactionParametersBuilder.java

@@ -1,43 +1,0 @@
-package com.github.rnowling.bps.datagenerator.builders.transaction;
-
-import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
-import com.google.common.collect.HashMultiset;
-import com.google.common.collect.Multiset;
-
-public class CustomerTransactionParametersBuilder
-{
-	private Multiset<PetSpecies> petCounts;
-	private double averageTransactionTriggerTime;
-	private double averagePurchaseTriggerTime;
-	
-	public CustomerTransactionParametersBuilder()
-	{
-		this.petCounts = HashMultiset.create();
-		this.averagePurchaseTriggerTime = 0.0;
-		this.averageTransactionTriggerTime = 0.0;
-	}
-	
-	public void addPet(PetSpecies species)
-	{
-		this.petCounts.add(species);
-	}
-
-	public void setAverageTransactionTriggerTime(
-			double averageTransactionTriggerTime)
-	{
-		this.averageTransactionTriggerTime = averageTransactionTriggerTime;
-	}
-
-	public void setAveragePurchaseTriggerTime(double averagePurchaseTriggerTime)
-	{
-		this.averagePurchaseTriggerTime = averagePurchaseTriggerTime;
-	}
-	
-	public CustomerTransactionParameters build()
-	{
-		return new CustomerTransactionParameters(this.petCounts,
-				this.averageTransactionTriggerTime,
-				this.averagePurchaseTriggerTime);
-	}
-}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/builders/transaction/TransactionPurchasesSamplerBuilder.java

@@ -1,55 +1,0 @@
-package com.github.rnowling.bps.datagenerator.builders.transaction;
-
-import java.util.Collection;
-
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerInventory;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.Purchase;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.PurchasingProfile;
-import com.github.rnowling.bps.datagenerator.samplers.transaction.PurchasingProcesses;
-import com.github.rnowling.bps.datagenerator.samplers.transaction.TransactionPurchasesHiddenMarkovModel;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-
-public class TransactionPurchasesSamplerBuilder
-{
-	final SeedFactory seedFactory;
-	final Collection<ProductCategory> productCategories;
-	final PurchasingProfile purchasingProfile;
-	
-	public TransactionPurchasesSamplerBuilder(Collection<ProductCategory> productCategories,
-			PurchasingProfile purchasingProfile,
-			SeedFactory seedFactory)
-	{
-		this.seedFactory = seedFactory;
-		this.productCategories = productCategories;
-		this.purchasingProfile = purchasingProfile;
-	}
-	
-	public Sampler<Purchase> build() throws Exception
-	{
-		CustomerTransactionParametersSamplerBuilder builder = new CustomerTransactionParametersSamplerBuilder(seedFactory);
-		CustomerTransactionParameters transactionParameters = builder.build().sample();
-		
-		CustomerInventoryBuilder inventoryBuilder = new CustomerInventoryBuilder(transactionParameters,
-				seedFactory);
-		inventoryBuilder.addAllProductCategories(productCategories);
-		CustomerInventory customerInventory = inventoryBuilder.build();
-		
-		PurchasingProcessesBuilder processesBuilder = new PurchasingProcessesBuilder(seedFactory);
-		processesBuilder.setPurchasingProfile(purchasingProfile);
-		PurchasingProcesses processes = processesBuilder.build();
-		
[…]

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/datamodels/simulation/Purchase.java

@@ -1,26 +1,0 @@
-package com.github.rnowling.bps.datagenerator.datamodels.simulation;
-
-import java.util.List;
-
-public class Purchase
-{
-	private final double purchaseTime;
-	private final List<Product> purchasedProducts;
-	
-	public Purchase(double purchaseTime, List<Product> purchasedProducts)
-	{
-		this.purchaseTime = purchaseTime;
-		this.purchasedProducts = purchasedProducts;
-	}
-	
-	public double getPurchaseTime()
-	{
-		return purchaseTime;
-	}
-
-	public List<Product> getPurchasedProducts()
-	{
-		return purchasedProducts;
-	}
-
-}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/datamodels/simulation/PurchasingProfile.java

@@ -1,27 +1,0 @@
-package com.github.rnowling.bps.datagenerator.datamodels.simulation;
-
-import java.util.Map;
-
-import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
-import com.google.common.collect.ImmutableMap;
-import com.google.common.collect.ImmutableSet;
-
-public class PurchasingProfile
-{
-	ImmutableMap<String, MarkovModel<Product>> productCategoryProfiles;
-	
-	public PurchasingProfile(Map<String, MarkovModel<Product>> productCategoryProfiles)
-	{
-		this.productCategoryProfiles = ImmutableMap.copyOf(productCategoryProfiles);
-	}
-	
-	public ImmutableSet<String> getProductCategories()
-	{
-		return productCategoryProfiles.keySet();
-	}
-	
-	public MarkovModel<Product> getProfile(String productCategory)
-	{
-		return productCategoryProfiles.get(productCategory);
-	}
-}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/framework/SeedFactory.java

@@ -1,0 +1,23 @@
+package com.github.rnowling.bps.datagenerator.framework;
+
+import java.util.Random;
+
+public class SeedFactory
+{
+	Random rng;
+	
+	public SeedFactory()
+	{
+		rng = new Random();
+	}
+	
+	public SeedFactory(long seed)
+	{
+		rng = new Random(seed);
+	}
+	
+	public long getNextSeed()
+	{
+		return rng.nextLong();
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/framework/markovmodels/MarkovModel.java

@@ -1,0 +1,29 @@
+package com.github.rnowling.bps.datagenerator.framework.markovmodels;
+
+import java.util.Map;
+
+import com.google.common.collect.ImmutableMap;
+import com.google.common.collect.ImmutableTable;
+import com.google.common.collect.Table;
+
+public class MarkovModel<T>
+{
+	final ImmutableTable<T, T, Double> transitionWeights;
+	final ImmutableMap<T, Double> startWeights;
+	
+	public MarkovModel(Table<T, T, Double> transitionWeights, Map<T, Double> startWeights)
+	{
+		this.transitionWeights = ImmutableTable.copyOf(transitionWeights);
+		this.startWeights = ImmutableMap.copyOf(startWeights);
+	}
+
+	public ImmutableTable<T, T, Double> getTransitionWeights()
+	{
+		return transitionWeights;
+	}
+
+	public ImmutableMap<T, Double> getStartWeights()
+	{
+		return startWeights;
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/framework/markovmodels/MarkovProcess.java

@@ -1,0 +1,53 @@
+package com.github.rnowling.bps.datagenerator.framework.markovmodels;
+
+import java.util.Map;
+
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.samplers.RouletteWheelSampler;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.google.common.collect.ImmutableMap;
+import com.google.common.collect.ImmutableTable;
+
+public class MarkovProcess<T> implements Sampler<T>
+{
+	final ImmutableMap<T, Sampler<T>> transitionSamplers;
+	final Sampler<T> startStateSampler;
+	
+	T currentState;
+	
+	
+	public MarkovProcess(MarkovModel<T> model, SeedFactory factory)
+	{
+		ImmutableTable<T, T, Double> transitionTable = model.getTransitionWeights();
+		
+		startStateSampler = RouletteWheelSampler.create(model.getStartWeights(), factory);
+		
+		ImmutableMap.Builder<T, Sampler<T>> builder = ImmutableMap.builder();
+		for(Map.Entry<T, Map<T, Double>> entry : transitionTable.rowMap().entrySet())
+		{
+			builder.put(entry.getKey(), RouletteWheelSampler.create(entry.getValue(), factory));
+		}
+		
+		
+		this.transitionSamplers = builder.build();
+		
+		currentState = null;
+	}
+	
+	public static <T> MarkovProcess<T> create(MarkovModel<T> model, SeedFactory factory)
+	{
+		return new MarkovProcess<T>(model, factory);
+	}
+	
+	public T sample() throws Exception
+	{
+		if(currentState == null)
+		{
+			currentState = startStateSampler.sample();
+			return currentState;
+		}
+		
+		currentState = transitionSamplers.get(currentState).sample();
+		return currentState;
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/framework/samplers/BoundedMultiModalGaussianSampler.java

@@ -1,0 +1,42 @@
+package com.github.rnowling.bps.datagenerator.framework.samplers;
+
+import java.util.List;
+import java.util.Random;
+
+import com.github.rnowling.bps.datagenerator.datamodels.Pair;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.google.common.collect.ImmutableList;
+
+public class BoundedMultiModalGaussianSampler implements Sampler<Double>
+{
+	ImmutableList<Pair<Double, Double>> distributions;
+	
+	double min;
+	double max;
+	Random rng;
+	
+	public BoundedMultiModalGaussianSampler(List<Pair<Double, Double>> distributions, double min, double max, SeedFactory seedFactory)
+	{
+		rng = new Random(seedFactory.getNextSeed());
+		this.distributions = ImmutableList.copyOf(distributions);
+		
+		this.min = min;
+		this.max = max;
+	}
+	
+	public Double sample()
+	{
+		int idx = rng.nextInt(distributions.size());
+		
+		double mean = distributions.get(idx).getFirst();
+		double var = distributions.get(idx).getSecond();
+		
+		double value = mean + rng.nextGaussian() * Math.sqrt(var);
+		
+		value = Math.min(value, this.max);
+		value = Math.max(value, this.min);
+		
+		return value;
+	}
+
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/framework/samplers/GaussianSampler.java

@@ -1,0 +1,24 @@
+package com.github.rnowling.bps.datagenerator.framework.samplers;
+
+import java.util.Random;
+
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+
+public class GaussianSampler implements Sampler<Double>
+{
+	double mean;
+	double var;
+	Random rng;
+	
+	public GaussianSampler(double mean, double var, SeedFactory seedFactory)
+	{
+		rng = new Random(seedFactory.getNextSeed());
+		this.mean = mean;
+		this.var = var;
+	}
+	
+	public Double sample()
+	{
+		return rng.nextGaussian() * Math.sqrt(var) + mean;
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/framework/samplers/SequenceSampler.java

@@ -1,0 +1,55 @@
+package com.github.rnowling.bps.datagenerator.framework.samplers;
+
+public class SequenceSampler implements Sampler<Integer>
+{
+	Integer start;
+	Integer end;
+	Integer step;
+	Integer next;
+	
+	public SequenceSampler()
+	{
+		start = 0;
+		end = null;
+		step = 1;
+		next = start;
+	}
+	
+	public SequenceSampler(Integer start)
+	{
+		this.start = start;
+		end = null;
+		step = 1;
+		next = start;
+	}
+	
+	public SequenceSampler(Integer start, Integer end)
+	{
+		this.start = start;
+		this.end = end;
+		step = 1;
+		next = start;
+	}
+	
+	public SequenceSampler(Integer start, Integer end, Integer step)
+	{
+		this.start = start;
+		this.end = end;
+		this.step = step;
+		next = start;
+	}
+	
+	public Integer sample() throws Exception
+	{
+		if(end == null || next < end)
+		{
+			Integer current = next;
+			next = current + step;
+			return current;
+		}
+		
+		throw new Exception("All values have been sampled");
+	}
+	
+	
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/framework/samplers/StatefulRejectSamplingMonteCarloSampler.java

@@ -1,0 +1,45 @@
+package com.github.rnowling.bps.datagenerator.framework.samplers;
+
+import java.util.Random;
+
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.pdfs.ConditionalProbabilityDensityFunction;
+
+
+public class StatefulRejectSamplingMonteCarloSampler<T> implements Sampler<T>
+{
+	private final Sampler<T> stateSampler;
+	private final Random rng;
+	private final ConditionalProbabilityDensityFunction<T, T> acceptancePDF;
+	private T currentState;
+	
+	public StatefulRejectSamplingMonteCarloSampler(Sampler<T> stateGenerator,
+			ConditionalProbabilityDensityFunction<T, T> acceptancePDF,
+			T initialState,
+			SeedFactory seedFactory)
+	{
+		this.acceptancePDF = acceptancePDF;
+		this.stateSampler = stateGenerator;
+		
+		rng = new Random(seedFactory.getNextSeed());
+		
+		this.currentState = initialState;
+	}
+
+	public T sample() throws Exception
+	{
+		while(true)
+		{
+			T proposedState = this.stateSampler.sample();
+			double probability = acceptancePDF.probability(proposedState, currentState);
+			double r = rng.nextDouble();
+			
+			if(r < probability)
+			{
+				this.currentState = proposedState;
+				return proposedState;
+			}
+		}
+	}
+	
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/generators/purchasingprofile/ProductCategoryMarkovModelSampler.java

@@ -1,0 +1,126 @@
+package com.github.rnowling.bps.datagenerator.generators.purchasingprofile;
+
+import java.util.Map;
+
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
+import com.github.rnowling.bps.datagenerator.framework.markovmodels.MarkovModel;
+import com.github.rnowling.bps.datagenerator.framework.markovmodels.MarkovModelBuilder;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.google.common.collect.Maps;
+
+public class ProductCategoryMarkovModelSampler implements Sampler<MarkovModel<Product>>
+{
+	final ProductCategory productCategory;
+	final Sampler<Double> fieldWeightSampler;
+	final Sampler<Double> fieldSimilarityWeightSampler;
+	final Sampler<Double> loopbackWeightSampler;
+	
+	Map<String, Double> fieldWeights;
+	Map<String, Double> fieldSimilarityWeights;
+	double loopbackWeight;
+	
+	public ProductCategoryMarkovModelSampler(ProductCategory productCategory, 
+			Sampler<Double> fieldWeightSampler, Sampler<Double> fieldSimilarityWeightSampler,
+			Sampler<Double> loopbackWeightSampler)
+	{
+		this.productCategory = productCategory;
+		
+		this.fieldSimilarityWeightSampler = fieldSimilarityWeightSampler;
+		this.fieldWeightSampler = fieldWeightSampler;
+		this.loopbackWeightSampler = loopbackWeightSampler;
+	}
+	
+	protected void generateWeights() throws Exception
+	{
+		fieldWeights = Maps.newHashMap();
+		fieldSimilarityWeights = Maps.newHashMap();
+		
+		for(String fieldName : productCategory.getFieldNames())
+		{
+			fieldWeights.put(fieldName, fieldWeightSampler.sample());
+			fieldSimilarityWeights.put(fieldName,fieldSimilarityWeightSampler.sample());
+		}
+		
+		loopbackWeight = loopbackWeightSampler.sample();
+	}
+	
+	protected <T> Map<T, Double> normalize(Map<T, Double> weights)
+	{
+		Map<T, Double> normalized = Maps.newHashMap();
+		
+		double weightSum = 0.0;
+		for(double weight : weights.values())
+		{
+			weightSum += weight;
+		}
+		
[…]

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/generators/purchasingprofile/PurchasingProfileBuilder.java

@@ -1,0 +1,27 @@
+package com.github.rnowling.bps.datagenerator.generators.purchasingprofile;
+
+import java.util.Map;
+
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
+import com.github.rnowling.bps.datagenerator.framework.markovmodels.MarkovModel;
+import com.google.common.collect.Maps;
+
+public class PurchasingProfileBuilder
+{
+	Map<String, MarkovModel<Product>> profiles;
+	
+	public PurchasingProfileBuilder()
+	{
+		profiles = Maps.newHashMap();
+	}
+	
+	public void addProfile(String productCategory, MarkovModel<Product> profile)
+	{
+		profiles.put(productCategory, profile);
+	}
+	
+	public PurchasingProfile build()
+	{
+		return new PurchasingProfile(profiles);
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/generators/purchasingprofile/PurchasingProfileSamplerBuilder.java

@@ -1,0 +1,57 @@
+package com.github.rnowling.bps.datagenerator.generators.purchasingprofile;
+
+import java.util.Collection;
+import java.util.Map;
+
+import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.markovmodels.MarkovModel;
+import com.github.rnowling.bps.datagenerator.framework.samplers.BoundedMultiModalGaussianSampler;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.google.common.collect.ImmutableList;
+import com.google.common.collect.Maps;
+
+public class PurchasingProfileSamplerBuilder
+{
+	final ImmutableList<ProductCategory> productCategories;
+	final SeedFactory seedFactory;
+	final Sampler<Double> fieldWeightSampler;
+	final Sampler<Double> fieldSimilarityWeightSampler;
+	final Sampler<Double> loopbackWeightSampler;
+	
+	public PurchasingProfileSamplerBuilder(Collection<ProductCategory> productCategories, SeedFactory seedFactory)
+	{
+		this.productCategories = ImmutableList.copyOf(productCategories);
+		this.seedFactory = seedFactory;
+		
+		this.fieldWeightSampler = new BoundedMultiModalGaussianSampler(Constants.PRODUCT_MSM_FIELD_WEIGHT_GAUSSIANS, 
+				Constants.PRODUCT_MSM_FIELD_WEIGHT_LOWERBOUND, 
+				Constants.PRODUCT_MSM_FIELD_WEIGHT_UPPERBOUND,
+				seedFactory);
+	
+	this.fieldSimilarityWeightSampler = new BoundedMultiModalGaussianSampler(Constants.PRODUCT_MSM_FIELD_SIMILARITY_WEIGHT_GAUSSIANS,
+			Constants.PRODUCT_MSM_FIELD_SIMILARITY_WEIGHT_LOWERBOUND, 
+			Constants.PRODUCT_MSM_FIELD_SIMILARITY_WEIGHT_UPPERBOUND,
+			seedFactory);
+	
[…]

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/generators/transaction/CustomerInventory.java

@@ -1,0 +1,49 @@
+package com.github.rnowling.bps.datagenerator.generators.transaction;
+
+import java.util.Map;
+
+import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
+import com.google.common.collect.ImmutableMap;
+import com.google.common.collect.Maps;
+
+public class CustomerInventory
+{
+	final private ImmutableMap<String, ProductCategoryInventory> productCategoryInventories;
+	
+	public CustomerInventory(Map<String, ProductCategoryInventory> productCategoryInventories)
+	{
+		this.productCategoryInventories = ImmutableMap.copyOf(productCategoryInventories);
+	}
+	
+	public void simulatePurchase(double time, Product product) throws Exception
+	{
+		String category = product.getFieldValueAsString(Constants.PRODUCT_CATEGORY);
+		ProductCategoryInventory inventory = productCategoryInventories.get(category);
+		inventory.simulatePurchase(time, product);
+	}
+	
+	public ImmutableMap<String, Double> getInventoryAmounts(double time)
+	{
+		Map<String, Double> amounts = Maps.newHashMap();
+		for(String category : productCategoryInventories.keySet())
+		{
+			double amount = productCategoryInventories.get(category).findRemainingAmount(time);
+			amounts.put(category, amount);
+		}
+		
+		return ImmutableMap.copyOf(amounts);
+	}
+	
+	public ImmutableMap<String, Double> getExhaustionTimes()
+	{
+		Map<String, Double> times = Maps.newHashMap();
+		for(String category : productCategoryInventories.keySet())
+		{
+			double time = productCategoryInventories.get(category).findExhaustionTime();
+			times.put(category, time);
+		}
+		
+		return ImmutableMap.copyOf(times);
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/generators/transaction/CustomerInventoryBuilder.java

@@ -1,0 +1,53 @@
+package com.github.rnowling.bps.datagenerator.generators.transaction;
+
+import java.util.Collection;
+import java.util.List;
+import java.util.Map;
+
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.google.common.collect.Lists;
+import com.google.common.collect.Maps;
+
+public class CustomerInventoryBuilder
+{
+	final private List<ProductCategory> productCategories;
+	final private SeedFactory seedFactory;
+	final private CustomerTransactionParameters parameters;
+	
+	public CustomerInventoryBuilder(CustomerTransactionParameters parameters,
+			SeedFactory seedFactory)
+	{
+		productCategories = Lists.newArrayList();
+		this.seedFactory = seedFactory;
+		this.parameters = parameters;
+	}
+	
+	public void addProductCategory(ProductCategory productCategory)
+	{
+		this.productCategories.add(productCategory);
+	}
+	
+	public void addAllProductCategories(Collection<ProductCategory> productCategories)
+	{
+		this.productCategories.addAll(productCategories);
+	}
+	
+	public CustomerInventory build()
+	{
+		Map<String, ProductCategoryInventory> inventories = Maps.newHashMap();
+		for(ProductCategory productCategory : productCategories)
+		{
+			if(parameters.countPetsBySpecies(productCategory.getApplicableSpecies()) > 0)
+			{
+				ProductCategoryInventory inventory = new ProductCategoryInventory(productCategory,
+					parameters, seedFactory);
+				inventories.put(productCategory.getCategoryLabel(), inventory);
+			}
+		}
+		
+		return new CustomerInventory(inventories);
+	}
+}
+
+

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/generators/transaction/PurchasingProcesses.java

@@ -1,0 +1,22 @@
+package com.github.rnowling.bps.datagenerator.generators.transaction;
+
+import java.util.Map;
+
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.google.common.collect.ImmutableMap;
+
+public class PurchasingProcesses
+{
+	ImmutableMap<String, Sampler<Product>> processes;
+	
+	public PurchasingProcesses(Map<String, Sampler<Product>> processes)
+	{
+		this.processes = ImmutableMap.copyOf(processes);
+	}
+	
+	public Product simulatePurchase(String productCategory) throws Exception
+	{
+		return this.processes.get(productCategory).sample();
+	}
+}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/pdfs/store/StoreLocationPopulationPDF.java

@@ -1,28 +1,0 @@
-package com.github.rnowling.bps.datagenerator.pdfs.store;
-
-import java.util.List;
-
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ZipcodeRecord;
-import com.github.rnowling.bps.datagenerator.statistics.pdfs.ProbabilityDensityFunction;
-
-public class StoreLocationPopulationPDF implements ProbabilityDensityFunction<ZipcodeRecord>
-{
-	double populationSum = 0.0;
-	
-	public StoreLocationPopulationPDF(List<ZipcodeRecord> zipcodeTable)
-	{
-		long populationSum = 0L;
-		for(ZipcodeRecord record : zipcodeTable)
-		{
-			populationSum += record.getPopulation();
-		}
-		
-		this.populationSum = ((double) populationSum);
-	}
-	
-	public double probability(ZipcodeRecord record)
-	{
-		return ((double) record.getPopulation()) / populationSum;
-	}
-
-}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/samplers/transaction/CustomerTransactionParametersSampler.java

@@ -1,48 +1,0 @@
-package com.github.rnowling.bps.datagenerator.samplers.transaction;
-
-import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerTransactionParametersBuilder;
-import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-
-public class CustomerTransactionParametersSampler implements Sampler<CustomerTransactionParameters>
-{
-	final private Sampler<Integer> nPetsSampler;
-	final private Sampler<PetSpecies> petSpeciesSampler;
-	final private Sampler<Double> purchaseTriggerTimeSampler;
-	final private Sampler<Double> transactionTriggerTimeSampler;
-
-	public CustomerTransactionParametersSampler(Sampler<Integer> nPetsSampler,
-			Sampler<PetSpecies> petSpeciesSampler, 
-			Sampler<Double> purchaseTriggerTimeSampler,
-			Sampler<Double> transactionTriggerTimeSampler)
-	{
-
-		this.nPetsSampler = nPetsSampler;
-		this.petSpeciesSampler = petSpeciesSampler;
-		this.purchaseTriggerTimeSampler = purchaseTriggerTimeSampler;
-		this.transactionTriggerTimeSampler = transactionTriggerTimeSampler;
-	}
-	
-	protected void generatePets(CustomerTransactionParametersBuilder builder) throws Exception
-	{
-		int nPets = this.nPetsSampler.sample();
-		
-		for(int i = 0; i < nPets; i++)
-		{
-			PetSpecies species = this.petSpeciesSampler.sample();
-			builder.addPet(species);	
-		}
-	}
-	
-	public CustomerTransactionParameters sample() throws Exception
-	{
-		CustomerTransactionParametersBuilder builder = new CustomerTransactionParametersBuilder();
-		
-		this.generatePets(builder);
-		builder.setAveragePurchaseTriggerTime(this.purchaseTriggerTimeSampler.sample());
-		builder.setAverageTransactionTriggerTime(this.transactionTriggerTimeSampler.sample());
-		
-		return builder.build();
-	}
-}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/samplers/transaction/TransactionSampler.java

@@ -1,38 +1,0 @@
-package com.github.rnowling.bps.datagenerator.samplers.transaction;
-
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Customer;
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Transaction;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.Purchase;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-
-public class TransactionSampler implements Sampler<Transaction>
-{
-	private final Sampler<Purchase> purchasesSampler;
-	private final Sampler<Store> storeSampler;
-	private final Sampler<Integer> idSampler;
-	private final Customer customer;
-	
-	public TransactionSampler(Customer customer, Sampler<Purchase> purchasesSampler,
-			Sampler<Store> storeSampler, Sampler<Integer> idSampler)
-	{
-		this.customer = customer;
-		this.purchasesSampler = purchasesSampler;
-		this.storeSampler = storeSampler;
-		this.idSampler = idSampler;
-	}
-	
-	
-	public Transaction sample() throws Exception
-	{		
-		Purchase purchase = purchasesSampler.sample();
-		Store store = this.storeSampler.sample();
-		Integer id = idSampler.sample();
-			
-		Transaction transaction = new Transaction(id, customer, store, purchase.getPurchaseTime(),
-					purchase.getPurchasedProducts());
-		
-		return transaction;
-	}
-
-}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/statistics/pdfs/ProbabilityDensityFunction.java

@@ -1,6 +1,0 @@
-package com.github.rnowling.bps.datagenerator.statistics.pdfs;
-
-public interface ProbabilityDensityFunction<T>
-{
-	public double probability(T datum);
-}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/statistics/samplers/DoubleSequenceSampler.java

@@ -1,55 +1,0 @@
-package com.github.rnowling.bps.datagenerator.statistics.samplers;
-
-public class DoubleSequenceSampler implements Sampler<Double>
-{
-	Double start;
-	Double end;
-	Double step;
-	Double next;
-	
-	public DoubleSequenceSampler()
-	{
-		start = 0.0;
-		end = null;
-		step = 1.0;
-		next = start;
-	}
-	
-	public DoubleSequenceSampler(Double start)
-	{
-		this.start = start;
-		end = null;
-		step = 1.0;
-		next = start;
-	}
-	
-	public DoubleSequenceSampler(Double start, Double end)
-	{
-		this.start = start;
-		this.end = end;
-		step = 1.0;
-		next = start;
-	}
-	
-	public DoubleSequenceSampler(Double start, Double end, Double step)
-	{
-		this.start = start;
-		this.end = end;
-		this.step = step;
-		next = start;
-	}
-	
-	public Double sample() throws Exception
-	{
-		if(end == null || next < end)
-		{
-			Double current = next;
-			next = current + step;
-			return current;
-		}
-		
-		throw new Exception("All values have been sampled");
-	}
-	
-	
-}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/statistics/samplers/GaussianSampler.java

@@ -1,24 +1,0 @@
-package com.github.rnowling.bps.datagenerator.statistics.samplers;
-
-import java.util.Random;
-
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-
-public class GaussianSampler implements Sampler<Double>
-{
-	double mean;
-	double var;
-	Random rng;
-	
-	public GaussianSampler(double mean, double var, SeedFactory seedFactory)
-	{
-		rng = new Random(seedFactory.getNextSeed());
-		this.mean = mean;
-		this.var = var;
-	}
-	
-	public Double sample()
-	{
-		return rng.nextGaussian() * Math.sqrt(var) + mean;
-	}
-}

### 2014-10-12 src/java/bps-data-generator/src/main/java/com/github/rnowling/bps/datagenerator/statistics/samplers/Sampler.java

@@ -1,6 +1,0 @@
-package com.github.rnowling.bps.datagenerator.statistics.samplers;
-
-public interface Sampler<T>
-{
-	public T sample() throws Exception;
-}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/builders/customer/TestCustomerSamplerBuilder.java

@@ -1,55 +1,0 @@
-package com.github.rnowling.bps.datagenerator.builders.customer;
-
-import static org.junit.Assert.assertNotNull;
-import static org.junit.Assert.assertTrue;
-
-import java.util.List;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.Names;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ZipcodeRecord;
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Customer;
-import com.github.rnowling.bps.datagenerator.datareaders.NameReader;
-import com.github.rnowling.bps.datagenerator.datareaders.ZipcodeReader;
-import com.github.rnowling.bps.datagenerator.samplers.customer.CustomerSampler;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.RouletteWheelSampler;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.SequenceSampler;
-
-public class TestCustomerSamplerBuilder
-{
-
-	@Test
-	public void testSample() throws Exception
-	{
-		ZipcodeReader zipcodeReader = new ZipcodeReader();
-		zipcodeReader.setCoordinatesFile(Constants.COORDINATES_FILE);
-		zipcodeReader.setIncomesFile(Constants.INCOMES_FILE);
-		zipcodeReader.setPopulationFile(Constants.POPULATION_FILE);
-		List<ZipcodeRecord> zipcodes = zipcodeReader.readData();
-		
-		NameReader nameReader = new NameReader(Constants.NAMEDB_FILE);
-		Names names = nameReader.readData();
-		
-		SeedFactory factory = new SeedFactory(1234);
-		
-		Sampler<Customer> sampler = new CustomerSampler(new SequenceSampler(),
-				RouletteWheelSampler.createUniform(names.getFirstNames().keySet(), factory), 
-				RouletteWheelSampler.createUniform(names.getLastNames().keySet(), factory), 
-				RouletteWheelSampler.createUniform(zipcodes, factory));
-		
-		Customer customer = sampler.sample();
-		
-		assertNotNull(customer);
-		assertTrue(customer.getId() >= 0);
[…]

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/builders/transaction/TestCustomerInventoryBuilder.java

@@ -1,75 +1,0 @@
-package com.github.rnowling.bps.datagenerator.builders.transaction;
-
-import static org.junit.Assert.assertEquals;
-import static org.junit.Assert.assertTrue;
-
-import java.util.Map;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerTransactionParametersSamplerBuilder;
-import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerInventory;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
-import com.github.rnowling.bps.datagenerator.datareaders.ProductCategoryBuilder;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-import com.google.common.collect.Maps;
-
-public class TestCustomerInventoryBuilder
-{
-	
-	@Test
-	public void testBuild() throws Exception
-	{
-		SeedFactory seedFactory = new SeedFactory(1234);
-		
-		CustomerTransactionParametersSamplerBuilder transParamsBuilder = new CustomerTransactionParametersSamplerBuilder(seedFactory);
-		Sampler<CustomerTransactionParameters> sampler = transParamsBuilder.build();
-		
-		CustomerTransactionParameters parameters = sampler.sample();
-		
-		ProductCategoryBuilder builder = new ProductCategoryBuilder();
-		builder.addApplicableSpecies(PetSpecies.DOG);
-		builder.setAmountUsedPetPetAverage(1.0);
-		builder.setAmountUsedPetPetVariance(1.0);
-		builder.setDailyUsageRate(2.0);
-		builder.setCategory("dog food");
-		
-		ProductCategory category = builder.build();
-		
-		CustomerInventoryBuilder inventoryBuilder = new CustomerInventoryBuilder(parameters, seedFactory);
-		inventoryBuilder.addProductCategory(category);
-		
[…]

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/builders/transaction/TestCustomerTransactionParametersBuilder.java

@@ -1,34 +1,0 @@
-package com.github.rnowling.bps.datagenerator.builders.transaction;
-
-import static org.junit.Assert.assertTrue;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerTransactionParametersBuilder;
-import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
-
-public class TestCustomerTransactionParametersBuilder
-{
-
-	@Test
-	public void testBuild() throws Exception
-	{
-		CustomerTransactionParametersBuilder builder = new CustomerTransactionParametersBuilder();
-		
-		builder.addPet(PetSpecies.DOG);
-		builder.addPet(PetSpecies.DOG);
-		builder.addPet(PetSpecies.DOG);
-		
-		builder.setAveragePurchaseTriggerTime(1.0);
-		builder.setAverageTransactionTriggerTime(2.0);
-		
-		CustomerTransactionParameters transParams = builder.build();
-		
-		assertTrue(transParams.countPetsBySpecies(PetSpecies.DOG) == 3);
-		assertTrue(transParams.countPetsBySpecies(PetSpecies.CAT) == 0);
-		assertTrue(transParams.getAveragePurchaseTriggerTime() == 1.0);
-		assertTrue(transParams.getAverageTransactionTriggerTime() == 2.0);
-	}
-
-}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/datamodels/simulation/TestCustomerTransactionParameters.java

@@ -1,54 +1,0 @@
-package com.github.rnowling.bps.datagenerator.datamodels.simulation;
-
-import static org.junit.Assert.assertEquals;
-
-import java.util.Arrays;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
-import com.google.common.collect.HashMultiset;
-import com.google.common.collect.Multiset;
-
-public class TestCustomerTransactionParameters
-{
-
-	@Test
-	public void testCountPetsBySpecies() throws Exception
-	{
-		Multiset<PetSpecies> petCounts = HashMultiset.create();
-		
-		petCounts.add(PetSpecies.CAT);
-		petCounts.add(PetSpecies.CAT);
-		petCounts.add(PetSpecies.CAT);
-		
-		
-		CustomerTransactionParameters transParams = new CustomerTransactionParameters(
-				petCounts, 0.0, 0.0);
-		
-		
-		assertEquals(transParams.countPetsBySpecies(PetSpecies.CAT), 3);
-		assertEquals(transParams.countPetsBySpecies(PetSpecies.DOG), 0);
-		assertEquals(transParams.countPets(), 3);
-	}
-	
-	@Test
-	public void testCountPetsByMultipleSpecies() throws Exception
-	{
-		Multiset<PetSpecies> petCounts = HashMultiset.create();
-		
-		petCounts.add(PetSpecies.CAT);
-		petCounts.add(PetSpecies.CAT);
-		petCounts.add(PetSpecies.DOG);
-		
-		
-		CustomerTransactionParameters transParams = new CustomerTransactionParameters(
-				petCounts, 0.0, 0.0);
-		
-		
-		assertEquals(transParams.countPetsBySpecies(Arrays.asList(PetSpecies.values())), 3);
-		assertEquals(transParams.countPets(), 3);
-	}
-
-}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/datamodels/simulation/TestProductCategoryInventory.java

@@ -1,60 +1,0 @@
-package com.github.rnowling.bps.datagenerator.datamodels.simulation;
-
-import static org.junit.Assert.assertEquals;
-import static org.junit.Assert.assertTrue;
-
-import java.util.Map;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.builders.transaction.CustomerTransactionParametersSamplerBuilder;
-import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.CustomerTransactionParameters;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.ProductCategoryInventory;
-import com.github.rnowling.bps.datagenerator.datareaders.ProductCategoryBuilder;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-import com.google.common.collect.Maps;
-
-public class TestProductCategoryInventory
-{
-	
-	@Test
-	public void testPurchase() throws Exception
-	{
-		SeedFactory seedFactory = new SeedFactory(1234);
-		
-		CustomerTransactionParametersSamplerBuilder transParamsBuilder = new CustomerTransactionParametersSamplerBuilder(seedFactory);
-		Sampler<CustomerTransactionParameters> sampler = transParamsBuilder.build();
-		
-		CustomerTransactionParameters parameters = sampler.sample();
-		
-		ProductCategoryBuilder builder = new ProductCategoryBuilder();
-		builder.addApplicableSpecies(PetSpecies.DOG);
-		builder.setAmountUsedPetPetAverage(1.0);
-		builder.setAmountUsedPetPetVariance(1.0);
-		builder.setDailyUsageRate(2.0);
-		
-		
-		ProductCategory category = builder.build();
-		
-		ProductCategoryInventory inventory = new ProductCategoryInventory(category, parameters, seedFactory);
-		
-		assertEquals(inventory.findExhaustionTime(), 0.0, 0.0001);
[…]

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/framework/markovmodels/TestMarkovModelBuilder.java

@@ -1,0 +1,62 @@
+package com.github.rnowling.bps.datagenerator.framework.markovmodels;
+
+import org.junit.Test;
+
+import static org.junit.Assert.assertThat;
+import static org.junit.Assert.assertEquals;
+import static org.junit.matchers.JUnitMatchers.*;
+
+import com.github.rnowling.bps.datagenerator.framework.markovmodels.MarkovModel;
+import com.github.rnowling.bps.datagenerator.framework.markovmodels.MarkovModelBuilder;
+
+public class TestMarkovModelBuilder
+{
+
+	@Test
+	public void testAddStateState()
+	{
+		MarkovModelBuilder<String> builder = MarkovModelBuilder.create();
+		
+		builder.addStartState("a", 1.0);
+		
+		MarkovModel<String> msm = builder.build();
+		
+		assertThat(msm.getStartWeights().keySet(), hasItem("a"));
+		assertEquals((double) msm.getStartWeights().get("a"), (double) 1.0, 0.000001);
+		
+	}
+	
+	@Test
+	public void testAddEdgeTransition()
+	{
+		MarkovModelBuilder<String> builder = MarkovModelBuilder.create();
+		
+		builder.addTransition("a", "b", 1.0);
+		
+		MarkovModel<String> msm = builder.build();
+		
+		assertThat(msm.getTransitionWeights().rowKeySet(), hasItem("a"));
+		assertThat(msm.getTransitionWeights().columnKeySet(), hasItem("b"));
+		assertEquals((double) msm.getTransitionWeights().get("a", "b"), (double) 1.0, 0.000001);	
+	}
+	
+	@Test
+	public void testBuildMSM()
+	{
+		MarkovModelBuilder<String> builder = MarkovModelBuilder.create();
+		
+		builder.addStartState("a", 1.0);
+		builder.addTransition("a", "b", 1.0);
+		builder.addTransition("a", "c", 1.0);
+		
+		MarkovModel<String> msm = builder.build();
+		
+		assertThat(msm.getStartWeights().keySet(), hasItem("a"));
+		assertThat(msm.getTransitionWeights().rowKeySet(), hasItem("a"));
+		assertThat(msm.getTransitionWeights().columnKeySet(), hasItem("b"));
+		assertThat(msm.getTransitionWeights().columnKeySet(), hasItem("c"));
[…]

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/generators/customer/TestCustomerLocationPDF.java

@@ -1,0 +1,51 @@
+package com.github.rnowling.bps.datagenerator.generators.customer;
+
+import static org.junit.Assert.assertTrue;
+
+import java.util.ArrayList;
+import java.util.List;
+
+import org.junit.Test;
+
+import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ZipcodeRecord;
+import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
+import com.github.rnowling.bps.datagenerator.datareaders.ZipcodeReader;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.github.rnowling.bps.datagenerator.generators.customer.CustomerLocationPDF;
+import com.github.rnowling.bps.datagenerator.generators.store.StoreSamplerBuilder;
+
+public class TestCustomerLocationPDF
+{
+
+	@Test
+	public void testProbability() throws Exception
+	{
+		ZipcodeReader zipcodeReader = new ZipcodeReader();
+		zipcodeReader.setCoordinatesFile(Constants.COORDINATES_FILE);
+		zipcodeReader.setIncomesFile(Constants.INCOMES_FILE);
+		zipcodeReader.setPopulationFile(Constants.POPULATION_FILE);
+		List<ZipcodeRecord> zipcodes = zipcodeReader.readData();
+		
+		SeedFactory factory = new SeedFactory(1234);
+		
+		StoreSamplerBuilder storeSamplerBuilder = new StoreSamplerBuilder(zipcodes, factory);
+		Sampler<Store> storeSampler = storeSamplerBuilder.build();
+		
+		List<Store> stores = new ArrayList<Store>();
+		for(int i = 0; i < 10; i++)
+		{
+			Store store = storeSampler.sample();
+			stores.add(store);
+		}
+		
+		CustomerLocationPDF customerLocationPDF = new CustomerLocationPDF(zipcodes, stores, 
+					Constants.AVERAGE_CUSTOMER_STORE_DISTANCE);
+		
+		double prob = customerLocationPDF.probability(zipcodes.get(0));
+		
+		assertTrue(prob > 0.0);
+	}
+
+}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/generators/customer/TestCustomerSamplerBuilder.java

@@ -1,0 +1,55 @@
+package com.github.rnowling.bps.datagenerator.generators.customer;
+
+import static org.junit.Assert.assertNotNull;
+import static org.junit.Assert.assertTrue;
+
+import java.util.List;
+
+import org.junit.Test;
+
+import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.Names;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ZipcodeRecord;
+import com.github.rnowling.bps.datagenerator.datamodels.outputs.Customer;
+import com.github.rnowling.bps.datagenerator.datareaders.NameReader;
+import com.github.rnowling.bps.datagenerator.datareaders.ZipcodeReader;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.samplers.RouletteWheelSampler;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.github.rnowling.bps.datagenerator.framework.samplers.SequenceSampler;
+import com.github.rnowling.bps.datagenerator.generators.customer.CustomerSampler;
+
+public class TestCustomerSamplerBuilder
+{
+
+	@Test
+	public void testSample() throws Exception
+	{
+		ZipcodeReader zipcodeReader = new ZipcodeReader();
+		zipcodeReader.setCoordinatesFile(Constants.COORDINATES_FILE);
+		zipcodeReader.setIncomesFile(Constants.INCOMES_FILE);
+		zipcodeReader.setPopulationFile(Constants.POPULATION_FILE);
+		List<ZipcodeRecord> zipcodes = zipcodeReader.readData();
+		
+		NameReader nameReader = new NameReader(Constants.NAMEDB_FILE);
+		Names names = nameReader.readData();
+		
+		SeedFactory factory = new SeedFactory(1234);
+		
+		Sampler<Customer> sampler = new CustomerSampler(new SequenceSampler(),
+				RouletteWheelSampler.createUniform(names.getFirstNames().keySet(), factory), 
+				RouletteWheelSampler.createUniform(names.getLastNames().keySet(), factory), 
+				RouletteWheelSampler.createUniform(zipcodes, factory));
+		
+		Customer customer = sampler.sample();
+		
+		assertNotNull(customer);
+		assertTrue(customer.getId() >= 0);
[…]

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/generators/purchasingprofile/TestPurchasingProfileSamplerBuilder.java

@@ -1,0 +1,47 @@
+package com.github.rnowling.bps.datagenerator.generators.purchasingprofile;
+
+import static org.junit.Assert.assertNotNull;
+import static org.junit.Assert.assertTrue;
+
+import java.util.List;
+
+import org.junit.Test;
+
+import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
+import com.github.rnowling.bps.datagenerator.datareaders.ProductsReader;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.github.rnowling.bps.datagenerator.generators.purchasingprofile.PurchasingProfileSamplerBuilder;
+
+public class TestPurchasingProfileSamplerBuilder
+{
+
+	@Test
+	public void testBuild() throws Exception
+	{
+		SeedFactory seedFactory = new SeedFactory(1245);
+		
+		ProductsReader reader = new ProductsReader(Constants.PRODUCTS_FILE);
+		
+		List<ProductCategory> productCategories = reader.readData();
+		
+		PurchasingProfileSamplerBuilder builder = new PurchasingProfileSamplerBuilder(productCategories, seedFactory);
+		Sampler<PurchasingProfile> sampler = builder.build();
+		PurchasingProfile profile = sampler.sample();
+		
+		assertNotNull(profile);
+		assertNotNull(profile.getProductCategories());
+		assertTrue(profile.getProductCategories().size() > 0);
+		
+		for(String label : profile.getProductCategories())
+		{
+			assertNotNull(profile.getProfile(label));
+			assertNotNull(profile.getProfile(label).getStartWeights());
+			assertTrue(profile.getProfile(label).getStartWeights().size() > 0);
+			assertNotNull(profile.getProfile(label).getTransitionWeights());
+			assertTrue(profile.getProfile(label).getTransitionWeights().size() > 0);
+		}
+	}
+
+}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/generators/transaction/TestCustomerInventory.java

@@ -1,0 +1,79 @@
+package com.github.rnowling.bps.datagenerator.generators.transaction;
+
+import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertTrue;
+
+import java.util.Map;
+
+import org.junit.Test;
+
+import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
+import com.github.rnowling.bps.datagenerator.datareaders.ProductCategoryBuilder;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.github.rnowling.bps.datagenerator.generators.transaction.CustomerInventory;
+import com.github.rnowling.bps.datagenerator.generators.transaction.CustomerTransactionParameters;
+import com.github.rnowling.bps.datagenerator.generators.transaction.CustomerTransactionParametersSamplerBuilder;
+import com.github.rnowling.bps.datagenerator.generators.transaction.ProductCategoryInventory;
+import com.google.common.collect.Maps;
+
+public class TestCustomerInventory
+{
+	
+	@Test
+	public void testPurchase() throws Exception
+	{
+		SeedFactory seedFactory = new SeedFactory(1234);
+		
+		CustomerTransactionParametersSamplerBuilder transParamsBuilder = new CustomerTransactionParametersSamplerBuilder(seedFactory);
+		Sampler<CustomerTransactionParameters> sampler = transParamsBuilder.build();
+		
+		CustomerTransactionParameters parameters = sampler.sample();
+		
+		ProductCategoryBuilder builder = new ProductCategoryBuilder();
+		builder.addApplicableSpecies(PetSpecies.DOG);
+		builder.setAmountUsedPetPetAverage(1.0);
+		builder.setAmountUsedPetPetVariance(1.0);
+		builder.setDailyUsageRate(2.0);
+		builder.setCategory("dog food");
+		
+		
+		ProductCategory category = builder.build();
+		
+		ProductCategoryInventory productInventory = new ProductCategoryInventory(category, parameters, seedFactory);
+		
[…]

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/generators/transaction/TestCustomerInventoryBuilder.java

@@ -1,0 +1,74 @@
+package com.github.rnowling.bps.datagenerator.generators.transaction;
+
+import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertTrue;
+
+import java.util.Map;
+
+import org.junit.Test;
+
+import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
+import com.github.rnowling.bps.datagenerator.datareaders.ProductCategoryBuilder;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.github.rnowling.bps.datagenerator.generators.transaction.CustomerInventoryBuilder;
+import com.github.rnowling.bps.datagenerator.generators.transaction.CustomerTransactionParametersSamplerBuilder;
+import com.google.common.collect.Maps;
+
+public class TestCustomerInventoryBuilder
+{
+	
+	@Test
+	public void testBuild() throws Exception
+	{
+		SeedFactory seedFactory = new SeedFactory(1234);
+		
+		CustomerTransactionParametersSamplerBuilder transParamsBuilder = new CustomerTransactionParametersSamplerBuilder(seedFactory);
+		Sampler<CustomerTransactionParameters> sampler = transParamsBuilder.build();
+		
+		CustomerTransactionParameters parameters = sampler.sample();
+		
+		ProductCategoryBuilder builder = new ProductCategoryBuilder();
+		builder.addApplicableSpecies(PetSpecies.DOG);
+		builder.setAmountUsedPetPetAverage(1.0);
+		builder.setAmountUsedPetPetVariance(1.0);
+		builder.setDailyUsageRate(2.0);
+		builder.setCategory("dog food");
+		
+		ProductCategory category = builder.build();
+		
+		CustomerInventoryBuilder inventoryBuilder = new CustomerInventoryBuilder(parameters, seedFactory);
+		inventoryBuilder.addProductCategory(category);
+		
+		CustomerInventory inventory = inventoryBuilder.build();
+		
[…]

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/generators/transaction/TestTransactionPurchasesHiddenMarkovModel.java

@@ -1,0 +1,189 @@
+package com.github.rnowling.bps.datagenerator.generators.transaction;
+
+import static org.junit.Assert.assertNotNull;
+import static org.junit.Assert.assertTrue;
+
+import java.util.List;
+import java.util.Map;
+
+import org.junit.Test;
+
+import com.github.rnowling.bps.datagenerator.Constants;
+import com.github.rnowling.bps.datagenerator.datamodels.PetSpecies;
+import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
+import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
+import com.github.rnowling.bps.datagenerator.datareaders.ProductCategoryBuilder;
+import com.github.rnowling.bps.datagenerator.framework.SeedFactory;
+import com.github.rnowling.bps.datagenerator.framework.markovmodels.MarkovModel;
+import com.github.rnowling.bps.datagenerator.framework.markovmodels.MarkovModelBuilder;
+import com.github.rnowling.bps.datagenerator.framework.samplers.DoubleSequenceSampler;
+import com.github.rnowling.bps.datagenerator.framework.samplers.Sampler;
+import com.github.rnowling.bps.datagenerator.generators.purchasingprofile.PurchasingProfile;
+import com.github.rnowling.bps.datagenerator.generators.purchasingprofile.PurchasingProfileBuilder;
+import com.github.rnowling.bps.datagenerator.generators.transaction.CustomerInventoryBuilder;
+import com.github.rnowling.bps.datagenerator.generators.transaction.CustomerTransactionParametersSamplerBuilder;
+import com.github.rnowling.bps.datagenerator.generators.transaction.PurchasingProcesses;
+import com.github.rnowling.bps.datagenerator.generators.transaction.PurchasingProcessesBuilder;
+import com.github.rnowling.bps.datagenerator.generators.transaction.TransactionPurchasesHiddenMarkovModel;
+import com.google.common.collect.Lists;
+import com.google.common.collect.Maps;
+
+public class TestTransactionPurchasesHiddenMarkovModel
+{
+	
+	private List<Product> createProducts(String category)
+	{
+		List<Product> products = Lists.newArrayList();
+		
+		for(int i = 0; i < 10; i++)
+		{
[…]

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/pdfs/customer/TestCustomerLocationPDF.java

@@ -1,51 +1,0 @@
-package com.github.rnowling.bps.datagenerator.pdfs.customer;
-
-import static org.junit.Assert.assertTrue;
-
-import java.util.ArrayList;
-import java.util.List;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.builders.store.StoreSamplerBuilder;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ZipcodeRecord;
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
-import com.github.rnowling.bps.datagenerator.datareaders.ZipcodeReader;
-import com.github.rnowling.bps.datagenerator.pdfs.customer.CustomerLocationPDF;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-
-public class TestCustomerLocationPDF
-{
-
-	@Test
-	public void testProbability() throws Exception
-	{
-		ZipcodeReader zipcodeReader = new ZipcodeReader();
-		zipcodeReader.setCoordinatesFile(Constants.COORDINATES_FILE);
-		zipcodeReader.setIncomesFile(Constants.INCOMES_FILE);
-		zipcodeReader.setPopulationFile(Constants.POPULATION_FILE);
-		List<ZipcodeRecord> zipcodes = zipcodeReader.readData();
-		
-		SeedFactory factory = new SeedFactory(1234);
-		
-		StoreSamplerBuilder storeSamplerBuilder = new StoreSamplerBuilder(zipcodes, factory);
-		Sampler<Store> storeSampler = storeSamplerBuilder.build();
-		
-		List<Store> stores = new ArrayList<Store>();
-		for(int i = 0; i < 10; i++)
-		{
-			Store store = storeSampler.sample();
-			stores.add(store);
-		}
-		
-		CustomerLocationPDF customerLocationPDF = new CustomerLocationPDF(zipcodes, stores, 
-					Constants.AVERAGE_CUSTOMER_STORE_DISTANCE);
-		
-		double prob = customerLocationPDF.probability(zipcodes.get(0));
-		
-		assertTrue(prob > 0.0);
-	}
-
-}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/pdfs/store/TestStoreLocationPopulationPDF.java

@@ -1,38 +1,0 @@
-package com.github.rnowling.bps.datagenerator.pdfs.store;
-
-import static org.junit.Assert.assertTrue;
-
-import java.util.List;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ZipcodeRecord;
-import com.github.rnowling.bps.datagenerator.datareaders.ZipcodeReader;
-import com.github.rnowling.bps.datagenerator.pdfs.store.StoreLocationIncomePDF;
-
-public class TestStoreLocationPopulationPDF
-{
-
-	@Test
-	public void testProbability() throws Exception
-	{
-		ZipcodeReader reader = new ZipcodeReader();
-		reader.setCoordinatesFile(Constants.COORDINATES_FILE);
-		reader.setIncomesFile(Constants.INCOMES_FILE);
-		reader.setPopulationFile(Constants.POPULATION_FILE);
-		
-		List<ZipcodeRecord> zipcodes = reader.readData();
-		
-		assertTrue(zipcodes.size() > 0);
-		
-		StoreLocationIncomePDF pdf = new StoreLocationIncomePDF(zipcodes, 100.0);
-		
-		for(ZipcodeRecord record : zipcodes)
-		{
-			assertTrue(pdf.probability(record) > 0.0);
-		}
-		
-	}
-
-}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/samplers/customer/TestCustomerSampler.java

@@ -1,66 +1,0 @@
-package com.github.rnowling.bps.datagenerator.samplers.customer;
-
-import static org.junit.Assert.assertNotNull;
-import static org.junit.Assert.assertTrue;
-
-import java.util.ArrayList;
-import java.util.List;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.builders.customer.CustomerSamplerBuilder;
-import com.github.rnowling.bps.datagenerator.builders.store.StoreSamplerBuilder;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.InputData;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.Names;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ZipcodeRecord;
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Customer;
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
-import com.github.rnowling.bps.datagenerator.datareaders.NameReader;
-import com.github.rnowling.bps.datagenerator.datareaders.ZipcodeReader;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-
-public class TestCustomerSampler
-{
-
-	@Test
-	public void testBuild() throws Exception
-	{
-		ZipcodeReader zipcodeReader = new ZipcodeReader();
-		zipcodeReader.setCoordinatesFile(Constants.COORDINATES_FILE);
-		zipcodeReader.setIncomesFile(Constants.INCOMES_FILE);
-		zipcodeReader.setPopulationFile(Constants.POPULATION_FILE);
-		List<ZipcodeRecord> zipcodes = zipcodeReader.readData();
-		
-		NameReader nameReader = new NameReader(Constants.NAMEDB_FILE);
-		Names names = nameReader.readData();
-		
-		InputData inputData = new InputData(zipcodes, names);
-		
-		SeedFactory factory = new SeedFactory(1234);
-		
-		StoreSamplerBuilder storeSamplerBuilder = new StoreSamplerBuilder(zipcodes, factory);
-		Sampler<Store> storeSampler = storeSamplerBuilder.build();
-		
-		List<Store> stores = new ArrayList<Store>();
-		for(int i = 0; i < 10; i++)
-		{
[…]

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/samplers/purchasingprofile/TestProductCategoryMarkovModelSampler.java

@@ -1,45 +1,0 @@
-package com.github.rnowling.bps.datagenerator.samplers.purchasingprofile;
-
-import static org.junit.Assert.assertNotNull;
-import static org.junit.Assert.assertTrue;
-
-import java.util.List;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ProductCategory;
-import com.github.rnowling.bps.datagenerator.datamodels.simulation.Product;
-import com.github.rnowling.bps.datagenerator.datareaders.ProductsReader;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.UniformSampler;
-
-public class TestProductCategoryMarkovModelSampler
-{
-
-	@Test
-	public void testSample() throws Exception
-	{
-		SeedFactory seedFactory = new SeedFactory(1245);
-		
-		ProductsReader reader = new ProductsReader(Constants.PRODUCTS_FILE);
-		
-		List<ProductCategory> productCategories = reader.readData();
-		
-		ProductCategory productCategory = productCategories.get(0);
-		
-		ProductCategoryMarkovModelSampler generator = new ProductCategoryMarkovModelSampler(productCategory, 
-				new UniformSampler(seedFactory), new UniformSampler(seedFactory), new UniformSampler(seedFactory)
-				);
-		
-		MarkovModel<Product> model = generator.sample();
-		
-		assertNotNull(model);
-		assertNotNull(model.getStartWeights());
-		assertNotNull(model.getTransitionWeights());
-		assertTrue(model.getStartWeights().size() > 0);
-		assertTrue(model.getTransitionWeights().size() > 0);
-	}
-
-}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/samplers/store/TestStoreSampler.java

@@ -1,56 +1,0 @@
-package com.github.rnowling.bps.datagenerator.samplers.store;
-
-import static org.junit.Assert.assertNotNull;
-import static org.junit.Assert.assertTrue;
-
-import java.util.List;
-import java.util.Map;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.Constants;
-import com.github.rnowling.bps.datagenerator.datamodels.inputs.ZipcodeRecord;
-import com.github.rnowling.bps.datagenerator.datamodels.outputs.Store;
-import com.github.rnowling.bps.datagenerator.datareaders.ZipcodeReader;
-import com.github.rnowling.bps.datagenerator.samplers.store.StoreSampler;
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.RouletteWheelSampler;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.SequenceSampler;
-import com.google.common.collect.Maps;
-
-public class TestStoreSampler
-{
-
-	@Test
-	public void testSampler() throws Exception
-	{
-		ZipcodeReader reader = new ZipcodeReader();
-		reader.setCoordinatesFile(Constants.COORDINATES_FILE);
-		reader.setIncomesFile(Constants.INCOMES_FILE);
-		reader.setPopulationFile(Constants.POPULATION_FILE);
-		
-		List<ZipcodeRecord> zipcodes = reader.readData();
-		
-		assertTrue(zipcodes.size() > 0);
-		
-		SeedFactory factory = new SeedFactory(1234);
-		
-		Map<ZipcodeRecord, Double> zipcodePDF = Maps.newHashMap();
-		for(ZipcodeRecord record : zipcodes)
-		{
-			zipcodePDF.put(record, 1.0);
-		}
-		
-		Sampler<Store> sampler = new StoreSampler(new SequenceSampler(), 
-				RouletteWheelSampler.create(zipcodePDF, factory));
-		
-		Store store = sampler.sample();
-		assertNotNull(store);
-		assertTrue(store.getId() >= 0);
-		assertNotNull(store.getName());
-		assertNotNull(store.getLocation());
-		
-	}
-
-}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/statistics/markovmodels/TestMarkovModelBuilder.java

@@ -1,62 +1,0 @@
-package com.github.rnowling.bps.datagenerator.statistics.markovmodels;
-
-import org.junit.Test;
-
-import static org.junit.Assert.assertThat;
-import static org.junit.Assert.assertEquals;
-import static org.junit.matchers.JUnitMatchers.*;
-
-import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModel;
-import com.github.rnowling.bps.datagenerator.statistics.markovmodels.MarkovModelBuilder;
-
-public class TestMarkovModelBuilder
-{
-
-	@Test
-	public void testAddStateState()
-	{
-		MarkovModelBuilder<String> builder = MarkovModelBuilder.create();
-		
-		builder.addStartState("a", 1.0);
-		
-		MarkovModel<String> msm = builder.build();
-		
-		assertThat(msm.getStartWeights().keySet(), hasItem("a"));
-		assertEquals((double) msm.getStartWeights().get("a"), (double) 1.0, 0.000001);
-		
-	}
-	
-	@Test
-	public void testAddEdgeTransition()
-	{
-		MarkovModelBuilder<String> builder = MarkovModelBuilder.create();
-		
-		builder.addTransition("a", "b", 1.0);
-		
-		MarkovModel<String> msm = builder.build();
-		
-		assertThat(msm.getTransitionWeights().rowKeySet(), hasItem("a"));
-		assertThat(msm.getTransitionWeights().columnKeySet(), hasItem("b"));
-		assertEquals((double) msm.getTransitionWeights().get("a", "b"), (double) 1.0, 0.000001);	
-	}
-	
-	@Test
-	public void testBuildMSM()
-	{
-		MarkovModelBuilder<String> builder = MarkovModelBuilder.create();
-		
-		builder.addStartState("a", 1.0);
-		builder.addTransition("a", "b", 1.0);
-		builder.addTransition("a", "c", 1.0);
-		
-		MarkovModel<String> msm = builder.build();
-		
-		assertThat(msm.getStartWeights().keySet(), hasItem("a"));
-		assertThat(msm.getTransitionWeights().rowKeySet(), hasItem("a"));
-		assertThat(msm.getTransitionWeights().columnKeySet(), hasItem("b"));
-		assertThat(msm.getTransitionWeights().columnKeySet(), hasItem("c"));
[…]

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/statistics/samplers/TestExponentialSampler.java

@@ -1,27 +1,0 @@
-package com.github.rnowling.bps.datagenerator.statistics.samplers;
-
-import static org.junit.Assert.assertTrue;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.ExponentialSampler;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-
-public class TestExponentialSampler
-{
-
-	@Test
-	public void testSample() throws Exception
-	{
-		double lambda = 1.0 / 2.0;
-		
-		SeedFactory seedFactory = new SeedFactory(1234);
-		
-		Sampler<Double> sampler = new ExponentialSampler(lambda, seedFactory);
-		
-		Double result = sampler.sample();
-		
-		assertTrue(result >= 0.0);
-	}
-}

### 2014-10-12 src/java/bps-data-generator/src/test/java/com/github/rnowling/bps/datagenerator/statistics/samplers/TestUniformIntSampler.java

@@ -1,46 +1,0 @@
-package com.github.rnowling.bps.datagenerator.statistics.samplers;
-
-import static org.junit.Assert.assertTrue;
-
-import org.junit.Test;
-
-import com.github.rnowling.bps.datagenerator.statistics.SeedFactory;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.Sampler;
-import com.github.rnowling.bps.datagenerator.statistics.samplers.UniformIntSampler;
-
-public class TestUniformIntSampler
-{
-
-	@Test
-	public void testSample() throws Exception
-	{
-		int upperbound = 10;
-		int lowerbound = 1;
-		
-		SeedFactory seedFactory = new SeedFactory(1234);
-		
-		Sampler<Integer> sampler = new UniformIntSampler(lowerbound, upperbound, seedFactory);
-		
-		Integer result = sampler.sample();
-		
-		assertTrue(result >= lowerbound);
-		assertTrue(result <= upperbound);
-	}
-	
-	@Test
-	public void testSampleInclusive() throws Exception
-	{
-		int upperbound = 2;
-		int lowerbound = 1;
-		
-		SeedFactory seedFactory = new SeedFactory(1234);
-		
-		Sampler<Integer> sampler = new UniformIntSampler(lowerbound, upperbound, seedFactory);
-		
-		Integer result = sampler.sample();
-		
-		assertTrue(result >= lowerbound);
-		assertTrue(result <= upperbound);
-	}
-
-}
